### 2025-03-13 manuscript_restrained_classification validation.Rmd
Updated Intro

---
title             : "Validation of quantifying infant all-day restrianed experiences at home using inertial sensors"
shorttitle        : "Quantify infant restrained time with sensors"
author: 
  - name          : "XX"
    affiliation   : "1"
    corresponding : XX    # Define only one corresponding author
    address       : "UC Riverside Department of Psychology, 900 University Avenue, Riverside, CA 92521"
    email         : "XX"
  - name          : "XX"
    affiliation   : "1"

authornote: |
  XX
abstract: |
  XX

keywords          : "infant, restraint, everyday experiences, inertial sensors, machine learning"
wordcount         : "XX"
bibliography      : "master.bib"
floatsintext      : no
linenumbers       : no
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
biblio-style      : "apa"
classoption       : "man"
output:
  papaja::apa6_pdf:
    citation_package: biblatex
header-includes:
  - \raggedbottom
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(papaja)
library(knitr)
library(patchwork)
library(scales)
library(ggforce)
library(hms)
library(tidyverse)
library(lubridate)
library(rstatix)
```

```{r analysis-preferences}
set.seed(2025)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
# pos_levels <- c("Supine", "Prone", "Sitting", "Upright", "Held")
# pal <-  c("#E69F00","#56B4E9", "#009E73","#F0E442", "#0072B2") %>% 
#   set_names(pos_levels)

# theme_update(text = element_text(size = 14),
#              axis.text.x = element_text(size = 14, color = "black"), 
#              axis.title.x = element_text(size = 16),
#              axis.text.y = element_text(size = 14,  color = "black"), 
#              axis.title.y = element_text(size = 16), 
#              panel.background = element_blank(),panel.border = element_blank(), 
#              panel.grid.major = element_blank(),
#              panel.grid.minor = element_blank(), axis.line = element_blank(), 
#              legend.key = element_rect(fill = "white")) 
```

During infancy, babies’ daily activities are largely decided by caregivers. When caregivers give infants freedom to move, infants have more autonomy over their activities, including practicing locomotor skills, exploring novel areas, and approach appealing targets. However, when caregivers restrain infants’ movement, infants have fewer choices about where they can go, what they can see or touch. Daily restraint practices include holding infants in arms and putting infants in devices like strollers, highchairs, car seats, exersaucers, etc. Being restrained has immediate impact on infant body movement, as well as cascading effect on infant language, visual, and whole-body learning experiences due to its constraint on infant learning context. Previous research has made endeavors to quantify infant daily restrained experiences, either by retrospective parental report or momentarily phone survey. However, both methods are subjected to report bias and lack time granularity. Therefore, the goal of this paper is to propose a new system utilizing inertial movement units (IMUs) to precisely quantify infant whole-day restrained experiences at home, overcoming previous methods’ limitations by collecting objective and continuous movement data.

<!-- I find it hard to connect between the two paragraphs. Also, I give up on subtitles, because I can’t say much about “temporary delay in motor development” -->

Restraint has effects on various domains of infant development. Previous research have found that restraint causes temporary delay in motor development (Abbott & Bartlett, 2001; Bartlett & Fanning, 2003; Carson et al., 2022; Pin et al., 2007). Infants need extensive practices on daily basis to master new motor skills (Adolph et al., 2012; XXX). But restraining infants deprive them of opportunities to practice motor skills or even move body parts (e.g., spine, legs, hip) (Siddicky et al., 2020). Beyond that, restraint alters infant visual learning experiences. For example, being seated in chairs probably limits walking infants’ visual field, but it may provide new visual experiences for infants who cannot sit straight (Franchak et al., 2018; Kretch et al., 2014; Luo & Franchak, 2020). Restraint also alters infant language learning experiences. For example, seated in devices is correlated to less consistent and less amount of adult language infants can hear (Malachowski et al., 2023). Additionally, restraint also alters infant manual exploration experiences, helping them to focus on a subgroup of objects from distractors (XXX). Therefore, given the extended effects of infant daily restrained experiences, it is important to quantify infants’ restrained experiences for whole-day long at home environment.

## Review of existing methods for measuring restraint

In the existing literature, retrospective parental report and ecological momentary assessment (EMA) are two commonly used methods to quantify infant daily restrained experiences. In most of the retrospective parental report studies, parents were asked to estimate the average daily restrained time of the infant (Abbott & Bartlett, 2001; Bartlett & Kneale Fanning, 2003; Callahan & Sisler, 1997; Carson et al., 2022; Hesketh et al., 2015; Karasik et al., 2018; Siddicky et al., 2020). Despite its widespread use, this method is high demanding for parents’ memory considering it is hard to track and remember every bout of restraint practice and then report an overall time estimation. There is an exception where researchers asked mothers to reflect on the previous day’s restraint practice every five minutes, which added more precision in estimation (Karasik et al., 2018). However, these methods are still sensitive to memory loss and thus not reliable in providing objective data. The recently developed method EMA lessens parents’ memory burden by making quick phone-survey prompts to parents a few times (e.g., 10 or 12 times) randomly throughout a day, asking infant body activities at the moment (Franchak, 2019; Franchak, Kadooka, et al., 2024; Malachowski et al., 2023). The total time proportion of infant being restrained during a day is then extrapolated based on the frequency of being restrained among all the prompts.

However, there are still shared limitations between the two methods. First, no matter whether the survey happens later or at the moment, they both rely on parental report, and thus are constrained by the parent’s presence with the infant. If the parent is systematically not aware of the infant’s certain activity (e.g., the caregiver is not at home in the afternoons), the report might not be fully representative of infant everyday experiences. Thus, a method not relied on parental report is needed. Second, both methods lack time granularity in quantifying infant daily restrained experiences. When is the infant restrained the most? Is there any time-sequence pattern in infant restrained experiences? These questions call for a method with higher time-resolution.

Measuring with IMUs has the promising potential to overcome these limitations. IMUs, light-weight sensors that can be tucked in special infant garments, can record infant body movement at home in an unobtrusive way. The sample rate of XX Hz allows for second-by-second observation of infant movement. The battery life of XX hours allows for whole-day long recording. An IMUs system, usually comprised of 3 to 4 sensors fixed in certain body places, records the fine-grained data of linear and angular acceleration for each sensor at every moment. The whole-day-long sensors’ data is impossible for humans to interpret and annotate painlessly. Thus, machine learning process is needed to help automatically classify movement categories. The IMUs system (i.e., sensors plus machine learning algorithms) has been validated in categorizing adults’ activities (XXX), children’s activities (XXX), and infants’ activities (XXX). However, no study before has validated the automatic classification of infant restrained status based on IMUs data, which is the primary goal of the current study.

## Current Study

Given the limitations in current methods in quantifying infant restrained experiences, and the potential of using sensors, the overall goal of this study is to evaluate the validity of using IMUs to automatically classify infant restrained status at home in a whole-day-long recording.

The current study uses the same system as in Franchak et al. (2024) with four IMUs embedded in infant leggings, two on knees and two on ankles. This IMUs system has been validated for classifying infant body position with accuracy of XXX. However, the automatic classification of infant restrained status remains to be a challenge. Consider an infant sitting on a couch independently versus an infant sitting on a highchair restrained; or an infant lying supine unrestrained versus an infant being tucked in cradle restrained; or an infant standing independently versus an infant being held in arms upright yet not moving; etc. Some forms of restrained activities might be hard to distinguish from unrestrained ones, if merely based on the infant’s knees and ankles’ movement data.

The IMUs system is tested by a total of XX sessions from XX infants who were aged from 7 to 14 months, covering a wide range from pre-sitting infants to walking infants. The sessions have an average recording time of XX hours (*SE* = XX). Sensors’ linear and angular acceleration at every moment is the raw data input, and human annotation of infant restrained status from the first 1.5-hour video is the ground truth data for the supervised machine learning classification.

To validate the IMUs system, we will first compare the supervised machine learning models’ prediction with human annotation results for the first 1.5-hour of the sessions. We will use the leave-one-out cross validation method (XXX): Leave one study session each time as test data and have the rest sessions as training data, and evaluate the model’s performance on the test data. Thus, the first aim of this study is to assess the models’ overall accuracy in predicting infant restrained status in the first 1.5-hour into the session. Afterwards, a new group model will be trained by feeding in all sessions’ human annotation results. We will then evaluate the group model’s performance on predicting infant whole-day-long restrained status, where no video is available. If the model preforms well, the estimation of the time proportion infant spent restrained should show similar pattern as previous studies. For example, infants spend less time restrained with age (XXX). Thus, the second aim of this study is to assess the group model’s prediction by testing the age-related change in infant restrained time proportion.

<!-- Haven't revised the following part -->
# Methods
## Participants
Our dataset is composed by infants from two non-overlapping age groups: XX infants (XX% female) from a younger group aged 4-7 months old and XX infants (XX% female) from an older group aged 11-14 months old. For each participant, we aimed to collect data monthly for consecutive four months. For example, we aimed to have younger infants participated when they were 4, 5, 6, and 7 months old). However, not all participants joined the study when they were 4 or 11 months old and some of them missed some sessions for various reasons, ending up with XX infants contributing to only one session and XX infants contributing to 2~4 sessions. We collected in total XX sessions from the younger group (age range from XX to XX mos) and XX sessions from the older group (age range from XX to XX mos). Caregivers reported the onset dates of the infant’s locomotor milestone in a structured survey (Adolph et al., 2003). We then categorized each session based on infants’ current locomotor status as non-sitters (n = XX), sitters (n = XX), crawlers (n = XX), and walkers (n = XX). See Figure XX for the whole data set distribution. Caregivers reported the infants’ race as as White (n = XX), African American (n = XX), Asian (n = XX), more than one (n = XX), or not fitting into the above races or chose not to answer (n = XX). There is no difference in sex, race, or ethnicity between the two age groups (ps >= XX). Families were recruited from Southern California via social media and community recruitment events, and received $30 for compensation for each visit. Caregivers consented before each session of the study. The whole procedures were approved by the University of California, Riverside Institutional Review Board.
<!-- split something here into different sections -->
At each session, we instructed caregivers to use recording devices without the presence of experimenters at the home. Four inertial movement units (IMUs) sewn in special leggings were used to record movement features (e.g., linear and rotational accelerations) of infants’ right thigh, left thigh, right ankle, and left ankle separately, at a sample rate of 62.5 Hz, for a whole day (Figure XX). A camera was used to record a 3-hour video at the beginning of each session of the infant’s activities for later human annotation.

## Procedure

## Data Processing

### Human Annotation of Restriant

### Machine Learning Classification of Restraint

### Validation Plan

# Results

## Data Sharing

## XXX

# Discussion

\newpage

# Declarations

## Funding

## Competing Interests

## Ethics Approval

## Consent

## Data, Material and Code availability

\newpage

<!-- # References -->

::: {#refs custom-style="Bibliography"}
:::

### 2025-04-21 manuscript_restrained_classification validation.qmd
wrote introduction

---
title: "Validation of quantifying infant all-day restrained experiences at home using inertial sensors"
blank-lines-above-title: 2
shorttitle: "Quantify infant restrained time with sensors"
fig-dpi: 300
execute: 
  echo: false
  warning: false
  error: false
  cache: false
author:
  - name: XX
    orcid: XX
    email: XX
    affiliations:
      - id: id1
        name: University of California, Riverside
        department: Department of Psychology
  - name: XX
    corresponding: true
    orcid: XX
    email: XX
    affiliations:
      - id: id1
        name: University of California, Riverside
        department: Department of Psychology

author-note:
  blank-lines-above-author-note: 1
  # Disclosures condensed to one paragraph, but you can start a field with two line breaks to break them up: \n\nNew Paragraph
  disclosures:
    # Example: This study was registered at ClinicalTrials.gov (Identifier NTC998877).
    study-registration: ~
    # Acknowledge and cite data/materials to be shared.
    data-sharing: ~
    # Example: This article is based on data published in Pulaski (2017).
    # Example: This article is based on the dissertation completed by Graham (2018).    
    # Example: Sally Jones earns royalties from the sale of Test X.
    conflict-of-interest: ~
    # Example: This study was supported by Grant A123 from the National Science Foundation.
    financial-support: ~
    # Example: The authors are grateful for the technical assistance of Dr. X during the initial design and setup of our lab equipment.
    gratitude: ~
    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
    authorship-agreements: ~
abstract: "XXX"
keywords: [infant, restraint, everyday experiences, inertial sensors, machine learning]
bibliography: [references.bib]
format:
  apaquarto-docx: 
    floatsintext: false
  apaquarto-html: default
  apaquarto-typst: default
  apaquarto-pdf: 
    # can be jou (journal), man (manuscript), stu (student), or doc (document)
    # for now, tables and figures do not render properly in jou mode. 
    documentmode: man
    # can be 10pt, 11pt, or 12pt
    fontsize: 12pt
    # Integrate tables and figures in text
    floatsintext: false
    # a4 paper if true, letter size if false
    a4paper: false
    # suppresses loading of the lmodern font package
    nolmodern: false
    # Suppresses loading of the fontenc package
    nofontenc: false
    # Suppress the title above the introduction
    donotrepeattitle: false
    # In jou mode, use times or pslatex instead of txfonts
    notxfonts: false
    # In jou mode, use Computer Modern font instead of times
    notimes: false
    # In jou mode, cancels automatic stretching of tables to column width 
    notab: false
    # Uses Helvetica font in stu and man mode
    helv: false
    # In man and stu mode, neutralizes the \helvetica command
    nosf: false
    # In man and stu mode, uses typewriter font
    tt: false
    # Puts draft watermark on first page
    draftfirst: false
    # Puts draft watermark on all pages
    draftall: false
    # Masks references that are marked as the author's own
    mask: false
    journal: ~
    volume: ~
    course: ~
    professor: ~
    duedate: ~
    # Hides correspondence text
    nocorrespondence: false
---

```{r}
#| include: false

library(papaja)
library(knitr)
library(patchwork)
library(scales)
library(ggforce)
library(hms)
library(tidyverse)
library(lubridate)
library(rstatix)
```

```{r}
#| include: false

set.seed(2025)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)

theme_update(text = element_text(size = 14),
             axis.text.x = element_text(size = 14, color = "black"),
             axis.title.x = element_text(size = 16),
             axis.text.y = element_text(size = 14,  color = "black"),
             axis.title.y = element_text(size = 16),
             panel.background = element_blank(),panel.border = element_blank(),
             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(), axis.line = element_blank(),
             legend.key = element_rect(fill = "white"))
```

Infants learn and develop through everyday experiences. One common yet underexplored aspect of infant everyday experiences is physical restraint. Restrained period is when infants are unable to change their body positions at will, such as when they are held by caregivers (e.g., in arms or in carriers) or placed in containment devices (e.g., strollers, highchairs, car seats, ExerSaucers). When unrestrained, infants have greater autonomy to explore, deciding where to go, and what to see, hear and touch. In contrast, restraint limits infants' ability to initiate movement and reduces their control over their exploration and thus learning input. However, at the same time, restraint can occasionally afford novel experiences that infants cannot otherwise access by themselves (e.g., going outdoors in a stroller). Because restraint helps shape the broader contexts in which infant learning takes place, it is critical to study infant restraint experiences in order to better understand when and how infant learning occurs and then inform practical caregiving guidance.

Restraint affects infant learning by altering infant daily experiences in many ways. First, restraint deprives infants of opportunities to move body parts (e.g., spine, legs, hip) [@Siddicky2020m], which is detrimental for their motor development. Indeed, prolonged restraint has been proven to cause temporary delay in motor development, even for ExerSaucers which are marketed as tools to promote motor development [@Abbott2001z; @Bartlett2003d; @Carson2022q; @Pin2007z; @Karasik2023l]. On the contrary, being unrestrained allows infants to produce XX steps per hour and fall XX times per day, which are both valuable practice for motor development (XXX). Second, restraint alters infant object interaction experience. Research found that infants hold objects less frequently when they are restrained compared to unrestrained, with restrained by caregivers providing the least opportunities [@Franchak2024v]. Third, restraint also alters infant language learning experiences. For example, seated in devices is correlated with less consistent and less amount of adult language infants can hear [@Malachowski2023z], potentially influencing language input and development.

However, restraint does not always reduce infant learning opportunities. Sometimes caregiver carrying infants or putting infants in devices can rescue infants from their body disadvantages, particularly during earlier stages of development. For example, young infants who cannot sit independently struggle to interactive with objects because they cannot visually examine and manually manipulate the object at the same time [@Soska2010q]. However, being seated in supportive devices helps stablize infant posture, free infant hands and allow for multimodal object interaction (XXX). Similarly, infant body height limits what they can see on their own. The situation is even worse for infants who cannot stand up because they see much fewer faces when they are in prone position [@Franchak2018i]. However, being held in a forward-facing carrier or seated in highchairs can elevate infants' viewpoint and increase opportunities for face-to-face interaction [@Yu2013h; @Kretch2015p; @Franchak2024d]. These examples demonstrate that restraint may provide infants with uncommon learning materials that they would not otherwise obtained due to their developmental constraint. These examples further imply that the effects of restraint may vary depending on infants’ age and motor abilities, reinforcing the need to precisely quantify restraint across developmental stages and examine how it changes with development.

Early research have made several attempts to quantify infant daily restrained time, but the estimates vary depending on the methods. Retrospective parental reports suggest relatively longer time of restraint. @Callahan1997s found infants younger than 5 months spent 5.7 hours per day (23.75%) on average in seating devices; @Bartlett2003d found that preterm infants at 8 months of age spent averaged 4.1 hours per day (17.08%) in containment equipment; @Siddicky2020m found that 2- to 6-month-olds spent 44% of daily *awake* time being held or carried. When recalculate based on infant average awake time (https://huckleberrycare.com/blog/baby-sleep-schedule-by-age-nap-and-sleep-chart; I will find the reference later), these findings suggest that infants from 2 to 8 months old may spend around 50% of their awake time restrained in devices, and nearly 50% of their awake time restrained by caregivers (i.e., carrying and holding). However, studies using ecological momentary assessment (EMA) found a smaller time amount of restraint. For example, @Malachowski2023z reported that 4- to 6-month-olds spent 41.13% of *total observation time* being held, and 45.43% in placement with physical support. The time proportion will presumably be less (and thus less than the above retrospective survey studies) when excluding infant sleep time. A more direct estimate comes from @Franchak2024v. They found that 10- to 13-month-olds spent 26.90% of *awake* time restrained in furniture or devices, and 18.79% restrained by caregivers, which is also lower than previous survey studies. In short, it appears that EMA gives lower estimates of infant daily restrained time compared to parental retrospective reports.

Does infant restrained time change with age? Only a handful of studeis have answered this question, but their findings are mixed as well. @Airaksinen2024z, using inertial measurement sensors (IMU) and machine learning algorithms, found that infants from Finland experienced *increased* holding and carrying time from 4 to 18 months of age. On the contrary, @Carson2022q, using parental retrospective report and longitudinal design, found that infants from Canada experienced *decreased* overall restrained time from 2 to 6 months of age. @Franchak2024v, using EMA and longitudinal design, found that infants from the US experienced *decreased* time held by caregivers from 10 to 13 months, yet the time restrained by furniture or devices did not change with age. The discrepancies in findings may be attributed to differences in the targeted behaviors (i.e., whether to include device restraint or not, and which devices to include), study design (i.e., cross-sectional vs. longitudinal), cultural contexts, as well as research methods. To address the current discrepancies in findings, the first step is to develop an accurate and reliable system that examines infant general restraint experiences, which can be further applied to different age ranges and culture contexts. Among the available methods, we propose IMU is the most objective, unobtrusive, and efficient way to caputure whole-day experiences in natural settings. But previous studies that have used IMU on infants only looked at holding behavior, excluding other extensive restrained practices such as device use [@Airaksinen2024z; @Franchak2024i]. Thus, a new customized machine learning algorithm is needed to identify a broader range of restraint experiences.

## Limitations of previous research methods

As illustrated above, survey is the most commonly used method for quantifying infants daily restrained experiences, either in the form of retrospective report or ecological momentary phone survey across days. In most of the retrospective parental report studies, parents were asked to estimate their infant's average daily restrained time over the past few weeks or month [@Abbott2001z; @Bartlett2003d; @Callahan1997s; @Carson2022q; @Hesketh2015q; @Karasik2018o; @Siddicky2020m]. Despite its widespread use, this method may not be precise because it highly demands parents’ memory—It is difficult to recall and sum every bout of restraint practice and then report an overall time estimation [@Bradburn1987g]. There are exceptions where researchers asked caregivers to complete structured daily diaries reflecting on the past day’s restraint practices in short intervals (e.g., every five minutes) [@Karasik2018o; @Majnemer2005g]. However, this method is still sensitive to memory loss and thus not reliable in providing objective data.

The recently developed method EMA lessens caregivers’ memory burden by making momentary phone-survey prompts to caregivers multiple times (e.g., 5, 10, or 12 times) randomly throughout a day [@Franchak2019v; @Franchak2024v; @Malachowski2023z]. Researchers then estimate the proportion of time infants spend restrained by calculating the frequency of restraint across all prompts. Presumably, EMA can improve accuracy compared to retrospective report. The estimate discrepancy between retrospective report and EMA that illustrated above indeed supports the potential difference between the two methods. However, the two methods still share crucial limitations. First, both retrospective reports and EMA rely on parental input, which assumes that caregivers are consistently or equally present to observe the infant throughout the day. If the infant’s certain activities are systematically unaware of by the caregiver (e.g., the caregiver is not at home in the afternoons), the report might not be fully representative of infant daily experiences. Thus, a method not relied on parental report is needed. Second, neither method provides time-sequence information with high granularity. They cannot answer finer-grained questions such as: When during the day is restraint most common? Are restraint bouts prolonged or brief and frequent? These questions call for a method with higher time-resolution.

Video recording and human annotation, as another method, offers more accurate and temporally detailed data than survey-based methods. However, this approach is labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies. As a result, most studies using video rely on short samples, normally around one hour (e.g., @Wang2025x). Accordingly, this method is sensitive to sampling bias because it is questionable for one-hour data to represent infant whole-day activities. Moreover, video recording can be obtrusive and may affect caregivers’ and infants’ behaviors, potentially introducing reactivity effects (XXX).

## Promises of IMU system

Measuring with IMU system has the potential to overcome the limitations listed above. IMU, light-weight sensors that can be tucked in specially designed garments (in our study, infant leggings), can record participants’ real-time movement of different body parts (e.g., hips and ankles). Easy synchronization between different sensors and local data storage eliminates the need for experimenters to collect data in person, making the method unobtrusive. A typical sample rate of 62.5 Hz allows for millisecond-level recording of motion data. A battery life of XX hours allows for whole-day-long observation, overcoming sampling bias to the greatest extent.

Each sensor collects raw data as accelerometer signals (in x, y, and z orientations) and gyroscope signals (in roll, pitch, and yaw orientations). By aggregating signals from a given time window, we can calculate customized motion features (statistical summaries like mean, skew, etc.) at every moment. These features can then be used as input for machine learning algorithms to automatically classify infant behaviors (e.g., restrained and unrestrained). Human annotation from video recordings serves as ground truth labels for model training. This pipeline has been validated in categorizing adults’ activities (XXX; I need to do more research on this and word it differently), children’s activities (XXX), and infants’ activities (Franchak et al., 2024 for infant body position). IMU have also been used to detect caregiver holding \[\@Airaksinen2024z\], but no prior studies have applied IMU to classify broader forms of restraint, including both holding and device restraining. Thus, we propose a new IMU system which collects objective, reliable, unobtrusive, continuous, and long-form infant movement data as well as automatically identifies infant restrained status.

## Current Study

Given the limitations in current methods and the potential of IMU technology, the overall goal of this study is to evaluate the validity of using IMUs to precisely quantify infant full-day restrained experiences at home. Despite previous validation of IMU system on identifying infant position and caregiver holding, identifying all restraint practices on infants is still challenging. Consider an infant sitting on a couch independently versus an infant sitting on a highchair restrained, or an infant lying supine unrestrained versus an infant being tucked in cradle restrained. The infants are in same body position and are very likely to show similar movement patterns, but the context is different, differentiating the restrained status. Furthermore, device restraint might be more difficult to detect than caregiver restraint because carrying and holding tend to generate dramatic movement (e.g., the infant moving in an adult speed). Thus, in the current study, we develop and validate the use of new machine-learning models that identify both caregiver restraint and device restraint, classifying infant moment-to-moment status as being restrained or unrestrained.

The models are tested by a total of 143 sessions from 66 infants who were aged from 7 to 14 months, covering a wide range from pre-sitting infants to walking infants. We sent caregivers special leggings with four IMUs embedded and a camera to record synchronized infant activities. Across the sessions, IMUs recorded an average of XX hours (*SE* = XX) of movement data, and the camera recorded the first 1.5 hours into the session. From the videos, human coders identify infant restrained status as ground truth for model training and testing. The **first aim** of the current study is to train the supervised machine-learning models of infant restrained status and evaluate the models’ accuracy using leave-one-out cross validation method (XXX) for the first 1.5 hours.

If the models perform well, we will use the same rationale to train a new group model by feeding in all sessions’ human annotation. This model will then be applied to the remaining, unannotated portions of the IMU recordings to estimate infants’ full-day restraint time. Unfortunately, neither did previous studies establish a clear convergent estimate of infant overall restrained time proportion, nor did they find consistent age-related change, probably due to method differences. Considering EMA is likely to be more accurate than parental retrospective reports, we will compare our current estimate with @Franchak2024v, who found that 10- to 13-month-old infants spent 45.69% awake time being restrained. We do not compare our estimate data with @Malachowski2023z, another EMA study, because their reported estimates are among total observation time rather than awake time. We will also compare our estimated age-related change with @Franchak2024v who found that infant experienced increased unrestrained time from 10 to 13 months. We do not compare our age-related change with @Airaksinen2024z, an IMU-based study, because they only looked at caregiver restraint and excluded device restraint. In short, our **second aim** is to validate our new system by comparing empirical results with a previous study. However, due to methodological differences again (IMU versus EMA), it remains an open question whether these estimates will converge. \[note, when I was writing here, I feel we should probably at least distinguish caregiver restraint and device restraint\]

# Method

### 2025-04-21 manuscript_restrained_classification validation.qmd
modified introduction

@@
 For example, young infants who cannot sit independently struggle to interactive with objects because they cannot visually examine and manually manipulate the object at the same time [@Soska2010q].
-However, being seated in supportive devices helps stablize infant posture, free infant hands and allow for multimodal object interaction (XXX).
+However, being seated in supportive devices helps stabilize infant posture, free infant hands and allow for multimodal object interaction (XXX).
 Similarly, infant body height limits what they can see on their own.
@@
 Does infant restrained time change with age?
-Only a handful of studeis have answered this question, but their findings are mixed as well.
+Only a handful of studies have answered this question, but their findings are mixed as well.
 @Airaksinen2024z, using inertial measurement sensors (IMU) and machine learning algorithms, found that infants from Finland experienced *increased* holding and carrying time from 4 to 18 months of age.
-On the contrary, @Carson2022q, using parental retrospective report and longitudinal design, found that infants from Canada experienced *decreased* overall restrained time from 2 to 6 months of age.
-@Franchak2024v, using EMA and longitudinal design, found that infants from the US experienced *decreased* time held by caregivers from 10 to 13 months, yet the time restrained by furniture or devices did not change with age.
+On the contrary, @Carson2022q, using parental retrospective report, found that infants from Canada experienced *decreased* overall restrained time from 2 to 6 months of age.
+@Franchak2019v, using EMA, found that infants from the US experienced *decreased* time held by caregivers from 3 to 12 months.
+The author further found 10- to 13-month-olds also experienced *decreased* held time, and *increased* overall unrestrained time [@Franchak2024v].
 The discrepancies in findings may be attributed to differences in the targeted behaviors (i.e., whether to include device restraint or not, and which devices to include), study design (i.e., cross-sectional vs.
@@
 To address the current discrepancies in findings, the first step is to develop an accurate and reliable system that examines infant general restraint experiences, which can be further applied to different age ranges and culture contexts.
-Among the available methods, we propose IMU is the most objective, unobtrusive, and efficient way to caputure whole-day experiences in natural settings.
+Among the available methods, we propose IMU is the most objective, unobtrusive, and efficient way to capture whole-day experiences in natural settings.
 But previous studies that have used IMU on infants only looked at holding behavior, excluding other extensive restrained practices such as device use [@Airaksinen2024z; @Franchak2024i].
@@
 
-Given the limitations in current methods and the potential of IMU technology, the overall goal of this study is to evaluate the validity of using IMUs to precisely quantify infant full-day restrained experiences at home.
+Given the limitations in current methods and the potential of IMU technology, the overall goal of this study is to evaluate the validity of using IMU to precisely quantify infant full-day restrained experiences at home.
 Despite previous validation of IMU system on identifying infant position and caregiver holding, identifying all restraint practices on infants is still challenging.
@@
 The models are tested by a total of 143 sessions from 66 infants who were aged from 7 to 14 months, covering a wide range from pre-sitting infants to walking infants.
-We sent caregivers special leggings with four IMUs embedded and a camera to record synchronized infant activities.
-Across the sessions, IMUs recorded an average of XX hours (*SE* = XX) of movement data, and the camera recorded the first 1.5 hours into the session.
+We sent caregivers special leggings with four IMU embedded and a camera to record synchronized infant activities.
+Across the sessions, IMU recorded an average of XX hours (*SE* = XX) of movement data, and the camera recorded the first 1.5 hours into the session.
 From the videos, human coders identify infant restrained status as ground truth for model training and testing.

### 2025-05-08 code_analyze_restrained.R
Delete code_analyze_restrained.R
Converted to qmd and chunked coded.

@@ -1,155 +1,0 @@
-library(tidyverse)
-library(timetk)
-library(here)
-library(rstatix)
-library(fastDummies)
-library(lme4)
-library(lmerTest)
-library(interactions)
-# i_am(".here")
-
-# preset
-theme_update(text = element_text(size = 20),
-             axis.text.x = element_text(size = 20, color = "black"), axis.title.x = element_text(size = 24),
-             axis.text.y = element_text(size = 20,  color = "black"), axis.title.y = element_text(size = 24), 
-             panel.background = element_blank(),panel.border = element_blank(), panel.grid.major = element_blank(),
-             panel.grid.minor = element_blank(), axis.line = element_blank(), 
-             legend.key = element_rect(fill = "white")) 
-# pos_levels <- c("Supine", "Prone", "Sitting", "Upright", "Held")
-# pal <-  c("#E69F00","#56B4E9", "#009E73","#F0E442", "#0072B2") %>% 
-  # set_names(pos_levels)
-
-# load data
-# synced_ppts <- read_csv("/Volumes/padlab/study_imu_long/code/project_status/project_dashboard.csv") %>% 
-#   mutate(id_uni = str_glue("{id}/{session}")) %>% 
-#   mutate(unique_id = paste0(id,"/",session)) %>% 
-#   filter(infant_unr == 2) %>% 
-#   filter (id != 104)
-
-# load data from local
-load("whole-model-prediction.RData")
-
-# function
-read_unrestrained <- function(temp_id) {
-  position <- read_csv(str_glue("/Volumes/padlab/study_imu_long/data/{temp_id}/synced_data/restrained_predictions_infant.csv"),
-                       show_col_types = FALSE) %>% 
-    mutate(unrestrained = factor(pos, levels = c("Unrestrained","Restrained")))
-  start_time <- position %>% slice_head %>% pull(time_start)
-  end_time <- position %>% slice_tail %>% pull(time_start)
-  position <- position %>% add_column(unique_id = temp_id, .before = 1) 
-}
-
-ds <- map_dfr(synced_ppts$unique_id, read_unrestrained)
-
[…]

### 2025-05-08 code_analyze_unrestrained.qmd
chunked and modified analysis code

---
title: "code_analyze_unrestrained"
author: "Hanzhi Wang"
format: html
editor: visual
---

## Import and preset

```{r}
#| echo: false
#| include: false
# load library
library(tidyverse)
library(timetk)
library(here)
library(rstatix)
library(fastDummies)
library(lme4)
library(lmerTest)
library(interactions)
```

```{r}
#| echo: false
#| include: false
# preset
theme_update(text = element_text(size = 20),
             axis.text.x = element_text(size = 20, color = "black"), axis.title.x = element_text(size = 24),
             axis.text.y = element_text(size = 20,  color = "black"), axis.title.y = element_text(size = 24), 
             panel.background = element_blank(),panel.border = element_blank(), panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(), axis.line = element_blank(), 
             legend.key = element_rect(fill = "white"))

loc_labels <- c("Non-Sitter", "Sitter", "Crawler", "Walker","Non-Crawler")
loc_colors <- c("tomato","chartreuse4","cyan3","mediumpurple1","orange") %>% set_names(loc_labels)
age_labels <- c("Younger","Older")
age_colors <- c("red","navy") %>% set_names(age_labels)

# function
read_restrained <- function(temp_id) {
  position <- read_csv(str_glue("/Volumes/padlab/study_imu_long/data/{temp_id}/synced_data/restrained_predictions_infant.csv"),
                       show_col_types = FALSE) %>% 
    mutate(unrestrained = factor(pos, levels = c("Unrestrained","Restrained")))
  start_time <- position %>% slice_head %>% pull(time_start)
  end_time <- position %>% slice_tail %>% pull(time_start)
  position <- position %>% add_column(unique_id = temp_id, .before = 1) 
}
```

```{r}
#| echo: false
# load data (1) from local
load("whole-model-prediction.RData")
```

```{r}
#| echo: false
#| eval: false
# OR load data (2) from server
synced_ppts <- read_csv("/Volumes/padlab/study_imu_long/code/project_status/project_dashboard.csv")%>% 
  mutate(id_uni = str_glue("{id}/{session}")) %>%
  mutate(unique_id = paste0(id,"/",session)) %>%
  filter(infant_unr == 2 |infant_unr == 1) %>% #have unrestrained and/or restrained predicted
  filter(unr_exported !=0) %>% # have human annotation
  filter (id != 104) %>% #104 is pilot data
  filter(!(id == 185 & session == 2)) #185/2 leggings were put on wrong baby
# id_info has session-level basic information
id_info <- synced_ppts %>% select(unique_id, id, agemo, session, sitter, crawler, walker) %>% 
  mutate(crawler = factor(crawler, levels = c("Non-Crawler", "Crawler")),
         walker = factor(walker, levels = c("Non-Walker", "Walker")),
         sitter = factor(sitter, levels = c("Non-Sitter","Sitter"))) %>% 
  mutate(age_group = factor(as.numeric(agemo < 9), levels = c(1,0), labels = c("Younger","Older"))) %>% 
  mutate(loc = ifelse(sitter=="Non-Sitter", "Non-Sitter", 
                      ifelse(crawler == "Non-Crawler", "Sitter",
                             ifelse(walker == "Non-Walker","Crawler","Walker")))) %>% 
  mutate(age_mo = floor(agemo+0.5))
id_info[id_info$unique_id=="10/1", "crawler"] <- "Non-Crawler"
id_info[id_info$unique_id=="10/1", "loc"] <- "Sitter"

# id_demo has other demographic information
id_demo <- synced_ppts %>%
  select(unique_id, id, agemo, infant_sex, infant_ethnicity, infant_race) %>% 
  mutate(age_group = factor(as.numeric(agemo < 9), levels = c(1,0), labels = c("Younger","Older"))) %>% 
  select(-agemo) %>% 
  distinct(id, .keep_all = TRUE)
id_demo[id_demo$id=="180", "infant_sex"] <- "male"

# ds has detailed prediction for every time window
ds <- map_dfr(synced_ppts$unique_id, read_restrained)
ds <- ds %>% left_join(id_info, by = "unique_id")

# calculate session-level data, and excluded unavailable time chunk
ds_sum <- ds %>% group_by(unique_id, agemo, walker, crawler,sitter, age_group,loc,age_mo) %>%
  filter(exclude_period == 0, nap_period == 0) %>%
  summarize(unrestrained_n = sum(as.numeric(unrestrained == "Unrestrained")),
            total_n = n()) %>%
  mutate(unrestrained_prop = unrestrained_n/total_n) %>% 
  mutate(restrained_prop = 1-unrestrained_prop) %>% 
  ungroup()

# outliers
# 178/3, younger min, 0.15
# 148/2, younger max, 0.96
# 181/3, older min, 0.75
outliers_list_older <- 
  DescTools::Outlier(ds_sum$restrained_prop[ds_sum$age_group=="Older"], method = "boxplot")
outliers_list_younger <- 
  DescTools::Outlier(ds_sum$restrained_prop[ds_sum$age_group=="Younger"], method = "boxplot")
ds_sum_excl <- 
  ds_sum[!(ds_sum$restrained_prop %in% outliers_list_older | ds_sum$restrained_prop %in% outliers_list_younger),]

# load human coding results
ds_human <- read_csv("Restrained human coding results.csv",
                     col_select = c("Subj","Session","Restrained Prop Agree","Restrained Kappa"),
                     na=c("/")) %>%
            mutate(unique_id = paste0(Subj,"/",Session,sep="")) %>% 
            drop_na()
ds_human <- left_join(ds_sum %>% select(unique_id),
                      ds_human, by="unique_id") %>% 
            select(-c("Subj",Session))
colnames(ds_human) <- c("unique_id","agree","kappa")

# save the data frames so that don't have to download data from server again
# save(ds, ds_sum, synced_ppts, id_info, id_demo, ds_human, ds_sum_excl, file = "whole-model-prediction.RData")
```

```{r}
#| echo: false
#| eval: false
# unmatched/missing sessions--all good now
full_file_names <- dir('predict_4_pos', full.names = F)
session_loocv_have <- str_replace(unique(unlist(str_extract_all(full_file_names, "\\d+_\\d+"))), "_", "/")
session_wholemodel_have <- ds_sum$unique_id
setdiff(session_loocv_have, session_wholemodel_have)
setdiff(session_wholemodel_have, session_loocv_have)
```

```{r}
#| echo: false
# demographic statistics
count(id_demo,infant_sex)
count(id_demo,infant_ethnicity)
count(id_demo,infant_race)
count(id_demo,age_group)
```

## Human coding rel

```{r}
# human coding reliability
library(psych)
(agree<-as.data.frame(describe(ds_human$agree)))
(kappa<-as.data.frame(describe(ds_human$kappa)))
```

## Data set

```{r}
# plot dataset
dataset <- ds_sum %>% 
  mutate(subj = str_split_i(ds_sum$unique_id, "/", 1)) %>% 
  add_count(subj, name = "sess_count") %>% 
  arrange(age_mo,sess_count)

ids <- data.frame(subj = dataset$subj)
ids <- ids[!duplicated(ids), , drop = FALSE] %>% 
  mutate(temp_id = row_number())

dataset <- dataset %>% 
  left_join(ids, by = "subj")

p_data <- ggplot(data=dataset,aes(x=age_mo, y=temp_id, color=factor(loc, levels=c("Non-Sitter", "Sitter", "Crawler","Walker"))))+
  geom_line(aes(group = temp_id), color = "black") +
  geom_point(shape=16, size=5, alpha=1)+
  scale_color_manual(values=loc_colors) +
  scale_x_continuous(breaks = seq(4, 14, by = 1))+
  scale_y_continuous(breaks = c(seq(1, 66, by = 1)),
                     labels = ids$subj) +
  labs(x="Age group (months)",
       y="Subject ID",
       color = "Locomotor status")+
  theme_update()
p_data
ggsave("figure/dataset.png",p_data, width = 10, height = 18)
ggsave("figure/dataset.eps",p_data, width = 10, height = 18)
```

## Total recording time

```{r}
# total recording time
ds_sum <- ds_sum %>% 
  mutate(total_time = total_n*15/60/60)

ds_sum %>%
  summarize(time_min = min(total_time),
            time_mean = mean(total_time), 
            time_sd = sd(total_time),
            time_median = median(total_time),
            time_max = max(total_time))

t.test(data=ds_sum, total_time~age_group)

p_ttltime <- ggplot(ds_sum, aes(x=total_time))+
  geom_histogram(color = "black",)+
  labs(x = "Total recording time (h)", y = "Count")+
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  scale_y_continuous(breaks = seq(0, 12, 2))
p_ttltime
ggsave("figure/total recording time dist.png",p_ttltime, width=10, height=10)
```

# Results

#Q1: Unrestrained time

```{r}
ggplot(ds_sum, aes(x = age_group, y = unrestrained_prop)) +
  geom_boxplot() + 
  geom_point(position = position_jitter(width = .1), size = 4, color = "red", shape = 1) + 
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
  labs(x="Age group",
       y="Unrestrained time %")
ggsave("figure/unrestrained time prop.png", width=6, height=5)

prop_un <- ds_sum %>% 
  group_by(age_group) %>% 
  summarize(min=min(unrestrained_prop),
            mean = mean(unrestrained_prop),
            sd = sd(unrestrained_prop),
            max = max(unrestrained_prop))
prop_un_excl <- ds_sum_excl %>% 
  group_by(age_group) %>% 
  summarize(min=min(unrestrained_prop),
            mean = mean(unrestrained_prop),
            sd = sd(unrestrained_prop),
            max = max(unrestrained_prop))
```

## Q2: Age-related change

```{r}
# across 4~14
ggplot(ds_sum) +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum))

ggplot(ds_sum_excl) +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change.png",width=6, height=4)
ggsave("figure/age_related_change.eps",width=6, height=4)
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl))

```

```{r}
# within younger group comparison
ds_sum %>% 
  filter(age_group=="Younger") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum %>% filter(age_group=="Younger")))

ds_sum_excl %>% 
  filter(age_group=="Younger") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change_young.png",width=6, height=4)
ggsave("figure/age_related_change_young.eps",width=6, height=4)
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl%>% filter(age_group=="Younger")))
```

```{r}
# within older group comparison
ds_sum %>% 
  filter(age_group=="Older") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum %>% filter(age_group=="Older")))

ds_sum_excl %>% 
  filter(age_group=="Older") %>% 
  ggplot() +
  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
ggsave("figure/age_related_change_old.png",width=6, height=4)
ggsave("figure/age_related_change_old.eps",width=6, height=4)
summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl%>% filter(age_group=="Older")))
```

```{r}
#| eval: false
#| echo: false
# re-do lm analysis using age_base and age_change
# For now, it would be too complicated to include age_base and age_change, so skip this distinction for now.
ds_sum2 <- ds_sum
ds_sum2 <- ds_sum2 %>% 
  mutate(id = map_chr(unique_id, ~ str_split_1(.x, "/")[1])) %>% 
  group_by(id) %>% 
  mutate(age_base = first(agemo)) %>% 
  ungroup()
ds_sum2$age_change <- ds_sum2$agemo-ds_sum2$age_base

summary(lm(unrestrained_prop ~ scale(age_base)+scale(age_change), data = ds_sum2))
summary(lm(unrestrained_prop ~ scale(age_base)+scale(age_change), data = ds_sum2 %>% filter(age_group=="Younger")))
summary(lm(unrestrained_prop ~ scale(age_base)+scale(age_change), data = ds_sum2 %>% filter(age_group=="Older")))

ds_sum_excl2 <- ds_sum_excl
ds_sum_excl2 <- ds_sum_excl2 %>% 
  mutate(id = map_chr(unique_id, ~ str_split_1(.x, "/")[1])) %>% 
  group_by(id) %>% 
  mutate(age_base = first(agemo)) %>% 
  ungroup()
ds_sum_excl2$age_change <- ds_sum_excl2$agemo-ds_sum_excl2$age_base

summary(lm(unrestrained_prop ~ scale(age_base)+scale(age_change), data = ds_sum_excl2))
summary(lm(unrestrained_prop ~ scale(age_base)+scale(age_change), data = ds_sum_excl2 %>% filter(age_group=="Younger"))) # there is a significant age-base change in younger group if excluding outliers
summary(lm(unrestrained_prop ~ scale(age_base)+scale(age_change), data = ds_sum_excl2 %>% filter(age_group=="Older")))
```

## Q3-1: sitting related change
```{r}
ds_sum %>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = sitter)) + 
    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = sitter), method = "lm", se = FALSE) + 
    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)
ggsave("figure/sit_related_change.png", width=6, height=4)
ggsave("figure/sit_related_change.eps", width=6, height=4)
summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum))
summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum_excl))

# ds_sum %>% 
#   filter(age_group=="Younger") %>% 
#   ggplot() +
#     geom_point(aes(x = agemo, y = unrestrained_prop, color = sitter)) + 
#     geom_smooth(aes(x = agemo, y = unrestrained_prop, color = sitter), method = "lm", se = FALSE) + 
#     scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
#     scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
#     scale_color_manual(values=loc_colors)
summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum %>% filter(age_group=="Younger")))
# summary(lm(unrestrained_prop ~ sitter + scale(age_base)+scale(age_change), data = ds_sum2 %>% filter(age_group=="Younger")))

```

## Q3-2: crawling related change
```{r}
ds_sum %>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = crawler)) + 
    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = crawler), method = "lm", se = FALSE) + 
    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)
ggsave("figure/crawl_related_change.png", width=6, height=4)
ggsave("figure/crawl_related_change.eps", width=6, height=4)
summary(lm(unrestrained_prop ~ crawler + scale(agemo), data = ds_sum))
summary(lm(unrestrained_prop ~ crawler + scale(agemo), data = ds_sum_excl))

```

## Q3-3: crawling + sitting
```{r}
ds_sum %>% 
  mutate(loc = ifelse(loc=="Walker","Crawler",loc)) %>% 
  ggplot() +
    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc)) + 
    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = loc), method = "lm", se = FALSE) + 
    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
    scale_color_manual(values=loc_colors)
ggsave("figure/crawl_sit_change.png", width=6, height=4)
ggsave("figure/crawl_sit_change.eps", width=6, height=4)
summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum))
summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum_excl))
# here, excluding outliers or not influences a lot
```

## Q4: unrestrained bouts--many short bouts or few long bouts?
```{r}
# create a bout level data frame
ds_bout <- ds %>% 
  mutate(unrestrained = recode(unrestrained, "Restrained" = "r", "Unrestrained" = "u"))%>% 
  mutate(unrestrained = ifelse(nap_period==1 | exclude_period==1, "e", as.character(unrestrained)))%>% 
  mutate(bouts = with(rle(unrestrained), rep(seq_along(values), lengths)))%>%
  group_by(bouts, unrestrained,unique_id) %>%
  summarise(start_time = first(time_start),
            end_time = last(time_start),
            .groups = "drop") %>% 
  mutate(end_time = end_time+15) %>% 
  mutate(dur = as.numeric(end_time - start_time)/60/60) %>% 
  filter(unrestrained != "e")

ds_bout_sess <- ds_bout %>% 
  group_by(unique_id,unrestrained) %>% 
  summarize(n_bout = n(),
            min = min(dur),
            median = median(dur),
            max = max(dur),
            .groups="drop") %>% 
  left_join(data.frame(unique_id = ds_sum$unique_id,
                      total_time = ds_sum$total_time,
                      age_mo = ds_sum$age_mo,
                      age_group=ds_sum$age_group,
                      loc = ds_sum$loc),
             by="unique_id") %>% 
  mutate(loc=factor(loc,levels=c("Non-Sitter", "Sitter", "Crawler","Walker"))) %>% 
  mutate(bout_h = n_bout/total_time)

ds_bout_wide <- ds_bout_sess %>% 
  select(-bout_h) %>% 
  pivot_wider(names_from = unrestrained, values_from = n_bout:max) %>% 
  mutate(total_bout = n_bout_r+n_bout_u)%>% 
  mutate(loc=factor(loc,levels=c("Non-Sitter", "Sitter", "Crawler","Walker"))) %>% 
  mutate(bout_h = total_bout/total_time)

# individual bout scatter 
for(i in unique(ds_bout$unique_id)){
  ggplot(data=ds_bout %>% filter(unique_id==i), aes(x=unique_id, y=dur, color=unrestrained))+
    geom_jitter(width=.05)
  ggsave(paste0("figure_bout/",str_replace(i, "/", "_"),".png", sep=""))
}
```

## Q4-1: bout number per hour, indicating transition frequency
```{r}
# first check, bout number per hour w/ recording time
ggplot(data=ds_bout_wide, aes(x=total_time, y=bout_h))+
  geom_point()+
  geom_smooth(method="lm")+
  labs(x="total recording time (h)", y="#bout per hour")
ggsave("figure/bout_hour_uniformity.png", width=6, height=4)
summary(lm(data=ds_bout_wide, bout_h~total_time))

# bout number per hour w/ age
p_b_age<-ggplot(data=ds_bout_wide,
       aes(x=factor(age_group,levels=c("Younger","Older")),y=bout_h,color=age_group))+
  geom_jitter(width=.1)+
  stat_summary(fun = mean, geom = "crossbar", width = 0.3, color = "black")+
  scale_color_manual(values=age_colors)+
  labs(x="", y="# bouts per hour")+
  theme(legend.position = "none")

# p_b_age2<-ggplot(data=ds_bout_wide,
#        aes(x=age_mo,y=bout_h))+
#   geom_jitter(width=.1,color="black")+
#   stat_summary(fun = mean, geom = "crossbar", width = 0.3, color = "red")+
#   labs(x="", y="# bouts per hour")+
#   theme(legend.position = "none")

# bout number per hour w/ loc status
p_b_loc <- ggplot(data=ds_bout_wide, aes(x=loc,y=bout_h,color=loc))+
  geom_jitter(width=.1)+
  stat_summary(fun = mean, geom = "crossbar", width = 0.3, color = "black")+
  scale_color_manual(values=loc_colors)+
  labs(x="", y="# bouts per hour")+
  theme(legend.position = "none")

ggpubr::ggarrange(p_b_age,p_b_loc, ncol = 2, nrow = 1, widths=c(4,8), heights=c(4))
ggsave("figure/bouts_per_hour.png", width=12, height=4)
ggsave("figure/bouts_per_hour.eps", width=12, height=4)

# summary(lm(data=ds_bout_wide, bout_h~age_mo))
# summary(lm(data=ds_bout_wide %>% filter(age_group=="Younger"), bout_h~age_mo))
# summary(lm(data=ds_bout_wide %>% filter(age_group=="Older"), bout_h~age_mo))
summary(lm(data=ds_bout_wide, bout_h~age_group))
summary(anova_result <- aov(data=ds_bout_wide, bout_h~loc))
TukeyHSD(anova_result)
# summary(lm(data=ds_bout_wide, bout_h~age_mo+loc))
# summary(lm(data=ds_bout_wide, bout_h~age_group+loc))
```

## Q4-2: bout median duration
```{r}
# bout number per hour w/ age
ds_bout_wide <- ds_bout_wide %>% 
  mutate(median = median_r+median_u)

p_bm_age1<-ggplot(data=ds_bout_wide,
       aes(x=factor(age_group,levels=c("Younger","Older")),y=median_r+median_u,color=age_group))+
  geom_jitter(width=.1)+
  stat_summary(fun = mean, geom = "crossbar", width = 0.3, color = "black")+
  scale_color_manual(values=age_colors)+
  labs(x="", y="Median Duration of Bouts (h)")+
  theme(legend.position = "none")

p_bm_loc1 <- ggplot(data=ds_bout_wide, aes(x=loc,y=median_r+median_u,color=loc))+
  geom_jitter(width=.1)+
  stat_summary(fun = mean, geom = "crossbar", width = 0.3, color = "black")+
  scale_color_manual(values=loc_colors)+
  labs(x="", y="Median Duration of Bouts (h)")+
  theme(legend.position = "none")

ggpubr::ggarrange(p_bm_age1,p_bm_loc1, ncol = 2, nrow = 1, widths=c(4,8), heights=c(4))
ggsave("figure/bouts_median.png", width=12, height=4)
ggsave("figure/bouts_median.eps", width=12, height=4)

summary(lm(data=ds_bout_wide, median~age_group))
summary(anova_result2 <- aov(data=ds_bout_wide, median~loc))
TukeyHSD(anova_result2)
```

## Q4-3: bout median dur unrestrained comp. restrained 
```{r}
# bout median w/ age
p_bm_age<-ds_bout_sess %>%
  mutate(ag_num=as.numeric(age_group),
         x_shift = case_when(
           unrestrained == "r" ~ ag_num - 0.3,
           unrestrained == "u" ~ ag_num + 0.3)) %>% 
  ggplot(aes(x=x_shift, y=median, color=age_group, shape = unrestrained))+
    geom_point(size=3)+
    geom_line(aes(group = unique_id), color = "black", alpha=.05)+
    # stat_summary(fun = mean, geom = "crossbar", width = 0.3, color = "black")+
    labs(x="", y ="Median Duration of Bouts (h)")+
    scale_x_continuous(breaks = c(1,2), 
                       labels = c("Younger","Older"))+
    scale_y_continuous(breaks=seq(0,0.6,0.1))+
    scale_color_manual(values=age_colors)+
    scale_shape_manual(values = c("r" = 1, "u" = 16))+
    theme(legend.position = "none")

t.test (x=ds_bout_sess$median[ds_bout_sess$unrestrained=="r"], 
        y=ds_bout_sess$median[ds_bout_sess$unrestrained=="u"], 
        paired=TRUE) # across groups, non-significant

t.test (x=ds_bout_sess$median[ds_bout_sess$unrestrained=="r" & ds_bout_sess$age_group=="Younger"], 
        y=ds_bout_sess$median[ds_bout_sess$unrestrained=="u" & ds_bout_sess$age_group=="Younger"],
        paired=TRUE) # Younger group, not significant

t.test (x=ds_bout_sess$median[ds_bout_sess$unrestrained=="r" & ds_bout_sess$age_group=="Older"], 
        y=ds_bout_sess$median[ds_bout_sess$unrestrained=="u" & ds_bout_sess$age_group=="Older"],
        paired=TRUE) # Older group, median duration of unrestrained time is significantly higher

# bout median w/ loc status
p_bm_loc<-ds_bout_sess %>%
  mutate(loc_num=as.numeric(loc),
         x_shift = case_when(
           unrestrained == "r" ~ loc_num - 0.15,
           unrestrained == "u" ~ loc_num + 0.15)) %>% 
  ggplot(aes(x=x_shift, y=median, color=loc, shape = unrestrained))+
    geom_point(size=3)+
    # geom_jitter(width=.1)+
    geom_line(aes(group = unique_id), color = "black", alpha=.05)+
    scale_shape_manual(values = c("r" = 1, "u" = 16)) +
    scale_color_manual(values=loc_colors)+
    scale_x_continuous(breaks = c(1,2,3,4), 
                       labels = c("Non-Sitter","Sitter","Crawler","Walker"))+
    labs(x="", y="Median Duration of Bouts (h)")

ggpubr::ggarrange(p_bm_age,p_bm_loc, ncol = 2, nrow = 1, widths=c(4,8), heights=c(4))
ggsave("figure/bouts_median_comp.png", width=12, height=4)
ggsave("figure/bouts_median_comp.eps", width=12, height=4)

# t.test(data=ds_bout_sess %>% filter(loc=="Non-Sitter"), median~unrestrained)
# t.test(data=ds_bout_sess %>% filter(loc=="Sitter"), median~unrestrained)
# t.test(data=ds_bout_sess %>% filter(loc=="Crawler"), median~unrestrained) # significant
# t.test(data=ds_bout_sess %>% filter(loc=="Walker"), median~unrestrained) # significant
# summary(aov(data=ds_bout_sess %>% filter(unrestrained=="r"), median~loc))
# summary(aov(data=ds_bout_sess %>% filter(unrestrained=="u"), median~loc))
# summary(aov(data=ds_bout_sess, median~loc*unrestrained))

for (l in loc_labels[1:4]){
  temp<- ds_bout_sess %>% filter(loc==l)
  print(l)
  print(t.test (x=temp$median[temp$unrestrained=="r"], 
          y=temp$median[temp$unrestrained=="u"], 
          paired=TRUE))
}

```

## Q5: dominant positions when unrestrained
```{r}
# need 30s-bin position data

```

### 2025-05-08 code_analyze_unrestrained.qmd
modified figure codes

@@
 # across 4~14
-ggplot(ds_sum) +
-  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
-  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
-  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
-  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
+# ggplot(ds_sum) +
+#   geom_point(aes(x = agemo, y = unrestrained_prop)) + 
+#   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
+#   scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
+#   scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
 summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum))
@@
 summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl))
+
+ggplot(ds_sum_excl) +
+  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
+  geom_smooth(aes(x = agemo, y = unrestrained_prop, group=age_group), method = "lm", se = T) + 
+  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
+ggsave("figure/age_related_change_ingroup.png",width=6, height=4)
+ggsave("figure/age_related_change_ingroup.eps",width=6, height=4)
 
@@
 ds_sum %>% 
-  mutate(loc = ifelse(loc=="Walker","Crawler",loc)) %>% 
+  mutate(loc = ifelse(loc=="Walker","Crawler",loc))%>%
+  mutate(loc = factor(loc, 
+                      levels = c("Non-Sitter","Sitter","Crawler"), 
+                      labels=c("Non-Sitter","Sitter","Crawler")))%>% 
   ggplot() +

### 2025-05-12 code_analyze_unrestrained.qmd
updated code

--- 2025-05-08 code_analyze_unrestrained.qmd
@@
 loc_labels <- c("Non-Sitter", "Sitter", "Crawler", "Walker","Non-Crawler")
-loc_colors <- c("tomato","chartreuse4","cyan3","mediumpurple1","orange") %>% set_names(loc_labels)
+loc_colors <- c("#D81B60","#004D40","#1E88E5","purple4","orange") %>% set_names(loc_labels)
 age_labels <- c("Younger","Older")
@@
 outliers_list_older <- 
-  DescTools::Outlier(ds_sum$restrained_prop[ds_sum$age_group=="Older"], method = "boxplot")
+  DescTools::Outlier(ds_sum$unrestrained_prop[ds_sum$age_group=="Older"], method = "boxplot")
 outliers_list_younger <- 
-  DescTools::Outlier(ds_sum$restrained_prop[ds_sum$age_group=="Younger"], method = "boxplot")
+  DescTools::Outlier(ds_sum$unrestrained_prop[ds_sum$age_group=="Younger"], method = "boxplot")
 ds_sum_excl <- 
-  ds_sum[!(ds_sum$restrained_prop %in% outliers_list_older | ds_sum$restrained_prop %in% outliers_list_younger),]
+  ds_sum[!(ds_sum$unrestrained_prop %in% outliers_list_older | ds_sum$unrestrained_prop %in% outliers_list_younger),]
+
+outliers_list <- 
+  DescTools::Outlier(ds_sum$unrestrained_prop, method = "boxplot")
+ds_sum_excl2 <- 
+  ds_sum[!(ds_sum$unrestrained_prop %in% outliers_list),]
 
@@
 # save the data frames so that don't have to download data from server again
-# save(ds, ds_sum, synced_ppts, id_info, id_demo, ds_human, ds_sum_excl, file = "whole-model-prediction.RData")
+# save(ds, ds_sum, synced_ppts, id_info, id_demo, ds_human, ds_sum_excl, ds_sum_excl2, file = "whole-model-prediction.RData")
 ```
@@
   geom_point(position = position_jitter(width = .1), size = 4, color = "red", shape = 1) + 
-  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
   labs(x="Age group",
@@
 ## Q2: Age-related change
+```{r}
+# descriptive results
+ds_sum %>% 
+  group_by(age_mo) %>% 
+  summarize(mean = mean(unrestrained_prop),
+            sd = sd(unrestrained_prop),
+            .groups="drop")
+```
 
@@
 # across 4~14
+# ggplot(ds_sum) +
+#   geom_point(aes(x = agemo, y = unrestrained_prop)) + 
+#   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
+#   scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+#   scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
+summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum))
+
 ggplot(ds_sum) +
@@
   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
-  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
-  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
-summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum))
+  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
+ggsave("figure/age_related_change.png",width=6, height=4)
+ggsave("figure/age_related_change.eps",width=6, height=4)
 
-ggplot(ds_sum_excl) +
+ggplot(ds_sum_excl2) +
   geom_point(aes(x = agemo, y = unrestrained_prop)) + 
   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
-  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
-  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
+  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
 ggsave("figure/age_related_change.png",width=6, height=4)
 ggsave("figure/age_related_change.eps",width=6, height=4)
-summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl))
+summary(lm(unrestrained_prop ~ agemo, data = ds_sum_excl2))
+
+ggplot(ds_sum_excl2) +
+  geom_point(aes(x = agemo, y = unrestrained_prop),size=4, shape=1, stroke=1.5) +
+  geom_smooth(aes(x = agemo, y = unrestrained_prop, group=age_group), method = "lm", se = T) +
+  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
+ggsave("figure/age_related_change_ingroup.png",width=6, height=4)
+ggsave("figure/age_related_change_ingroup.eps",width=6, height=4)
 
@@
   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
-  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
-  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
+  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
 summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum %>% filter(age_group=="Younger")))
 
-ds_sum_excl %>% 
+ds_sum_excl2 %>%
   filter(age_group=="Younger") %>%
@@
   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) +
-  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
-  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
+  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
 ggsave("figure/age_related_change_young.png",width=6, height=4)
 ggsave("figure/age_related_change_young.eps",width=6, height=4)
-summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl%>% filter(age_group=="Younger")))
+
+summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl2%>% filter(age_group=="Younger")))
 ```
@@
   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
-  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
-  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
+  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
 summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum %>% filter(age_group=="Older")))
 
-ds_sum_excl %>% 
+ds_sum_excl2 %>% 
   filter(age_group=="Older") %>% 
@@
   geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
-  scale_x_continuous(name = "Age (mos)", breaks = c(4,7,11,14), limits = c(3,15)) +
-  scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))
+  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
 ggsave("figure/age_related_change_old.png",width=6, height=4)
 ggsave("figure/age_related_change_old.eps",width=6, height=4)
-summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl%>% filter(age_group=="Older")))
+summary(lm(unrestrained_prop ~ scale(agemo), data = ds_sum_excl2%>% filter(age_group=="Older")))
 ```
@@
 
-## Q3-1: sitting related change
+## Q3 locomotor-related change: crawling + sitting( + walking)
 ```{r}
 ds_sum %>% 
+  mutate(loc = ifelse(loc=="Walker","Crawler",loc))%>%
+  mutate(loc = factor(loc, 
+                      levels = c("Non-Sitter","Sitter","Crawler"), 
+                      labels=c("Non-Sitter","Sitter","Crawler")))%>% 
   ggplot() +
-    geom_point(aes(x = agemo, y = unrestrained_prop, color = sitter)) + 
-    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = sitter), method = "lm", se = FALSE) + 
+    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc)) + 
     scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
-    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
+    scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
     scale_color_manual(values=loc_colors)
-ggsave("figure/sit_related_change.png", width=6, height=4)
-ggsave("figure/sit_related_change.eps", width=6, height=4)
-summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum))
-summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum_excl))
 
-# ds_sum %>% 
-#   filter(age_group=="Younger") %>% 
-#   ggplot() +
-#     geom_point(aes(x = agemo, y = unrestrained_prop, color = sitter)) + 
-#     geom_smooth(aes(x = agemo, y = unrestrained_prop, color = sitter), method = "lm", se = FALSE) + 
-#     scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
-#     scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
-#     scale_color_manual(values=loc_colors)
-summary(lm(unrestrained_prop ~ sitter + scale(agemo), data = ds_sum %>% filter(age_group=="Younger")))
-# summary(lm(unrestrained_prop ~ sitter + scale(age_base)+scale(age_change), data = ds_sum2 %>% filter(age_group=="Younger")))
+# sitting + crawling
+ds_sum_excl2 %>% 
+  mutate(loc = ifelse(loc=="Walker","Crawler",loc))%>%
+  mutate(loc = factor(loc, 
+                      levels = c("Non-Sitter","Sitter","Crawler"), 
+                      labels=c("Non-Sitter","Sitter","Crawler")))%>% 
+  ggplot() +
+    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc)) + 
+    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
+    scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
+    scale_color_manual(values=loc_colors)
 
+# sitting + crawling + walking
+mean_unr_loc <- ds_sum_excl2 %>% 
+  group_by(loc) %>% 
+  summarize(age = mean(agemo),
+            prop = mean(unrestrained_prop),
+            .groups="drop") %>% 
+  mutate(x=age-0.8,
+         xend=age+0.8,
+         y=prop,
+         yend=prop,)
+
+ds_sum_excl2 %>% 
+  ggplot() +
+    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc),size=4,shape=1, stroke=1.5) + 
+    scale_x_continuous(name = "Age (mon)", breaks = c(4,7,11,14), limits = c(3,15)) +
+    scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
+    scale_color_manual(values=loc_colors)
+ggsave("figure/sit_crawl_walk_change.png", width=8, height=4)
+ggsave("figure/sit_crawl_walk_change.eps", width=8, height=4)
+# summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum))
+# summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum_excl))
+
+summary(lm(unrestrained_prop ~ agemo, data = ds_sum_excl2))
+summary(lm(unrestrained_prop ~ sitter + agemo, data = ds_sum_excl2))
+summary(lm(unrestrained_prop ~ sitter + crawler + agemo, data = ds_sum_excl2))
+summary(lm(unrestrained_prop ~ sitter + crawler + walker + agemo, data = ds_sum_excl2))
 ```
 
-## Q3-2: crawling related change
 ```{r}
-ds_sum %>% 
-  ggplot() +
-    geom_point(aes(x = agemo, y = unrestrained_prop, color = crawler)) + 
-    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = crawler), method = "lm", se = FALSE) + 
-    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
-    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
-    scale_color_manual(values=loc_colors)
-ggsave("figure/crawl_related_change.png", width=6, height=4)
-ggsave("figure/crawl_related_change.eps", width=6, height=4)
-summary(lm(unrestrained_prop ~ crawler + scale(agemo), data = ds_sum))
-summary(lm(unrestrained_prop ~ crawler + scale(agemo), data = ds_sum_excl))
+# description
+t <- ds_sum_excl2 %>% 
+  group_by(loc) %>% 
+  summarize(mean=mean(unrestrained_prop),
+            sd=sd(unrestrained_prop))
 
-```
-
-## Q3-3: crawling + sitting
-```{r}
-ds_sum %>% 
-  mutate(loc = ifelse(loc=="Walker","Crawler",loc)) %>% 
-  ggplot() +
-    geom_point(aes(x = agemo, y = unrestrained_prop, color = loc)) + 
-    geom_smooth(aes(x = agemo, y = unrestrained_prop, color = loc), method = "lm", se = FALSE) + 
-    scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
-    scale_y_continuous(name = "Prop(unrestrained)", breaks = seq(0, 1, .25), limits = c(0,1))+
-    scale_color_manual(values=loc_colors)
-ggsave("figure/crawl_sit_change.png", width=6, height=4)
-ggsave("figure/crawl_sit_change.eps", width=6, height=4)
-summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum))
-summary(lm(unrestrained_prop ~ crawler + sitter + agemo, data = ds_sum_excl))
-# here, excluding outliers or not influences a lot
 ```

### 2025-05-12 code_analyze_time-series.R
Delete code_analyze_time-series.R
move to new repo

@@ -1,90 +1,0 @@
-# load library
-library(tidyverse)
-library(here)
-library(lubridate)
-library(janitor)
-library(scales)
-library(hms)
-
-# pre-set
-theme_update(text = element_text(size = 14),
-             plot.title = element_text(size = 12),
-             axis.text.x = element_text(size = 14, color = "black"), axis.title.x = element_text(size = 16),
-             axis.text.y = element_text(size = 12,  color = "black"), axis.title.y = element_text(size = 16), 
-             panel.background = element_blank(),panel.border = element_blank(), panel.grid.major = element_blank(),
-             panel.grid.minor = element_blank(), axis.line = element_blank(), 
-             legend.key = element_rect(fill = "white")) 
-lims <- as_hms(c('09:00:00', '21:59:00'))
-hour_breaks = as_hms(c('09:00:00', '10:00:00', '11:00:00', '12:00:00', '13:00:00', '14:00:00', '15:00:00', '16:00:00', '17:00:00', '18:00:00', '19:00:00', '20:00:00', '21:00:00'))
-label_breaks = c("9am","","11am","","1pm","","3pm","","5pm","","7pm","","9pm")
-pos_levels <- c("Supine", "Prone", "Sitting", "Upright", "Held")
-pal <-  c("#E69F00","#56B4E9", "#009E73","#F0E442", "#0072B2") %>% 
-  set_names(pos_levels)
-
-loc_levels <- c("Crawling", "Walking")
-loc_pal <-  c("brown","blue") %>% 
-  set_names(loc_levels)
-
-unrestrained_levels <- c("Yes", "No")
-unrestrained_pal <-  c("purple","white") %>% 
-  set_names(unrestrained_levels)
-
-# load data
-ds <- read_csv("!import/lena-imu-compiled.csv")
-
-# data wrangling
-ds <- ds %>% 
-  mutate(id_uni = factor(ds$id*100+ds$session)) %>% 
-  select(-c(roll.key:Bin.Mins)) %>% 
-  rename_with(janitor::make_clean_names)
-
-ds$age_centered <- scale(ds$age, scale = F)
-ds$age_group <- factor(as.numeric(ds$age > 9), levels = 0:1, labels = c("Younger","Older"))
-
-ds <- ds %>% mutate(across(sit_time:restrained_time, ~ ifelse(nap_time > 0 | exclude_time > 0, 0, .x))) %>% 
[…]

### 2025-08-27 code_compare_models.R
used check_predictions_restraint_loocv instead

@@ -1,205 +1,0 @@
-# load packages
-library(tidyverse)
-library(dplyr)
-library(here)
-
-# update theme
-theme_update(text = element_text(size = 14),
-             axis.text.x = element_text(size = 14, color = "black"),
-             axis.title.x = element_text(size = 16),
-             axis.text.y = element_text(size = 14,  color = "black"),
-             axis.title.y = element_text(size = 16),
-             panel.background = element_blank(),panel.border = element_blank(),
-             panel.grid.major = element_blank(),
-             panel.grid.minor = element_blank(), axis.line = element_blank(),
-             legend.key = element_rect(fill = "white"))
-
-# load metrics from csv
-ds_30_pos <- read_csv("metrics_restrained_30_pos.csv") %>% 
-  mutate(id = ifelse(str_detect(id, "-"), format(as.Date(id, format="%d-%b"),"%m/%d"),id)) %>% 
-  arrange(id) %>% 
-  rename(id=id, acc=overall_acc, sens=overall_sens, spec=overall_spec, posi=overall_posi_pred) %>% 
-  mutate(sens = 1-sens, spec = 1-spec, posi = 1-posi)
-
-ds_30_no_pos <- read_csv("metrics_restrained_30_no_pos.csv") %>% 
-  mutate(id = ifelse(str_detect(id, "-"), format(as.Date(id, format="%d-%b"),"%m/%d"),id)) %>% 
-  arrange(id)%>% 
-  rename(id=id, acc=overall_acc, sens=overall_sens, spec=overall_spec, posi=overall_posi_pred) %>% 
-  mutate(sens = 1-sens, spec = 1-spec, posi = 1-posi)
-
-ds_4_pos <- read_csv("metrics_restrained_4_pos.csv") %>% 
-  mutate(id = ifelse(str_detect(id, "-"), format(as.Date(id, format="%d-%b"),"%m/%d"),id)) %>% 
-  arrange(id)%>% 
-  rename(id=id, acc=overall_acc, sens=overall_sens, spec=overall_spec, posi=overall_posi_pred) %>% 
-  mutate(sens = 1-sens, spec = 1-spec, posi = 1-posi)
-
-ds_4_no_pos <- read_csv("metrics_restrained_4_no_pos.csv") %>% 
-  mutate(id = ifelse(str_detect(id, "-"), format(as.Date(id, format="%d-%b"),"%m/%d"),id)) %>% 
-  arrange(id)%>% 
[…]

### 2025-08-27 code_compare_models_4_16_30.R

@@ -1,199 +1,0 @@
-# load packages
-library(tidyverse)
-library(dplyr)
-library(here)
-library(car)
-# update theme
-theme_update(text = element_text(size = 14),
-             axis.text.x = element_text(size = 14, color = "black"),
-             axis.title.x = element_text(size = 16),
-             axis.text.y = element_text(size = 14,  color = "black"),
-             axis.title.y = element_text(size = 16),
-             panel.background = element_blank(),panel.border = element_blank(),
-             panel.grid.major = element_blank(),
-             panel.grid.minor = element_blank(), axis.line = element_blank(),
-             legend.key = element_rect(fill = "white"))
-
-# load metrics from csv
-ds_30_no_pos <- read_csv("metrics_restrained_30_no_pos.csv") %>% 
-  mutate(id = ifelse(str_detect(id, "-"), format(as.Date(id, format="%d-%b"),"%m/%d"),id)) %>% 
-  arrange(id)%>% 
-  rename(id=id, acc=overall_acc, sens=overall_sens, spec=overall_spec, posi=overall_posi_pred) %>% 
-  mutate(sens = 1-sens, spec = 1-spec, posi = 1-posi)
-
-ds_4_no_pos <- read_csv("metrics_restrained_4_no_pos.csv") %>% 
-  mutate(id = ifelse(str_detect(id, "-"), format(as.Date(id, format="%d-%b"),"%m/%d"),id)) %>% 
-  arrange(id)%>% 
-  rename(id=id, acc=overall_acc, sens=overall_sens, spec=overall_spec, posi=overall_posi_pred) %>% 
-  mutate(sens = 1-sens, spec = 1-spec, posi = 1-posi)
-
-ds_16_no_pos <- read_csv("metrics_restrained_16_no_pos.csv") %>% 
-  mutate(id = ifelse(str_detect(id, "-"), format(as.Date(id, format="%d-%b"),"%m/%d"),id)) %>% 
-  arrange(id)%>% 
-  rename(id=id, acc=overall_acc, sens=overall_sens, spec=overall_spec, posi=overall_posi_pred) %>% 
-  mutate(sens = 1-sens, spec = 1-spec, posi = 1-posi)
-
-
-# human coded 146 sessions. There are only 139 sessions for 30s models, and 126 sessions for 4s models.
-# check which sessions are missing
-# load human coding results
[…]

### 2025-08-27 check_predictions_restraint_bulk_loocv.R
Copied the code from server and modified it to calculate the performance matrix

@@ -1,0 +1,29 @@
+# Calculate performance matrices for all models across sessions
+# Called "check_predictions_restraint_loocv.R" and data on the server
+library(tidyverse)
+library(rstatix)
+library(here)
+library(cvms)
+library(janitor)
+i_am(".here")
+
+
+source(here("code","project_status","check_predictions_restraint_loocv.R"))
+synced_ppts <- read_csv(here("code","project_status","project_dashboard.csv")) %>% 
+  filter(infant_synced_unr > 0) %>% 
+  filter(!(sessions_dir %in%  c("/Volumes/padlab/study_imu_long/data/104/1",
+                                "/Volumes/padlab/study_imu_long/data/185/2",
+                                "/Volumes/padlab/study_imu_long/data/185/3")))
+# "/Volumes/padlab/study_imu_long/data/148/1",
+# "/Volumes/padlab/study_imu_long/data/178/3")))
+# "/Volumes/padlab/study_imu_long/data/132/3")))
+
+# Calculate performance matrices
+matrix_30 <- 
+  map2_dfr(synced_ppts$id, synced_ppts$session, ~ check_predictions_restraint_loocv(model = "30_no_pos", .x, .y))
+matrix_16 <- 
+  map2_dfr(synced_ppts$id, synced_ppts$session, ~ check_predictions_restraint_loocv(model = "16_no_pos", .x, .y))
+matrix_4 <- 
+  map2_dfr(synced_ppts$id, synced_ppts$session, ~ check_predictions_restraint_loocv(model = "4_no_pos", .x, .y))
+
+save(matrix_4, matrix_16, matrix_30, file = "matrices.RData")

### 2025-08-27 check_predictions_restraint_loocv.R

@@ -1,0 +1,35 @@
+# Calculate performance matrix for each model on each session based on loocv predictions
+check_predictions_restraint_loocv <- function(model, id, session) {
+  require(tidyverse)
+  require(here)
+  require(cvms)
+  i_am(".here")
+
+  code_file = str_glue("/Volumes/padlab/study_imu_long/restraint/1 validation/model prediction/predict_{model}/{id}_{session}_actual.csv") #human coding after filtering out NAs
+  pos_file = str_glue("/Volumes/padlab/study_imu_long/restraint/1 validation/model prediction/predict_{model}/{id}_{session}_predict.csv") #loocv prediction. both files should have same length
+  
+  rm(code,predictions)
+  code =read_csv(code_file)
+  ref1 <- str_match(code$pool[1],"\\[\\\"(.*?)\\\"")[,2]
+  code <- code %>% 
+    mutate(code = ifelse((ref1=="Unrestrained" & ref == 1 | ref1=="Restrained" & ref == 2), "u","r"))
+  
+  predictions = read_csv(pos_file)
+  ref2 <- str_match(predictions$pool[1],"\\[\\\"(.*?)\\\"")[,2]
+  predictions <- predictions %>% 
+    mutate(pos = ifelse((ref2=="Unrestrained" & ref == 1 | ref2=="Restrained" & ref == 2), "u","r"))
+  
+  ds <- data.frame(code = code$code, pos = predictions$pos) %>% 
+    mutate(code = factor(code, levels=c("u","r")),
+           pos = factor(pos, levels=c("u","r")))
+  
+  res_u <- confusion_matrix(ds$code, ds$pos, positive="u")
+  res_r <- confusion_matrix(ds$code, ds$pos, positive="r")
+
+  out <- rbind(res_u, res_r) %>% 
+    select(`Positive Class`:`Prevalence`) %>% 
+    mutate(model = model, id = id, session = session) %>% 
+    select(model, id, session, everything())
+  
+  return(out)
+}

### 2025-08-27 manuscript_restrained_classification validation.qmd
Updated result analysis and writing framework

--- 2025-04-21 manuscript_restrained_classification validation.qmd
@@
 #| include: false
-
 library(papaja)
@@
 library(rstatix)
+library(janitor)
+library(flextable)
+library(effectsize)
 ```
@@
              legend.key = element_rect(fill = "white"))
+
+model_labels <- c("4_no_pos", "16_no_pos", "30_no_pos")
+model_shapes <- c(5,1,2) %>% set_names(model_labels)
+class_labels <- c("r", "u")
+class_shapes <- c(5,1) %>% set_names(class_labels)
+class_colors <- c("#0072B2","#E69F00") %>% set_names(class_labels)
+
+indx <- c("balanced_accuracy","sensitivity","pos_pred_value","f1","prevalence_df")
+indices <- c("Sensitivity", "Positive Predictive Value", "F1", "Prevalence Difference")
+
 ```
 
-Infants learn and develop through everyday experiences.
-One common yet underexplored aspect of infant everyday experiences is physical restraint.
-Restrained period is when infants are unable to change their body positions at will, such as when they are held by caregivers (e.g., in arms or in carriers) or placed in containment devices (e.g., strollers, highchairs, car seats, ExerSaucers).
-When unrestrained, infants have greater autonomy to explore, deciding where to go, and what to see, hear and touch.
-In contrast, restraint limits infants' ability to initiate movement and reduces their control over their exploration and thus learning input.
-However, at the same time, restraint can occasionally afford novel experiences that infants cannot otherwise access by themselves (e.g., going outdoors in a stroller).
-Because restraint helps shape the broader contexts in which infant learning takes place, it is critical to study infant restraint experiences in order to better understand when and how infant learning occurs and then inform practical caregiving guidance.
+# Results
+We first compared the model’s performance to human annotation of the windowed data set of the first 1.5 hours across all 142 sessions.
+We then compared the model’s estimate to previous works regarding infants’ full day restrained time and developmental change.
 
-Restraint affects infant learning by altering infant daily experiences in many ways.
-First, restraint deprives infants of opportunities to move body parts (e.g., spine, legs, hip) [@Siddicky2020m], which is detrimental for their motor development.
-Indeed, prolonged restraint has been proven to cause temporary delay in motor development, even for ExerSaucers which are marketed as tools to promote motor development [@Abbott2001z; @Bartlett2003d; @Carson2022q; @Pin2007z; @Karasik2023l].
-On the contrary, being unrestrained allows infants to produce XX steps per hour and fall XX times per day, which are both valuable practice for motor development (XXX).
-Second, restraint alters infant object interaction experience.
-Research found that infants hold objects less frequently when they are restrained compared to unrestrained, with restrained by caregivers providing the least opportunities [@Franchak2024v].
-Third, restraint also alters infant language learning experiences.
-For example, seated in devices is correlated with less consistent and less amount of adult language infants can hear [@Malachowski2023z], potentially influencing language input and development.
+## Model Performance Compared to Human Annotation
+```{r}
+load("matrices.RData")
+matrix_comb <- rbind(matrix_30, matrix_16, matrix_4)
+matrix_comb <- matrix_comb %>%  
+  clean_names() %>% 
+  mutate(model = factor(model, levels=c("4_no_pos","16_no_pos","30_no_pos")),
+         positive_class = factor(positive_class, levels=c('r','u'))) %>% 
+  mutate(prevalence_df = detection_prevalence - prevalence)
 
-However, restraint does not always reduce infant learning opportunities.
-Sometimes caregiver carrying infants or putting infants in devices can rescue infants from their body disadvantages, particularly during earlier stages of development.
-For example, young infants who cannot sit independently struggle to interactive with objects because they cannot visually examine and manually manipulate the object at the same time [@Soska2010q].
-However, being seated in supportive devices helps stablize infant posture, free infant hands and allow for multimodal object interaction (XXX).
-Similarly, infant body height limits what they can see on their own.
-The situation is even worse for infants who cannot stand up because they see much fewer faces when they are in prone position [@Franchak2018i].
-However, being held in a forward-facing carrier or seated in highchairs can elevate infants' viewpoint and increase opportunities for face-to-face interaction [@Yu2013h; @Kretch2015p; @Franchak2024d].
-These examples demonstrate that restraint may provide infants with uncommon learning materials that they would not otherwise obtained due to their developmental constraint.
-These examples further imply that the effects of restraint may vary depending on infants’ age and motor abilities, reinforcing the need to precisely quantify restraint across developmental stages and examine how it changes with development.
+# Descriptives
+stats <- matrix_comb %>% 
+  group_by(model, positive_class) %>% 
+  summarize (mean_acc = mean(balanced_accuracy, na.rm=T),
+             sd_acc = sd(balanced_accuracy, na.rm=T),
+             mean_sen = mean(sensitivity, na.rm=T),
+             sd_sen = sd(sensitivity, na.rm=T),
+             mean_ppv = mean(pos_pred_value, na.rm=T),
+             sd_ppv = sd(pos_pred_value, na.rm=T),
+             mean_f1 = mean(f1, na.rm=T),
+             sd_f1 = sd(f1, na.rm=T),
+             mean_code_prev = mean(prevalence, na.rm=T),
+             mean_pred_prev = mean(detection_prevalence, na.rm=T),
+             .groups="drop")
 
-Early research have made several attempts to quantify infant daily restrained time, but the estimates vary depending on the methods.
-Retrospective parental reports suggest relatively longer time of restraint.
-@Callahan1997s found infants younger than 5 months spent 5.7 hours per day (23.75%) on average in seating devices; @Bartlett2003d found that preterm infants at 8 months of age spent averaged 4.1 hours per day (17.08%) in containment equipment; @Siddicky2020m found that 2- to 6-month-olds spent 44% of daily *awake* time being held or carried.
-When recalculate based on infant average awake time (https://huckleberrycare.com/blog/baby-sleep-schedule-by-age-nap-and-sleep-chart; I will find the reference later), these findings suggest that infants from 2 to 8 months old may spend around 50% of their awake time restrained in devices, and nearly 50% of their awake time restrained by caregivers (i.e., carrying and holding).
-However, studies using ecological momentary assessment (EMA) found a smaller time amount of restraint.
-For example, @Malachowski2023z reported that 4- to 6-month-olds spent 41.13% of *total observation time* being held, and 45.43% in placement with physical support.
-The time proportion will presumably be less (and thus less than the above retrospective survey studies) when excluding infant sleep time.
-A more direct estimate comes from @Franchak2024v.
-They found that 10- to 13-month-olds spent 26.90% of *awake* time restrained in furniture or devices, and 18.79% restrained by caregivers, which is also lower than previous survey studies.
-In short, it appears that EMA gives lower estimates of infant daily restrained time compared to parental retrospective reports.
+```
 
-Does infant restrained time change with age?
-Only a handful of studeis have answered this question, but their findings are mixed as well.
-@Airaksinen2024z, using inertial measurement sensors (IMU) and machine learning algorithms, found that infants from Finland experienced *increased* holding and carrying time from 4 to 18 months of age.
-On the contrary, @Carson2022q, using parental retrospective report and longitudinal design, found that infants from Canada experienced *decreased* overall restrained time from 2 to 6 months of age.
-@Franchak2024v, using EMA and longitudinal design, found that infants from the US experienced *decreased* time held by caregivers from 10 to 13 months, yet the time restrained by furniture or devices did not change with age.
-The discrepancies in findings may be attributed to differences in the targeted behaviors (i.e., whether to include device restraint or not, and which devices to include), study design (i.e., cross-sectional vs.
-longitudinal), cultural contexts, as well as research methods.
-To address the current discrepancies in findings, the first step is to develop an accurate and reliable system that examines infant general restraint experiences, which can be further applied to different age ranges and culture contexts.
-Among the available methods, we propose IMU is the most objective, unobtrusive, and efficient way to caputure whole-day experiences in natural settings.
-But previous studies that have used IMU on infants only looked at holding behavior, excluding other extensive restrained practices such as device use [@Airaksinen2024z; @Franchak2024i].
-Thus, a new customized machine learning algorithm is needed to identify a broader range of restraint experiences.
+```{r tbl-indexbetweenmodels}
+# Compare between models
 
-## Limitations of previous research methods
+table_mean <- matrix_comb %>%
+  group_by(positive_class, model) %>%
+  summarise(
+    mean_acc = mean(balanced_accuracy, na.rm = TRUE),
+    mean_sen = mean(sensitivity, na.rm = TRUE),
+    mean_ppv = mean(pos_pred_value, na.rm = TRUE),
+    mean_f1 = mean(f1, na.rm = TRUE),
+    mean_prev_df = mean(prevalence_df, na.rm = TRUE),
+    .groups="drop"
+  ) %>%
+  pivot_longer(mean_acc:mean_prev_df, names_to = "Indices", values_to = "mean") %>%
+  pivot_wider(names_from = model, values_from = mean) %>% 
+  mutate(Indices = recode(Indices,
+                        "mean_acc" = "Accuracy",
+                        "mean_sen" = "Sensitivity",
+                        "mean_ppv" = "Positive Predictive Value",
+                        "mean_f1" = "F1",
+                        "mean_prev_df" = "Prevalence Difference")) %>%
+  relocate("Indices", .before = "positive_class") %>% 
+  mutate(Indices = factor(Indices, levels=c("Accuracy", 
+                              "Sensitivity", 
+                              "Positive Predictive Value", 
+                              "F1", 
+                              "Prevalence Difference"))) %>% 
+  arrange(Indices) %>% 
+  rename("Category" = "positive_class",
+         "4s model" = "4_no_pos",
+         "16s model" = "16_no_pos",
+         "30s model" = "30_no_pos")
 
-As illustrated above, survey is the most commonly used method for quantifying infants daily restrained experiences, either in the form of retrospective report or ecological momentary phone survey across days.
-In most of the retrospective parental report studies, parents were asked to estimate their infant's average daily restrained time over the past few weeks or month [@Abbott2001z; @Bartlett2003d; @Callahan1997s; @Carson2022q; @Hesketh2015q; @Karasik2018o; @Siddicky2020m].
-Despite its widespread use, this method may not be precise because it highly demands parents’ memory—It is difficult to recall and sum every bout of restraint practice and then report an overall time estimation [@Bradburn1987g].
-There are exceptions where researchers asked caregivers to complete structured daily diaries reflecting on the past day’s restraint practices in short intervals (e.g., every five minutes) [@Karasik2018o; @Majnemer2005g].
-However, this method is still sensitive to memory loss and thus not reliable in providing objective data.
+aov_mdl <- function(index, ctgr){
+  temp_ds <- matrix_comb %>% filter(positive_class==ctgr)
+  fmla <- reformulate("model", response = index)  # response ~ model
+  anova <- tidy(aov(fmla, data = temp_ds))
+}
+aov_r <- map(indx, ~aov_mdl(index=.x, ctgr="r")) %>% 
+  map2_dfr(indx, ~mutate(as.data.frame(.x), Indices = .y)) %>% 
+  mutate(Category = "r")
+aov_u <- map(indx, ~aov_mdl(index=.x, ctgr="u")) %>% 
+  map2_dfr(indx, ~mutate(as.data.frame(.x), Indices = .y)) %>% 
+  mutate(Category = "u")
+table_aov <- rbind(aov_r, aov_u) %>% 
+  mutate(df_resi = lead(df, n = 1)) %>% 
+  filter(term=="model") %>% 
+  mutate(df = str_glue("({df}, {df_resi})")) %>% 
+  select(-term, -df_resi) %>% 
+  rename("SS" = "sumsq",
+         "MS" = "meansq",
+         "F" = "statistic",
+         "p" = "p.value") %>% 
+  relocate("Indices", .before = "df") %>% 
+  relocate("Category", .before = "df") %>% 
+  relocate("df", .before = "p") %>%
+  mutate(Indices = recode(Indices,
+                        "balanced_accuracy" = "Accuracy",
+                        "sensitivity" = "Sensitivity",
+                        "pos_pred_value" = "Positive Predictive Value",
+                        "f1" = "F1",
+                        "prevalence_df" = "Prevalence Difference")) %>%
+  mutate(Indices = factor(Indices, levels=c("Accuracy", 
+                              "Sensitivity", 
+                              "Positive Predictive Value", 
+                              "F1", 
+                              "Prevalence Difference"))) %>% 
+  arrange(Indices)
 
-The recently developed method EMA lessens caregivers’ memory burden by making momentary phone-survey prompts to caregivers multiple times (e.g., 5, 10, or 12 times) randomly throughout a day [@Franchak2019v; @Franchak2024v; @Malachowski2023z].
-Researchers then estimate the proportion of time infants spend restrained by calculating the frequency of restraint across all prompts.
-Presumably, EMA can improve accuracy compared to retrospective report.
-The estimate discrepancy between retrospective report and EMA that illustrated above indeed supports the potential difference between the two methods.
-However, the two methods still share crucial limitations.
-First, both retrospective reports and EMA rely on parental input, which assumes that caregivers are consistently or equally present to observe the infant throughout the day.
-If the infant’s certain activities are systematically unaware of by the caregiver (e.g., the caregiver is not at home in the afternoons), the report might not be fully representative of infant daily experiences.
-Thus, a method not relied on parental report is needed.
-Second, neither method provides time-sequence information with high granularity.
-They cannot answer finer-grained questions such as: When during the day is restraint most common?
-Are restraint bouts prolonged or brief and frequent?
-These questions call for a method with higher time-resolution.
+table_btw_mdl <- cbind(table_mean, table_aov %>% select(-Indices, -Category)) %>% 
+  slice(-2) %>% 
+  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
+         p = str_remove(p, "^0")) %>% 
+  mutate(across(where(is.numeric), ~ round(.x, 3)))
+  # Still need to combine some rows, and add a heading as "accuracy" for three models 
+```
 
-Video recording and human annotation, as another method, offers more accurate and temporally detailed data than survey-based methods.
-However, this approach is labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies.
-As a result, most studies using video rely on short samples, normally around one hour (e.g., @Wang2025x).
-Accordingly, this method is sensitive to sampling bias because it is questionable for one-hour data to represent infant whole-day activities.
-Moreover, video recording can be obtrusive and may affect caregivers’ and infants’ behaviors, potentially introducing reactivity effects (XXX).
+```{r tbl-indexbetweenctgr}
+# Compare between categories while controlling for models
+lm_ctgr <- function(model_select){
+  temp_ds <- matrix_comb %>% filter(model==model_select)
+  list(
+    sen = summary(lm(sensitivity ~ positive_class, data=temp_ds))$coefficients,
+    ppv = summary(lm(pos_pred_value ~ positive_class, data=temp_ds))$coefficients,
+    f1  = summary(lm(f1 ~ positive_class, data=temp_ds))$coefficients,
+    prev_df = summary(lm(prevalence_df ~ positive_class, data=temp_ds))$coefficients
+  )
+}
 
-## Promises of IMU system
+tbl_ctgr_30 <- lm_ctgr("30_no_pos")%>% 
+  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
+  relocate("Indices", .before = "Estimate")%>% 
+  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
+  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
+         p = str_remove(p, "^0")) %>% 
+  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
+  add_column(Term = rep(c("Intercept","Prediction of Unrestrained"), 4), .before = "Estimate") %>% 
+  select(-SE)
 
-Measuring with IMU system has the potential to overcome the limitations listed above.
-IMU, light-weight sensors that can be tucked in specially designed garments (in our study, infant leggings), can record participants’ real-time movement of different body parts (e.g., hips and ankles).
-Easy synchronization between different sensors and local data storage eliminates the need for experimenters to collect data in person, making the method unobtrusive.
-A typical sample rate of 62.5 Hz allows for millisecond-level recording of motion data.
-A battery life of XX hours allows for whole-day-long observation, overcoming sampling bias to the greatest extent.
+tbl_ctgr_16 <- lm_ctgr("16_no_pos") %>% 
+  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
+  relocate("Indices", .before = "Estimate")%>% 
+  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
+  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
+         p = str_remove(p, "^0")) %>% 
+  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
+  add_column(Term = rep(c("Intercept","Prediction of Unrestrained"), 4), .before = "Estimate") %>% 
+  select(-SE)
 
-Each sensor collects raw data as accelerometer signals (in x, y, and z orientations) and gyroscope signals (in roll, pitch, and yaw orientations).
-By aggregating signals from a given time window, we can calculate customized motion features (statistical summaries like mean, skew, etc.) at every moment.
-These features can then be used as input for machine learning algorithms to automatically classify infant behaviors (e.g., restrained and unrestrained).
-Human annotation from video recordings serves as ground truth labels for model training.
-This pipeline has been validated in categorizing adults’ activities (XXX; I need to do more research on this and word it differently), children’s activities (XXX), and infants’ activities (Franchak et al., 2024 for infant body position).
-IMU have also been used to detect caregiver holding \[\@Airaksinen2024z\], but no prior studies have applied IMU to classify broader forms of restraint, including both holding and device restraining.
-Thus, we propose a new IMU system which collects objective, reliable, unobtrusive, continuous, and long-form infant movement data as well as automatically identifies infant restrained status.
+tbl_ctgr_4 <- lm_ctgr("4_no_pos") %>% 
+  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
+  relocate("Indices", .before = "Estimate")%>% 
+  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
+  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
+         p = str_remove(p, "^0")) %>% 
+  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
+  add_column(Term = rep(c("Intercept","Prediction of Unrestrained"), 4), .before = "Estimate") %>% 
+  select(-SE)
 
-## Current Study
+tbl_ctgr <- cbind(tbl_ctgr_4, tbl_ctgr_16 %>% select(-Indices,-Term)) %>% 
+  cbind(tbl_ctgr_30 %>% select(-Indices, -Term))
 
-Given the limitations in current methods and the potential of IMU technology, the overall goal of this study is to evaluate the validity of using IMUs to precisely quantify infant full-day restrained experiences at home.
-Despite previous validation of IMU system on identifying infant position and caregiver holding, identifying all restraint practices on infants is still challenging.
-Consider an infant sitting on a couch independently versus an infant sitting on a highchair restrained, or an infant lying supine unrestrained versus an infant being tucked in cradle restrained.
-The infants are in same body position and are very likely to show similar movement patterns, but the context is different, differentiating the restrained status.
-Furthermore, device restraint might be more difficult to detect than caregiver restraint because carrying and holding tend to generate dramatic movement (e.g., the infant moving in an adult speed).
-Thus, in the current study, we develop and validate the use of new machine-learning models that identify both caregiver restraint and device restraint, classifying infant moment-to-moment status as being restrained or unrestrained.
+# ctgr_4 <- ctgr_4 %>% 
+#   flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
+#   merge_v(j = "Indices") %>% fix_border_issues() %>% 
+#   hline(i = seq(2,8,2))%>% 
+#   padding(padding.top = 2, padding.bottom = 2) %>% 
+#   set_table_properties(layout = "autofit", width = .6) %>% 
+#   align(align = "center", part = "all")
 
-The models are tested by a total of 143 sessions from 66 infants who were aged from 7 to 14 months, covering a wide range from pre-sitting infants to walking infants.
-We sent caregivers special leggings with four IMUs embedded and a camera to record synchronized infant activities.
-Across the sessions, IMUs recorded an average of XX hours (*SE* = XX) of movement data, and the camera recorded the first 1.5 hours into the session.
-From the videos, human coders identify infant restrained status as ground truth for model training and testing.
-The **first aim** of the current study is to train the supervised machine-learning models of infant restrained status and evaluate the models’ accuracy using leave-one-out cross validation method (XXX) for the first 1.5 hours.
+```
 
-If the models perform well, we will use the same rationale to train a new group model by feeding in all sessions’ human annotation.
-This model will then be applied to the remaining, unannotated portions of the IMU recordings to estimate infants’ full-day restraint time.
-Unfortunately, neither did previous studies establish a clear convergent estimate of infant overall restrained time proportion, nor did they find consistent age-related change, probably due to method differences.
-Considering EMA is likely to be more accurate than parental retrospective reports, we will compare our current estimate with @Franchak2024v, who found that 10- to 13-month-old infants spent 45.69% awake time being restrained.
-We do not compare our estimate data with @Malachowski2023z, another EMA study, because their reported estimates are among total observation time rather than awake time.
-We will also compare our estimated age-related change with @Franchak2024v who found that infant experienced increased unrestrained time from 10 to 13 months.
-We do not compare our age-related change with @Airaksinen2024z, an IMU-based study, because they only looked at caregiver restraint and excluded device restraint.
-In short, our **second aim** is to validate our new system by comparing empirical results with a previous study.
-However, due to methodological differences again (IMU versus EMA), it remains an open question whether these estimates will converge.
-\[note, when I was writing here, I feel we should probably at least distinguish caregiver restraint and device restraint\]
+```{r fig-indexcomparison}
+# Sensitivity
+p_sen <- ggplot(data=matrix_comb, aes(x=model, y=sensitivity, color=positive_class, shape=model))+
+  geom_point(position = position_jitterdodge(
+              jitter.width = 0.4,
+              jitter.height = 0,
+              dodge.width = 0.6),
+             size = 5)+
+  scale_color_manual(values=class_colors,labels=c("restrained","unrestrained")) +
+  scale_shape_manual(values=model_shapes) +
+  stat_summary(fun = mean,
+               geom = "crossbar",
+               aes(group = positive_class),  
+               position = position_dodge(width = 0.6), 
+               width = 0.4,
+               color = "black") +
+  labs(x="",
+       y="Sensitivity",
+       color="Category")+
+  scale_x_discrete(labels = c("4s","16s","30s"))+
+  scale_y_continuous(
+    limits = c(0, 1),  
+    breaks = c(0, 0.25, 0.5, 0.75, 1) 
+  ) +
+  guides(color = "none",shape="none")+
+  theme_update()
+p_sen
 
-# Method
+# Positive Predictive Value
+p_ppv <- ggplot(data=matrix_comb, aes(x=model, y=pos_pred_value, color=positive_class, shape=model))+
+  geom_point(position = position_jitterdodge(
+    jitter.width = 0.4,
+    jitter.height = 0,
+    dodge.width = 0.6),
+    size = 5)+
+  scale_color_manual(values=class_colors,labels=c("restrained","unrestrained")) +
+  scale_shape_manual(values=model_shapes) +
+  stat_summary(fun = mean,
+               geom = "crossbar",
+               aes(group = positive_class),  
+               position = position_dodge(width = 0.6), 
+               width = 0.4,
+               color = "black") +
+  labs(x="",
+       y="Positive Predictive Value",
+       color="Category")+
+  scale_x_discrete(labels = c("4s","16s","30s"))+
+  scale_y_continuous(
+    limits = c(0, 1),  
+    breaks = c(0, 0.25, 0.5, 0.75, 1) 
+  ) +
+  guides(shape="none")+
+  theme_update()
+p_ppv
+# F1 score
+p_f1 <- ggplot(data=matrix_comb, aes(x=model, y=f1, shape=model, color=positive_class))+
+  geom_point(position = position_jitterdodge(
+    jitter.width = 0.4,
+    jitter.height = 0,
+    dodge.width = 0.6),
+    size = 5)+
+  scale_color_manual(values=class_colors,labels=c("restrained","unrestrained")) +
+  scale_shape_manual(values=model_shapes) +
+  stat_summary(fun = mean,
+               geom = "crossbar",
+               aes(group = positive_class),  
+               position = position_dodge(width = 0.6), 
+               width = 0.4,
+               color = "black") +
+  labs(x="Models' Window Size",
+       y="F1",
+       color="Category")+
+  scale_x_discrete(labels = c("4s","16s","30s"))+
+  scale_y_continuous(
+    limits = c(0, 1),  
+    breaks = c(0, 0.25, 0.5, 0.75, 1) 
+  ) +
+  guides(color = "none",shape="none")+
+  theme_update()
+p_f1
+# Prevalence Difference
+p_prev_df <- ggplot(data=matrix_comb, aes(x=model, y=prevalence_df, shape=model, color=positive_class))+
+  geom_point(position = position_jitterdodge(
+    jitter.width = 0.4,
+    jitter.height = 0,
+    dodge.width = 0.6),
+    size = 5)+
+  geom_hline(yintercept = 0, linetype = 'dashed')+
+  scale_color_manual(values=class_colors,labels=c("restrained","unrestrained")) +
+  scale_shape_manual(values=model_shapes) +
+  stat_summary(fun = mean,
+               geom = "crossbar",
+               aes(group = positive_class),  
+               position = position_dodge(width = 0.6), 
+               width = 0.4,
+               color = "black") +
+  labs(x="Models' Window Size",
+       y="Prevalence Difference",
+       color="Category")+
+  scale_x_discrete(labels = c("4s","16s","30s"))+
+  scale_y_continuous(
+    limits = c(-.8, .8),
+    breaks = c(-.8, -.4, 0, .4, .8)
+  ) +
+  guides(shape = "none")+
+  theme_update()
+p_prev_df
+
+# figure arrangement
+ggpubr::ggarrange(p_sen, p_ppv, p_f1, p_prev_df, ncol = 2, nrow = 2, 
+                  labels = c("A", "B", "C", "D"), 
+                  label.x = 0, 
+                  label.y = 1,
+                  font.label = list(size = 16, face = "bold"),
+                  widths=c(6,7), heights=c(4,4))
+ggsave("figure/indices.png", width=13, height=8)
+ggsave("figure/indices.eps", width=13, height=8)
+```
+
+We evaluated the following indices: (1) accuracy, (2) sensitivity, (3) positive predictive value, (4) F1, and (5) prevalence difference for all three models.
+Prevalence difference is a subtraction of a model prediction’s prevalence of certain category from the human annotation’s prevalence, a positive value indicating the models’ overall tendency to overestimate a category.
+
+Models performed equally well in all indices. (Table) Indices mean and ANOVA results. XXX
+
+Despite the overall good performance, the models presented predicting bias towards unrestrained category.
+(Table + Figure).
+The sensitivity for predicting unrestrained periods was significantly higher for all models.
+The positive predictive value for predicting unrestrained periods was generally lower for all models, but 16s and 30s models did not significantly over-predict the unrestrained periods as 4s model did.
+The F1 value for both categories was comparable regardless of the models’ window size.
+The prevalence difference value for unrestrained category was significantly higher for all models, but 30s model did not significantly overestimate unrestrained periods or underestimate restrained periods.
+
+An ideal model should present high accuracy, balanced performance across categories, and minimal offset in overall prevalence compared to ground truth.
+Based on these criteria, we selected the 30-s window model and applied it to process the full-day data.
+
+## Model Estimate Compared to Previous Studies
+```{r}
+load("whole-model-prediction.RData")
+rm(ds, ds_sum_excl, ds_sum_excl0, id_demo, id_info)
+# exclude sessions where total recording time is <3 hours
+ds_excl <- ds_sum %>% filter(total_time < 3) #11 sessions 
+ds_sum_excl <- ds_sum %>% filter(total_time >=3) #132 sessions left
+ds_sum_excl <- ds_sum_excl %>% 
+  filter(unique_id != "148/1",
+         unique_id != "178/3",
+         unique_id != "185/3")
+```
+
+```{r}
+ds_sum_excl %>% 
+  group_by(age_mo) %>% 
+  summarize(mean = mean(unrestrained_prop),
+            sd = sd(unrestrained_prop),
+            .groups="drop")
+ds_sum_excl %>% 
+  group_by(age_group) %>% 
+  summarize(mean = mean(unrestrained_prop),
+            sd = sd(unrestrained_prop),
+            .groups="drop")
+
+summary(lm(unrestrained_prop ~ age_group, data = ds_sum_excl)) # t test
+
+ggplot(ds_sum_excl) +
+  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
+  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
+  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
+ggsave("figure/age_related_change.png",width=6, height=4)
+ggsave("figure/age_related_change.eps",width=6, height=4)
+
+```
+
+Next, we compared the 30s model’s estimates of full day restrained time duration with previous studies (XX).
+We excluded eleven sessions where the total recording time was less than 3 hours.
+We excluded another two sessions where a big chunk of duration should have been excluded due to sensors removal, but we did not have reliable time stamp of when to start excluding.
+We also excluded another one session where the family put the leggings on a wrong infant later into the session.
+
+Using the rest 129 sessions, the model reported an average of XX% of awake time (*SD* = XX) being restrained for younger infants (4-7 months) and XX% (*SD* = XX) for older infants (11-14 months).
+Figure XX shows a clear decrease in the percent of restrained time from younger infants to older infants.
+These results are consistent with previous findings where XXX.
+
+```{r}
+# ARCHIVE
+### Prevalence correlation scatter plot
+p_prev_scatter_r <- ggplot(data=matrix_comb %>% 
+                             filter(positive_class=="r" & model=="30_no_pos"), 
+                           aes(x = prevalence, y = detection_prevalence))+
+  geom_point(size = 5, color = class_colors[1]) +
+  geom_abline(intercept = 0, slope = 1, color = "black") +
+  scale_x_continuous(limits = c(0, 1)) +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_update() +
+  labs(x = "Ground Truth Prevalence", y = "Detection Prevalence",
+       title = "Restrained Prediction by 30s Model")
+p_prev_scatter_r
+
+p_prev_scatter_u <- ggplot(data=matrix_comb %>% 
+                             filter(positive_class=="u" & model=="30_no_pos"), 
+                           aes(x = prevalence, y = detection_prevalence))+
+  geom_point(size = 5, color = class_colors[2]) +
+  geom_abline(intercept = 0, slope = 1, color = "black") +
+  scale_x_continuous(limits = c(0, 1)) +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_update() +
+  labs(x = "Ground Truth Prevalence", y = "Detection Prevalence",
+       title = "Unrestrained Pred by 30s Model")
+p_prev_scatter_u
+
+ggpubr::ggarrange(p_prev_scatter_r,p_prev_scatter_u, ncol = 2, nrow = 1, 
+                  labels = c("A", "B"), 
+                  label.x = 0, 
+                  label.y = 1,
+                  font.label = list(size = 16, face = "bold"),
+                  widths=c(5,5), heights=c(5))
+ggsave("figure/prevalence_scatter.png", width=10, height=5)
+ggsave("figure/prevalence_scatter.eps", width=10, height=5)
+```

### 2025-10-01 manuscript_restrained_classification validation.qmd
updated table 1

@@
 ---
-title: "Validation of quantifying infant all-day restrained experiences at home using inertial sensors"
+title: "Quantifying Infants' Everyday Restrained Experiences in the Home Using Wearable Inertial Sensors"
 blank-lines-above-title: 2
@@
 author:
-  - name: XX
+  - name: Hanzhi Wang
+    orcid: 0009-0000-0520-5935
+    email: hanzhi.wang@email.ucr.edu
+    affiliations:
+      - id: id1
+        name: University of California, Riverside
+        department: Department of Psychology
+  - name: Hailey N. Rousey
     orcid: XX
@@
         department: Department of Psychology
-  - name: XX
+  - name: John M. Franchak
     corresponding: true
-    orcid: XX
-    email: XX
+    orcid: 0000-0002-0751-2864
+    email: franchak@ucr.edu
     affiliations:
@@
     # Example: This study was registered at ClinicalTrials.gov (Identifier NTC998877).
-    study-registration: ~
+    study-registration: null
     # Acknowledge and cite data/materials to be shared.
-    data-sharing: ~
+    data-sharing: null
     # Example: This article is based on data published in Pulaski (2017).
@@
     # Example: Sally Jones earns royalties from the sale of Test X.
-    conflict-of-interest: ~
+    conflict-of-interest: null
     # Example: This study was supported by Grant A123 from the National Science Foundation.
-    financial-support: ~
+    financial-support: null
     # Example: The authors are grateful for the technical assistance of Dr. X during the initial design and setup of our lab equipment.
-    gratitude: ~
+    gratitude: null
     # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
-    authorship-agreements: ~
+    authorship-agreements: null
 abstract: "XXX"
-keywords: [infant, restraint, everyday experiences, inertial sensors, machine learning]
+keywords: [infant, restraint, everyday experiences, wearable sensors, machine learning]
 bibliography: [references.bib]
@@
     mask: false
-    journal: ~
-    volume: ~
-    course: ~
-    professor: ~
-    duedate: ~
+    journal: null
+    volume: null
+    course: null
+    professor: null
+    duedate: null
     # Hides correspondence text
     nocorrespondence: false
+editor_options: 
+  chunk_output_type: console
 ---
@@
 
-theme_update(text = element_text(size = 14),
-             axis.text.x = element_text(size = 14, color = "black"),
-             axis.title.x = element_text(size = 16),
-             axis.text.y = element_text(size = 14,  color = "black"),
-             axis.title.y = element_text(size = 16),
+theme_update(text = element_text(size = 11),
+             axis.text.x = element_text(size = 11, color = "black"),
+             axis.title.x = element_text(size = 12),
+             axis.text.y = element_text(size = 11,  color = "black"),
+             axis.title.y = element_text(size = 12),
              panel.background = element_blank(),panel.border = element_blank(),
@@
 
-indx <- c("balanced_accuracy","sensitivity","pos_pred_value","f1","prevalence_df")
+indx <- c("balanced_accuracy","kappa", "sensitivity","pos_pred_value","f1")
 indices <- c("Sensitivity", "Positive Predictive Value", "F1", "Prevalence Difference")
@@
 
+```{r}
+# demographics
+load("whole-model-prediction.RData")
+rm(ds, ds_sum_excl, ds_sum_excl0, id_demo, id_info)
+ds_sum <- ds_sum %>% 
+  filter(unique_id != "185/3")
+ds_human <- ds_human %>% 
+  filter(unique_id != "185/3")
+demo <- read.csv("lena-imu-compiled.csv")
+demo[demo$id=="180", "infant_sex"] <- "male"
+
+temp <- ds_sum
+temp <- temp %>% 
+  separate(unique_id, into=c("id","sess"),sep="/",convert = TRUE) %>% 
+  distinct(id, .keep_all = TRUE)
+
+temp <- left_join(temp, demo %>% select("id","infant_sex","infant_race","infant_ethnicity") %>% distinct(id, .keep_all = TRUE), by="id", unmatched="drop")
+
+age<-temp %>% 
+  group_by(age_group) %>% 
+  summarize(n=n(),
+            .groups="drop")
+
+sex<-temp %>% 
+  group_by(infant_sex) %>% 
+  summarize(n=n(),
+            .groups="drop")
+race<-temp %>% 
+  group_by(infant_race) %>% 
+  summarize(n=n(),
+            .groups="drop")
+
+```
+
+```{r}
+# Human coding reliability
+colnames(ds_human) <- c("unique_id","agree","kappa")
+library(psych)
+stats_human <- ds_human %>% 
+  summarize(mean_agree = mean(agree),
+            min_agree = min(agree),
+            max_agree = max(agree),
+            mean_kappa = mean(kappa),
+            min_kappa = min(kappa),
+            max_kappa = max(kappa))
+```
+
+Physical restraint is a common feature in infants’ everyday life.
+Infants are routinely held or carried by caregivers, or restrained in devices like highchairs, walkers, and strollers for the sake of comforting, safety, transportation, etc [@Birken2015s].
+Restraint influences infants’ motor development by constraining their body movement [e.g., @Pin2007z].
+Restraint also frames infants’ learning experiences for perceptual and cognitive development.
+For example, the frequency of object holding experiences differs between restrained and unrestrained times [@Franchak2024l], the amount and quality of language infants receive decrease as infants spend more time restrained in devices [@Malachowski2023a], yet the frequency of joint attention and
+
+Holding and carrying restraint practices may facilitate infants' social cognitive development [e.g., @Berecz2020y].
+Restraint, as an externally regulated aspect of infants’ daily life, decides the scope of input and experiences infants have access to.
+Therefore, characterizing infants restrained experiences helps profile infants’ daily learning contexts and situate their motor, perceptual, and cognitive development within everyday experiences.
+
+Previous studies made several attempts to document infant restrained time at home.
+Some used parental surveys, including questionnaires, diaries, or ecological momentary assessment [e.g., @Abbott2001u; @Karasik2018t; @Franchak2019o].
+Some used video recording and human annotation [e.g., @Wang2025v].
+However, these methods cannot provide a continuous and prolonged account of infants’ experiences.
+In the current study, we propose that wearable sensors can be a solution for recording continuous, full-day movement data.
+Moreover, previous studies had different emphases on restraint practices.
+Some only focused on caregiver restraint like holding and carrying [e.g., @Airaksinen2024z], some only focused on seating devices use [e.g., @Callahan1997s], some focused broadly on containment equipment [e.g., @Bartlett2003w; @Carson2022n; @Hesketh2015i; @Karasik2022a], and others covered a wider range of restraint like furniture placement, seating devices and holding [e.g., @Siddicky2020f; @Franchak2024l; @Malachowski2023a].
+In the current study, we combine different aspects of restraint that have been separately examined into one investigation.
+
+## Previous Measurements of Restrained Experiences
+
+Despite different emphases on infant restrained experiences, past research shows that both caregiver restraint and device restraint are common in infant daily life, and the time infants spend restrained changes over development.
+In terms of caregiver restraint, infants aged from 2 to 6 months spent around 40% of their awake time being held or carried [@Malachowski2023a; @Siddicky2020f].
+Older infants aged from 10 to 13 months spent around 19% of their awake time being held or carries [@Franchak2024l].
+In terms of device restraint, infants aged from 4 to 6 months spent around 20% of awake time in seating devices, and 35% in placement with physical support [@Callahan1997s; @Malachowski2023a].
+Older infants aged from 10 to 13 months spent around 27% of awake time in restraining furniture or devices [@Franchak2024l].
+
+Based on the statistics above, it appears that caregiver restraint and device restraint decrease with age.
+However, other work that compared across ages within a single study presented contradictory results.
+@Airaksinen2024z, using wearable sensors and machine learning modeling, found that infants from Finland experienced increased holding and carrying time from 4 to 18 months of age.
+On the contrary, @Carson2022n, using parental retrospective report, found that infants from Canada experienced decreased overall restrained time from 2 to 6 months of age.
+@Franchak2019o, using ecological momentary assessment (EMA), also found that infants from the US experienced decreased time held by caregivers from 3 to 12 months.
+The author and their colleagues further expanded the findings to 10- to 13-month-olds, showing that infants experienced decreased held time and increased overall unrestrained time with age [@Franchak2024l].
+The discrepancies may be attributed to study design (i.e., longitudinal versus cross-sectional), cultural differences, as well as research methods (i.e., wearable sensors versus parental reports).
+The current study, which uses wearable sensors on U.S.
+infants aged from 4 to 14 months, might help clarify the inconsistencies found in prior research.
+
+## Restrained Practices Shape Infant Learning Opportunities in Real Time
+
+Beyond measuring total duration of infants’ restrained time, prior research also examined how real-time restrained experiences may constrain or support infant learning.
+On the one hand, restraint practices limit infants’ opportunities for learning.
+First, restraint practices prevent infants from moving their body parts (e.g., spine, legs, hip) and thus may cause temporary delay in motor development [@Abbott2001u; @Bartlett2003w; @Carson2022n; @Karasik2023d; @Pin2007z; @Siddicky2020f].
+Second, infants with relatively more restrained time may produce fewer object interactions, with longest caregiver restrained periods related to the least object interaction [@Franchak2024l].
+Third, infants with relatively more device-restrained time may hear less amount of and less consistent adult language [@Malachowski2023a].
+
+On the other hand, restrained periods can enrich certain kinds of learning, especially for younger infants, through stabilizing their body posture.
+Consider young infants who cannot sit up independently yet and thus cannot free up their hands to explore objects [@Soska2010t].
+Restraint devices allow these pre-sitting infants to experience the advantages of a sitting posture by steadying their bodies and enabling eye-hand coordination.
+Moreover, infants in sitting or upright positions have better opportunities to look at faces compared to infants in prone position [@Franchak2018r].
+Restraint devices or carriers that rescue infants from prone position can elevate their viewpoint and afford richer visual input.
+These examples demonstrate that restraint practices might provide infants with learning experiences that they would not otherwise obtain due to their body constraints, especially for younger infants.
+These examples also underscore that real-time observation of restraint is essential for understanding how context shapes infants’ multimodal learning opportunities.
+
+## Limitations of Previous Research Methods
+
+Researchers have used survey methods including traditional retrospective report and ecological momentary assessment (EMA) to quantify infants’ restrained time, but both methods have similar limitations.
+Retrospective parental report, which was most commonly used, asks parents to estimate their infant’s average daily restrained time over the past few weeks or month [@Abbott2001u; @Bartlett2003w; @Callahan1997s; @Carson2022n; @Hesketh2015i; @Karasik2018t; @Siddicky2020f].
+Despite its widespread use, this method may not be precise enough because it highly relies on parents’ memory—it is difficult to recall and sum every bout of restraint practices and then report an overall time estimation [@Bradburn1987g].
+There are exceptions where caregivers were asked to complete structured daily diaries reflecting on the previous day’s restraint practices in short intervals (e.g., every five minutes) [@Karasik2018t; @Majnemer2005j].
+However, this method might still be sensitive to memory loss and thus not reliable in providing accurate data.
+
+The recently developed method EMA reduces caregivers’ memory burden by making momentary phone-survey prompts to caregivers multiple times (e.g., 5, 10, or 12 times) randomly throughout a day [@Franchak2019o; @Franchak2024l; @Malachowski2023a].
+However, same limitations still remain.
+First, both methods assume that the respondent is consistently present to observe the infant throughout the day.
+The report might not be fully representative of full-day activities if the infant happens to be restrained more frequently when the respondent is absent (e.g., the caregiver restrains the infant at home while they goes out on errands).
+Second, neither method provides continuous time-series data.
+They cannot answer granular questions such as: When throughout the day does restraint happen most frequently?
+Are restrained bouts usually prolonged or brief?
+Thus, a method not relying on parental report and providing time-sequence information is needed.
+
+Video recording and human annotation, as another method, overcomes the above drawbacks of survey-based methods.
+However, one of the limitations of this method is being obtrusive.
+The presence of a researcher in the room may affect the caregiver’s and infant’s behaviors.
+Another limitation is that this approach is labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies.
+Indeed, these studies usually do not have long duration of recording (e.g., one hour in @Wang2025v).
+Within the short duration of recording, researchers would tend to record the period when the infants are awake and active.
+In contrast, the periods when the infants are restrained, like seated in front of TV or in highchairs, would likely be skipped.
+Thus, this method is sensitive to sampling bias and might not be representative of infants’ whole-day experiences.
+
+## Promises and Challenges of Wearable Sensors
+
+A system that uses wearable sensors (i.e., inertial movement units) to record real-time movement data and machine learning models to classify movement categories can overcome all the limitations listed above.
+First, by collecting infants’ movement data through sensors, the system can objectively capture infants’ position, not relying on parental report.
+Second, the sensors can provide moment-to-moment observation thatd is continuous and rich in time-sequence information.
+Third, the sensors are embedded in baby garments, making the system less obtrusive.
+Fourth, the system can record and process full-day or multiple-day data at a time due to the sensors’ sufficient battery life and automative classification of machine learning models.
+
+The past decade has witnessed an uptake of using wearable sensors to record and classify movement categories, but challenges remain.
+The application of wearable sensors progressed from adults [@Arif2015u; @Preece2009n], to children [@Nam2013y; @Ren2016-px; @Stewart2018h], and then to infants [@Airaksinen2020j; @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q].
+For infants, there have been established models that classify infants’ body postures [@Airaksinen2020j; @Franchak2021k; @Franchak2024e], and models that detect holding and carrying periods [@Airaksinen2024z; @Yao2019q].
+Prior works concluded valuable experience, such as using multiple sensors, and including gyroscope data together with accelerator data [@Airaksinen2025l].
+
+However, one of the challenges researchers continue to face is the reduced model performance in naturalistic home settings [@Airaksinen2024z; @Franchak2024e; @Yao2019q], compared to more controlled laboratory environments [@Airaksinen2020j; @Franchak2021k].
+Indeed, infants may exhibit more complex and fluid body positions at home versus in lab, which requires the models to be able to generalize across scenarios.
+Moreover, since no prior work has used wearable sensors to detect infants’ overall restrained periods beyond caregiver restraint, the current study faces a unique challenge to detect this new behavior category.
+
+Detecting restrained periods would be even more challenging than classifying body postures or detecting holding periods only.
+First, infants may show similar body movements whether restrained or not.
+For example, infants would kick legs when they are lying without restraint and also when they are strapped in a swing.
+The differences in body movement between restrained and unrestrained periods may be subtle to capture.
+Possible distinctions might be manifested in reduced range of motion or repetitive motion due to constraints, which requires high-precision sensors and sensitive models to detect.
+Second, infants may present completely different body movements across various restrained scenarios.
+Holding is relatively easy to detect because it produces distinguishable patterns such as infants getting lifted or put down at an unusual speed, and infants moving at an adult’s pace with minimal leg movement.
+However, other kinds of restraint have diverse movement patterns.
+For example, infants seated in highchairs would be in a sitting position with occasional limbs swinging, whereas infants positioned in walkers would be in an upright position with extended legs and slow walking.
+Accordingly, the wearable sensors system must identify the underlying kinematic signatures common to various restraint types while accounting for their contextual variability.
+
+A promising strategy to address this untested classification lies in the flexibility of modeling approaches.
+One strategy is adjusting the window size within which movement data are aggregated.
+With a fixed sampling rate, larger windows provide more data for the model to leverage to predict the current category.
+Prior studies have used window sizes ranging from 2.3 to 4 seconds for infant posture category—a multiclass classification [@Airaksinen2020j; @Franchak2024e], and from 1.15 to 10 seconds for infant holding and carrying detection—a binary task [@Airaksinen2024z].
+Considering that transitions between restraint states are typically slower than transitions between postures, we set window sizes longer than 4 seconds.
+Theoretically, shorter windows have higher temporal resolution and are sensitive to abrupt changes but may not be robust enough to make holistic judgement.
+In contrast, longer windows can smooth noises and are more informative but may not be sensitive enough or may introduce irrelevant data from heterogeneous activities.
+Determining the optimal window size is therefore a key step for the model to reliably detect infants’ restrained periods.
+
+## Current Study
+
+In short, we propose that wearable sensors offer a more objective, unobtrusive, information-rich, and efficient approach to capturing infants’ full-day experiences in natural settings compared to prior methods.
+However, unique challenges remain in our new effort to detect infants’ restrained periods.
+We used an existing data set with infants aged from 4 to 14 months.
+Infants wore a special garment with four sensors tucked in pockets—two on each side of thighs and two on each side of ankles.
+In each visit, the family received the equipment dropped off at their doorstep in the morning.
+The caregiver was instructed to put the garment on the infant to record their whole-day movement and set up a third-person camera to record the first 1.5 hours of the study as video ground truth for human annotation.
+The current study tested whether a supervised machine learning model trained on human annotated labels can accurately classify infants’ restrained status.
+We varied the window size as 4s, 16s, and 30s and tested the model performance.
+
+We tested the model in two ways.
+First, we compared how accurately model predictions matched human annotation for the 1.5 hours of the video-recorded session.
+We defined \>85% agreement as a good accuracy, based on prior studies reporting 85–96% accuracy in classifying infants’ movement at home.
+Second, we applied the model to full-day data to estimate the proportion of time each infant was restrained.
+We tested the convergent validity of the model against past research to determine whether restrained time decreased with age [@Carson2022n; @Franchak2019o; @Franchak2024l].
+
+# Method
+
+## Dataset
+
+The data set was published on Databrary (Franchak, 2023) and used in other empirical works (XX).
+The data set consists of two cohorts of infants: `r apa_num(age[1,2])` younger infants aged from 4 to 7 months of age, and `r apa_num(age[2,2])` older infants aged from 11 to 14 months of age.
+Among the total of `r apa_num(age[1,2]+age[2,2])` infants, 26 of them participated in the study for only once, while 39 of them participated from 2 to 4 times, resulting in a total of 142 sessions.
+Thirty-four of the infants are female and thirty-one are male.
+Caregivers reported the infants’ race as White (*n* = 23), Hispanic or Latinx (*n* = 8), African American (*n* = 2), Asian (*n* = 2), more than one (*n* = 15), or not fitting into the above races or chose not to answer (*n* = 15).
+Families were recruited from Southern California via social media and community recruitment events and received \$30 for compensation for each visit.
+Caregivers provided informed consent before each session of the study.
+The whole procedures were approved by the University of California, Riverside Institutional Review Board.
+
+## Apparatus
+
+As shown in figure XX, a special infant garment was customized to secure four lightweight wearable sensors (MC10 Biostamp) at four specific locations (i.e., right thigh, left thigh, right ankle, and left ankle).
+Each sensor collects raw data as accelerometer signals (in x, y, and z orientations) and gyroscope signals (in roll, pitch, and yaw orientations).
+The sampling rate of 62.5 Hz allows for millisecond-level granularity, and the sufficient battery life and built-in storage allows for whole-day-long recording.
+An action camera fixed on a mini tripod (GoPro HERO9 Black) was used to record a third-person video of the first 1.5 hours into the session as ground truth.
+The camera and sensors were synchronized at the beginning of each session.
+
+## Procedure
+
+On the day of a session, a researcher arrived at the family’s house in the morning with a bucket of all equipment.
+The researcher turned on the camera and synchronized it with the sensors by dropping the garment in front of the camera—the action would leave a distinctive timestamp on both the camera and the sensors.
+The researcher then called the caregiver to pick up the bucket and instructed the caregiver to set up the equipment.
+The caregiver was told to place the camera to where it could capture the main room and to move the camera if needed to ensure the infant was always in view.
+The caregiver was instructed to put the garment on the infant and position the infant in various postures (including supine, prone, sitting, crawling, standing, walking, being held, and being restrained in seating devices) for at least one minute each for model training purpose.
+Next, the caregiver played with the infant as usual in front of the camera for 10 minutes.
+Afterwards, the family went about their day with the garment on the infant.
+The caregiver logged the time when the infant napped or that the garment was removed for reasons such as diaper changes, baths, going out on errands, etc.
+The researcher picked up the equipment the next day.
+
+## Human Annotation of Restraint
+
+Trained human coders watched the 1.5-hour videos and annotated the periods when the infant was restrained or unrestrained using Datavyu software (https://datavyu.org/).
+Restrained and unrestrained is defined as a binary category, meaning that the infant could only be in either of the two statuses unless they were out of view.
+We define restrained periods as when infants cannot initiate or control their own movements, which includes being held or carried by caregivers, or being restrained in devices that constrain them from changing body position.
+Unrestrained periods, on the contrary, are when infants could freely change their body position or locomote.
+Each video file was coded by two coders: one primary coder annotated the whole 1.5-hour video, and an independent reliability coder annotated the first 30 minutes of the video.
+The inter-rater reliability was calculated as the proportion of video frames where the two coders coded the same label.
+The human annotation reached a high agreement with an overall accuracy of `r apa_num(stats_human[1,1]*100)`%, ranging from `r apa_num(stats_human[1,2]*100)`% to `r apa_num(stats_human[1,3]*100)`%, and an average kappa of `r apa_num(stats_human[1,4])`, ranging from `r apa_num(stats_human[1,5])` to `r apa_num(stats_human[1,6])`.
+
+## Model Classification of Restraint
+
+The machine-learning classification process followed procedures similar to [@Franchak2021k; @Franchak2024e].
+First, we aligned the time series of human annotations and sensor data by matching the synchronization time stamp in the video and sensors.
+This resulted in a data set with human-coded labels and motion data from four sensors—each capturing acceleration (x, y, z) and angular velocity (roll, pitch, yaw)—available at each sampling point.
+
+Building on previous methods, we focused on short temporal windows to capture the dynamic nature of movement.
+Accordingly, we aggregated raw IMU data into motion features within sliding windows of n seconds.
+For each window, we computed statistical features—mean, median, minimum, 25th percentile, 75th percentile, maximum, skewness, kurtosis, standard deviation, and sum—for all six signals across the four sensors, resulting in 240 features.
+We then added 196 additional features based on cross-sensor and cross-orientation metrics such as correlations and pairwise differences, resulting in 436 features per window.
+
+We started the windowing process from the syncing point and slide the window every *n*/2 seconds.
+Consequently, each time point spaced every *n*/2 seconds contains 436 motion features, aggregated from the *n*/2 seconds before and after that point.
+For model training and testing, we included only time windows where the infant remained in a single status, so that each window could be assigned one unambiguous label.
+The window size *n* varied from 4s, to 16s and 30s.
+
+The third step was to train and validate supervised machine learning models that classify the motion features into restrained status.
+We applied the random forest algorithm (Breiman, 2001) in modeling using the *DecisionTree* package (Sadeghi et al., 2022) in Julia.
+We used a “training” set of the windowed data set and validated the models by comparing their predictions on a separate “testing” set to the corresponding human annotations.
+Specifically, we used a leave-one-out cross-validation approach—each session was held out once and only as the testing set—to evaluate how well the model generalized to unseen data.
+After validation, an optimal window size was picked and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.
+
+## Data Sharing
+
+NOT SURE HOW WE ARE GOING TO SHARE THE DATA
+
 # Results
-We first compared the model’s performance to human annotation of the windowed data set of the first 1.5 hours across all 142 sessions.
-We then compared the model’s estimate to previous works regarding infants’ full day restrained time and developmental change.
+
+We first compared the model’s performance to the human annotation of the first 1.5 hours across all 142 sessions.
+We then compared the model’s estimate of infants' full day restrained time and its developmental change to previous works.
 
 ## Model Performance Compared to Human Annotation
+
 ```{r}
@@
 
-# Descriptives
-stats <- matrix_comb %>% 
-  group_by(model, positive_class) %>% 
-  summarize (mean_acc = mean(balanced_accuracy, na.rm=T),
-             sd_acc = sd(balanced_accuracy, na.rm=T),
-             mean_sen = mean(sensitivity, na.rm=T),
-             sd_sen = sd(sensitivity, na.rm=T),
-             mean_ppv = mean(pos_pred_value, na.rm=T),
-             sd_ppv = sd(pos_pred_value, na.rm=T),
-             mean_f1 = mean(f1, na.rm=T),
-             sd_f1 = sd(f1, na.rm=T),
-             mean_code_prev = mean(prevalence, na.rm=T),
-             mean_pred_prev = mean(detection_prevalence, na.rm=T),
-             .groups="drop")
-
 ```
@@
 ```{r tbl-indexbetweenmodels}
+#| apa-note: Age in months served as a continuous predictor and was centered within each age group; skill was a categorical predictor with non-sitters/non-walkers serving as the reference categories
+#| tbl-cap: Performance indices and ANOVA comparison across models
+#| ft.align: left
+
 # Compare between models
-
+# Descriptive
 table_mean <- matrix_comb %>%
@@
     mean_acc = mean(balanced_accuracy, na.rm = TRUE),
+    mean_kappa = mean(kappa, na.rm = TRUE),
     mean_sen = mean(sensitivity, na.rm = TRUE),
@@
     mean_f1 = mean(f1, na.rm = TRUE),
-    mean_prev_df = mean(prevalence_df, na.rm = TRUE),
     .groups="drop"
   ) %>%
-  pivot_longer(mean_acc:mean_prev_df, names_to = "Indices", values_to = "mean") %>%
+  pivot_longer(mean_acc:mean_f1, names_to = "Metric", values_to = "mean") %>%
   pivot_wider(names_from = model, values_from = mean) %>% 
-  mutate(Indices = recode(Indices,
-                        "mean_acc" = "Accuracy",
+  mutate(Metric = recode(Metric,
+                        "mean_acc" = "Balanced Accuracy",
+                        "mean_kappa" = "Kappa",
                         "mean_sen" = "Sensitivity",
                         "mean_ppv" = "Positive Predictive Value",
-                        "mean_f1" = "F1",
-                        "mean_prev_df" = "Prevalence Difference")) %>%
-  relocate("Indices", .before = "positive_class") %>% 
-  mutate(Indices = factor(Indices, levels=c("Accuracy", 
+                        "mean_f1" = "F1")) %>%
+  relocate("Metric", .before = "positive_class") %>% 
+  mutate(Metric = factor(Metric, levels=c("Balanced Accuracy", 
+                                            "Kappa",
                                             "Sensitivity", 
                                             "Positive Predictive Value", 
-                              "F1", 
-                              "Prevalence Difference"))) %>% 
-  arrange(Indices) %>% 
+                                            "F1"))) %>% 
+  arrange(Metric) %>% 
   rename("Category" = "positive_class",
-         "4s model" = "4_no_pos",
-         "16s model" = "16_no_pos",
-         "30s model" = "30_no_pos")
+         "4 s" = "4_no_pos",
+         "16 s" = "16_no_pos",
+         "30 s" = "30_no_pos")
 
+# Between models: ANOVA
 aov_mdl <- function(index, ctgr){
@@
 aov_r <- map(indx, ~aov_mdl(index=.x, ctgr="r")) %>% 
-  map2_dfr(indx, ~mutate(as.data.frame(.x), Indices = .y)) %>% 
+  map2_dfr(indx, ~mutate(as.data.frame(.x), Metric = .y)) %>% 
   mutate(Category = "r")
 aov_u <- map(indx, ~aov_mdl(index=.x, ctgr="u")) %>% 
-  map2_dfr(indx, ~mutate(as.data.frame(.x), Indices = .y)) %>% 
+  map2_dfr(indx, ~mutate(as.data.frame(.x), Metric = .y)) %>% 
   mutate(Category = "u")
@@
   mutate(df = str_glue("({df}, {df_resi})")) %>% 
-  select(-term, -df_resi) %>% 
-  rename("SS" = "sumsq",
-         "MS" = "meansq",
-         "F" = "statistic",
+  select(-term, -df_resi, -sumsq, -meansq) %>% 
+  rename("F" = "statistic",
          "p" = "p.value") %>% 
-  relocate("Indices", .before = "df") %>% 
+  relocate("Metric", .before = "df") %>% 
   relocate("Category", .before = "df") %>% 
   relocate("df", .before = "p") %>%
-  mutate(Indices = recode(Indices,
-                        "balanced_accuracy" = "Accuracy",
+  mutate(Metric = recode(Metric,
+                        "balanced_accuracy" = "Balanced Accuracy",
+                        "kappa" = "Kappa",
                         "sensitivity" = "Sensitivity",
                         "pos_pred_value" = "Positive Predictive Value",
-                        "f1" = "F1",
-                        "prevalence_df" = "Prevalence Difference")) %>%
-  mutate(Indices = factor(Indices, levels=c("Accuracy", 
+                        "f1" = "F1")) %>%
+  mutate(Metric = factor(Metric, levels=c("Balanced Accuracy", 
+                                            "Kappa",
                                             "Sensitivity", 
                                             "Positive Predictive Value", 
-                              "F1", 
-                              "Prevalence Difference"))) %>% 
-  arrange(Indices)
+                                            "F1"))) %>% 
+  arrange(Metric)
 
-table_btw_mdl <- cbind(table_mean, table_aov %>% select(-Indices, -Category)) %>% 
-  slice(-2) %>% 
+# Combine descriptive and ANOVA results and put them in a flex table
+table_btw_mdl <- cbind(table_mean, table_aov %>% select(-Metric, -Category)) %>% 
+  slice(-2,-4) %>% 
   mutate(p = ifelse(p < .001,"<.001",round(p,3)),
@@
   mutate(across(where(is.numeric), ~ round(.x, 3)))
-  # Still need to combine some rows, and add a heading as "accuracy" for three models 
+
+table_btw_mdl[1,2] <- ""
+table_btw_mdl[2,2] <- ""
+
+table_flex_btw_mdl <- table_btw_mdl %>% 
+  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
+  merge_v(j = "Metric") %>% fix_border_issues() %>% 
+  hline(i = c(1,2,4,6,8)) %>% 
+  padding(padding.top = 2, padding.bottom = 2) %>% 
+  set_table_properties(layout = "autofit", width = .99) %>% 
+  align(align = "center", part = "all")
+table_flex_btw_mdl
 ```
@@
 ```{r tbl-indexbetweenctgr}
+#| apa-note: Age in months served as a continuous predictor and was centered within each age group; skill was a categorical predictor with non-sitters/non-walkers serving as the reference categories
+#| tbl-cap: POSTPONE-Models' prediction performance for unrestrained periods compared to restrained periods
+#| ft.align: left
+
 # Compare between categories while controlling for models
@@
 
+#
+t <- matrix_comb %>% 
+  filter (model=="30_no_pos") %>% 
+  select(id, session, positive_class, sensitivity) %>% 
+  pivot_wider(names_from = positive_class, values_from = sensitivity) %>% 
+  mutate(df = u-r)
+summary(lm(df ~ 1, data = t))
+
+#
 tbl_ctgr_30 <- lm_ctgr("30_no_pos")%>% 
@@
   mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
-  add_column(Term = rep(c("Intercept","Prediction of Unrestrained"), 4), .before = "Estimate") %>% 
+  add_column(Term = rep(c("Int.","Unrest. vs Rest."), 4), .before = "Estimate") %>% 
   select(-SE)
@@
   mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
-  add_column(Term = rep(c("Intercept","Prediction of Unrestrained"), 4), .before = "Estimate") %>% 
+  add_column(Term = rep(c("Int.","Unrest. vs Rest."), 4), .before = "Estimate") %>% 
   select(-SE)
 
-tbl_ctgr <- cbind(tbl_ctgr_4, tbl_ctgr_16 %>% select(-Indices,-Term)) %>% 
-  cbind(tbl_ctgr_30 %>% select(-Indices, -Term))
+tbl_ctgr <- cbind(tbl_ctgr_4, tbl_ctgr_16 %>% select(-Indices,-Term) %>% rename(Estimate_16=Estimate,t_16=t,p_16=p)) %>% 
+  cbind(tbl_ctgr_30 %>% select(-Indices, -Term) %>% rename(Estimate_30=Estimate,t_30=t,p_30=p))
 
-# ctgr_4 <- ctgr_4 %>% 
-#   flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
-#   merge_v(j = "Indices") %>% fix_border_issues() %>% 
-#   hline(i = seq(2,8,2))%>% 
-#   padding(padding.top = 2, padding.bottom = 2) %>% 
-#   set_table_properties(layout = "autofit", width = .6) %>% 
-#   align(align = "center", part = "all")
-
+tbl_flex_ctgr <- tbl_ctgr %>% 
+  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>%
+  set_header_labels(Estimate_16 = "Estimate", t_16 = "t", p_16 = "p",
+                    Estimate_30 = "Estimate", t_30 = "t", p_30 = "p") %>%
+  merge_v(j = "Indices") %>% fix_border_issues() %>% 
+  hline(i = c(2,4,6,8)) %>% 
+  padding(padding.top = 2, padding.bottom = 2) %>% 
+  set_table_properties(layout = "autofit", width = .99) %>% 
+  align(align = "center", part = "all") %>% 
+  add_header_row(values = c("", "4s Model", "16s Model", "30s Model"), colwidths = c(2,3,3,3))
+tbl_flex_ctgr
 ```
@@
 ```{r fig-indexcomparison}
+#| fig-cap: Performance indices across models 
+#| fig-width: 13
+#| fig-height: 8
+#| apa-note: Each point shows the percent of awake time spent in a body position at each session.
+Younger infants (4-7 months) are plotted separately for non-sitters (light blue) compared to sitters (dark blue) and older infants (11-14 months) are plotted separately for non-walkers (orange) versus walkers (brown).
+Linear best-fit lines are shown within each age and skill group.
+
 # Sensitivity
+sig_sen <- tibble(x = c(1,2,3),
+                  y = c(1.05,1.05,1.05))
+
 p_sen <- ggplot(data=matrix_comb, aes(x=model, y=sensitivity, color=positive_class, shape=model))+
@@
                color = "black") +
+  geom_text(data = sig_sen, mapping = aes(x = x, y = y, label = "*"), 
+            inherit.aes = FALSE, size = 15, color="red") +
   labs(x="",
@@
   scale_y_continuous(
-    limits = c(0, 1),  
+    limits = c(0, 1.1),  
     breaks = c(0, 0.25, 0.5, 0.75, 1) 
@@
   theme_update()
-p_sen
 
 # Positive Predictive Value
+sig_ppv <- tibble(x = c(1),
+                  y = c(1.05))
 p_ppv <- ggplot(data=matrix_comb, aes(x=model, y=pos_pred_value, color=positive_class, shape=model))+
@@
                color = "black") +
+  geom_text(data = sig_ppv, mapping = aes(x = x, y = y, label = "*"), 
+            inherit.aes = FALSE, size = 15, color="red") +
   labs(x="",
@@
   theme_update()
-p_ppv
+
 # F1 score
@@
   theme_update()
-p_f1
+
 # Prevalence Difference
+sig_prev <- tibble(x = c(1, 1.35, 2, 2.35, 3),
+                  y = c(0.75,0,0.75,0,0.75))
 p_prev_df <- ggplot(data=matrix_comb, aes(x=model, y=prevalence_df, shape=model, color=positive_class))+
@@
                color = "black") +
+  geom_text(data = sig_prev, mapping = aes(x = x, y = y, label = "*"), 
+            inherit.aes = FALSE, size = 15, color="red") +
   labs(x="Models' Window Size",
@@
   theme_update()
-p_prev_df
 
+(p_sen + p_ppv)/(p_f1 + p_prev_df)
 # figure arrangement
-ggpubr::ggarrange(p_sen, p_ppv, p_f1, p_prev_df, ncol = 2, nrow = 2, 
-                  labels = c("A", "B", "C", "D"), 
-                  label.x = 0, 
-                  label.y = 1,
-                  font.label = list(size = 16, face = "bold"),
-                  widths=c(6,7), heights=c(4,4))
-ggsave("figure/indices.png", width=13, height=8)
-ggsave("figure/indices.eps", width=13, height=8)
+# ggpubr::ggarrange(p_sen, p_ppv, p_f1, p_prev_df, ncol = 2, nrow = 2, 
+#                   labels = c("A", "B", "C", "D"), 
+#                   label.x = 0, 
+#                   label.y = 1,
+#                   font.label = list(size = 16, face = "bold"),
+#                   widths=c(6,7), heights=c(4,4))
+# ggsave("figure/indices.png", width=13, height=8)
+# ggsave("figure/indices.eps", width=13, height=8)
 ```
 
-We evaluated the following indices: (1) accuracy, (2) sensitivity, (3) positive predictive value, (4) F1, and (5) prevalence difference for all three models.
-Prevalence difference is a subtraction of a model prediction’s prevalence of certain category from the human annotation’s prevalence, a positive value indicating the models’ overall tendency to overestimate a category.
-
-Models performed equally well in all indices. (Table) Indices mean and ANOVA results. XXX
+@tbl-indexbetweenmodels shows that all three models reached high accuracy in predicting restrained periods (`r apa_num(table_btw_mdl[1,3]*100)`% for 4s model, `r apa_num(table_btw_mdl[1,4]*100)`% for 16s model, and `r apa_num(table_btw_mdl[1,5]*100)`% for 30s model).
+All three models showed high sensitivity (ranging from `r apa_num(min(table_btw_mdl[2,3:5]))` to `r apa_num(max(table_btw_mdl[2,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[3,3:5]))` to `r apa_num(max(table_btw_mdl[3,3:5]))` for unrestrained prediction), high positive predictive value (ranging from `r apa_num(min(table_btw_mdl[4,3:5]))` to `r apa_num(max(table_btw_mdl[4,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[5,3:5]))` to `r apa_num(max(table_btw_mdl[5,3:5]))` for unrestrained prediction), and high F1 score (ranging from `r apa_num(min(table_btw_mdl[6,3:5]))` to `r apa_num(max(table_btw_mdl[6,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[7,3:5]))` to `r apa_num(max(table_btw_mdl[7,3:5]))` for unrestrained prediction).
+Moreover, the three models do not have significant differences in either of the performance indices for each prediction category.
 
 Despite the overall good performance, the models presented predicting bias towards unrestrained category.
-(Table + Figure).
-The sensitivity for predicting unrestrained periods was significantly higher for all models.
-The positive predictive value for predicting unrestrained periods was generally lower for all models, but 16s and 30s models did not significantly over-predict the unrestrained periods as 4s model did.
-The F1 value for both categories was comparable regardless of the models’ window size.
-The prevalence difference value for unrestrained category was significantly higher for all models, but 30s model did not significantly overestimate unrestrained periods or underestimate restrained periods.
+@tbl-indexbetweenctgr and @fig-indexcomparison A shows that all three models were more sensitive predicting unrestrained periods (*p*s \<= `r apa_num(tbl_ctgr[2,11])`).
+It means the models tended to successfully detect an unrestrained period when it actually happened, but could not successfully detect a true restrained period.
+@fig-indexcomparison B shows that the 4s model had a significantly lower positive predictive value for unrestrained periods compared to restrained periods (*p* = `r apa_num(tbl_ctgr[4,5])`), while the 16s model and 30s model presented an insignificant tendency.
+It means the models, especially 4s model, tended to unreliably report an unrestrained period when it did not happen.
+@fig-indexcomparison D shows the prevalence difference of models, defined as the model’s predicted percentage of time of a certain category minus the human’s predicted percentage of time—a positive value means the model predicts more time for certain category than human annotation.
+All three models tended to estimate more unrestrained periods than restrained periods (*p*s \<= `r apa_num(tbl_ctgr[8,11])`).
+Moreover, the 4s model and 16s model's predicted prevalence for unrestrained periods was significantly higher than human annotated prevalence (*p*s \<= `r apa_num(tbl_ctgr[7,8])`).
+The 30s model did not show a significant bias for overpredicting unrestrained periods.
+The possible reasons for performance differences are explained in discussion section.
+Additionally, all three models presented comparable F1 values for both categories (*p*s \<= `r apa_num(tbl_ctgr[6,5])`).
 
-An ideal model should present high accuracy, balanced performance across categories, and minimal offset in overall prevalence compared to ground truth.
-Based on these criteria, we selected the 30-s window model and applied it to process the full-day data.
+With all indices taken into consideration, the 30s model turns out to be a relatively good model out of three, with a high accuracy, a balanced performance across categories, and a minimal offset in overall prevalence compared to human annotation.
+We then applied the 30s model to process the full-day sensor data.
 
 ## Model Estimate Compared to Previous Studies
+
 ```{r}
-load("whole-model-prediction.RData")
-rm(ds, ds_sum_excl, ds_sum_excl0, id_demo, id_info)
 # exclude sessions where total recording time is <3 hours
@@
   filter(unique_id != "148/1",
-         unique_id != "178/3",
-         unique_id != "185/3")
-```
+         unique_id != "178/3")
 
-```{r}
-ds_sum_excl %>% 
-  group_by(age_mo) %>% 
-  summarize(mean = mean(unrestrained_prop),
-            sd = sd(unrestrained_prop),
+stats_fullday <- ds_sum_excl %>% 
+  group_by(age_group) %>% 
+  summarize(mean_restrained = mean(restrained_prop),
+            sd_restrained = sd(restrained_prop),
             .groups="drop")
-ds_sum_excl %>% 
-  group_by(age_group) %>% 
-  summarize(mean = mean(unrestrained_prop),
-            sd = sd(unrestrained_prop),
-            .groups="drop")
-
-summary(lm(unrestrained_prop ~ age_group, data = ds_sum_excl)) # t test
-
-ggplot(ds_sum_excl) +
-  geom_point(aes(x = agemo, y = unrestrained_prop)) + 
-  geom_smooth(aes(x = agemo, y = unrestrained_prop), method = "lm", se = T) + 
-  scale_x_continuous(name = "Age (mo.)", breaks = c(4,7,11,14), limits = c(3,15)) +
-  scale_y_continuous(name = "Unrestrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))
-ggsave("figure/age_related_change.png",width=6, height=4)
-ggsave("figure/age_related_change.eps",width=6, height=4)
 
@@
 
-Next, we compared the 30s model’s estimates of full day restrained time duration with previous studies (XX).
-We excluded eleven sessions where the total recording time was less than 3 hours.
-We excluded another two sessions where a big chunk of duration should have been excluded due to sensors removal, but we did not have reliable time stamp of when to start excluding.
-We also excluded another one session where the family put the leggings on a wrong infant later into the session.
+```{r fig-agechange}
+#| fig-cap: Restrained time proportion by age 
+#| fig-width: 4
+#| fig-height: 3
+#| apa-note: Each point shows the percent of awake time spent in a body position at each session.
+Younger infants (4-7 months) are plotted separately for non-sitters (light blue) compared to sitters (dark blue) and older infants (11-14 months) are plotted separately for non-walkers (orange) versus walkers (brown).
+Linear best-fit lines are shown within each age and skill group.
 
-Using the rest 129 sessions, the model reported an average of XX% of awake time (*SD* = XX) being restrained for younger infants (4-7 months) and XX% (*SD* = XX) for older infants (11-14 months).
-Figure XX shows a clear decrease in the percent of restrained time from younger infants to older infants.
-These results are consistent with previous findings where XXX.
+t <- summary(lm(restrained_prop ~ age_group, data = ds_sum_excl))$coefficients # t test
 
-```{r}
-# ARCHIVE
-### Prevalence correlation scatter plot
-p_prev_scatter_r <- ggplot(data=matrix_comb %>% 
-                             filter(positive_class=="r" & model=="30_no_pos"), 
-                           aes(x = prevalence, y = detection_prevalence))+
-  geom_point(size = 5, color = class_colors[1]) +
-  geom_abline(intercept = 0, slope = 1, color = "black") +
-  scale_x_continuous(limits = c(0, 1)) +
-  scale_y_continuous(limits = c(0, 1)) +
-  theme_update() +
-  labs(x = "Ground Truth Prevalence", y = "Detection Prevalence",
-       title = "Restrained Prediction by 30s Model")
-p_prev_scatter_r
+p_age <- ggplot(ds_sum_excl) +
+  geom_point(aes(x = agemo, y = restrained_prop),shape = 1,size = 3) + 
+  geom_smooth(aes(x = agemo, y = restrained_prop), method = "lm", se = T) + 
+  scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
+  scale_y_continuous(name = "Restrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
+  theme_update()
+p_age
+# ggsave("figure/age_related_change.png",width=6, height=4)
+# ggsave("figure/age_related_change.eps",width=6, height=4)
 
-p_prev_scatter_u <- ggplot(data=matrix_comb %>% 
-                             filter(positive_class=="u" & model=="30_no_pos"), 
-                           aes(x = prevalence, y = detection_prevalence))+
-  geom_point(size = 5, color = class_colors[2]) +
-  geom_abline(intercept = 0, slope = 1, color = "black") +
-  scale_x_continuous(limits = c(0, 1)) +
-  scale_y_continuous(limits = c(0, 1)) +
-  theme_update() +
-  labs(x = "Ground Truth Prevalence", y = "Detection Prevalence",
-       title = "Unrestrained Pred by 30s Model")
-p_prev_scatter_u
+```
 
-ggpubr::ggarrange(p_prev_scatter_r,p_prev_scatter_u, ncol = 2, nrow = 1, 
-                  labels = c("A", "B"), 
-                  label.x = 0, 
-                  label.y = 1,
-                  font.label = list(size = 16, face = "bold"),
-                  widths=c(5,5), heights=c(5))
-ggsave("figure/prevalence_scatter.png", width=10, height=5)
-ggsave("figure/prevalence_scatter.eps", width=10, height=5)
-```
+In order to compare the model estimate of infants' full day experience with previous studies, we excluded eleven sessions whose total recording time was less than 3 hours.
+We also excluded another two sessions where the family took the garment off for a substantial portion of time but did not log the time stamp of removal.
+The rest 129 sessions have an average recording time of `r apa_num(mean(ds_sum_excl$total_time))` hours (*SD* = `r apa_num(sd(ds_sum_excl$total_time))`).
+
+The model prediction on the 129 full-day sessions converged with previous studies [@Carson2022n; @Franchak2019o; @Franchak2024l].
+Younger infants aged from 4 to 7 months spent *M* = `r apa_num(stats_fullday[1,2]*100)`% (*SD* = `r apa_num(stats_fullday[1,3]*100)`%) of awake time restrained.
+Older infants aged from 11 to 14 months spent *M* = `r apa_num(stats_fullday[2,2]*100)`% (*SD* = `r apa_num(stats_fullday[2,3]*100)`%) of awake time restrained.
+Recall the previous EMA study [@Franchak2024l] that found infants aged from 10 to 13 months spent around 46% of time restrained by both caregiver and devices.
+The current study found a smaller portion of time in infants slightly older using wearable sensors.
+The possible reason behind the discrepancy is explained in the discussion section.
+Moreover, @fig-agechange shows a significant decrease in restrained time in older infants (*t* = `r apa_num(t[2,3])`, *p* = `r apa_num(t[2,4])`), corroborating the convergent validity of the current system.
+
+# Discussions
+
+# References

### 2025-10-02 manuscript_restrained_classification validation.qmd
updated figures and tables

@@
 library(effectsize)
+library(ggbeeswarm)
 ```
@@
 model_shapes <- c(5,1,2) %>% set_names(model_labels)
+model_colors <- c("#0072B2","#E69F00", "#009E73") %>% set_names(c("4 s", "16 s", "30 s"))
 class_labels <- c("r", "u")
 class_shapes <- c(5,1) %>% set_names(class_labels)
-class_colors <- c("#0072B2","#E69F00") %>% set_names(class_labels)
+# class_colors <- c("#0072B2","#E69F00") %>% set_names(class_labels)
 
 indx <- c("balanced_accuracy","kappa", "sensitivity","pos_pred_value","f1")
-indices <- c("Sensitivity", "Positive Predictive Value", "F1", "Prevalence Difference")
+per_class_indx <- c("sensitivity","pos_pred_value","f1")
 
@@
 
-```{r tbl-indexbetweenctgr}
-#| apa-note: Age in months served as a continuous predictor and was centered within each age group; skill was a categorical predictor with non-sitters/non-walkers serving as the reference categories
-#| tbl-cap: POSTPONE-Models' prediction performance for unrestrained periods compared to restrained periods
+```{r tbl-cibtwctg}
+#| apa-note: Compare Between Categories
+#| tbl-cap: XX
 #| ft.align: left
+# For per-class metrics (i.e., sens, ppv, f1), calculate the CI of the metric difference btw classes (i.e., r and u)
 
-# Compare between categories while controlling for models
-lm_ctgr <- function(model_select){
-  temp_ds <- matrix_comb %>% filter(model==model_select)
+# Function to bootstrap CI of gap for each model and each metric
+bootstrap_gap <- function(df, metric, B = 2000) {
+  df_wide <- df %>%
+    select(session, id, positive_class, !!sym(metric)) %>%
+    pivot_wider(names_from = positive_class, values_from = !!sym(metric)) %>%
+    mutate(gap = r - u,
+           id_session = paste(id, session, sep = "_"))
+
+  sessions <- unique(df_wide$id_session)
+  n <- length(sessions)
+  res <- numeric(B)
+
+  for (b in seq_len(B)) {
+    sampled <- sample(sessions, size = n, replace = TRUE)
+    df_b <- df_wide %>% filter(id_session %in% sampled)
+    res[b] <- mean(df_b$gap, na.rm = TRUE)
+  }
+
+  ci <- quantile(res, c(0.025, 0.975), na.rm = TRUE)
+
   list(
-    sen = summary(lm(sensitivity ~ positive_class, data=temp_ds))$coefficients,
-    ppv = summary(lm(pos_pred_value ~ positive_class, data=temp_ds))$coefficients,
-    f1  = summary(lm(f1 ~ positive_class, data=temp_ds))$coefficients,
-    prev_df = summary(lm(prevalence_df ~ positive_class, data=temp_ds))$coefficients
+    metric = metric,
+    mean_gap = mean(df_wide$gap, na.rm = TRUE),
+    ci = ci
   )
@@
 
-#
-t <- matrix_comb %>% 
-  filter (model=="30_no_pos") %>% 
-  select(id, session, positive_class, sensitivity) %>% 
-  pivot_wider(names_from = positive_class, values_from = sensitivity) %>% 
-  mutate(df = u-r)
-summary(lm(df ~ 1, data = t))
+table_btw_ctg <- crossing(model = model_labels, metric = per_class_indx) %>%   # all combinations
+  mutate(result = map2(model, metric, 
+                       ~ bootstrap_gap(df = matrix_comb %>% filter(model == .x), 
+                                       metric = .y))) %>%
+  mutate(result = map(result, ~ tibble(
+    mean_gap = .x$mean_gap,
+    ci_low   = .x$ci[1],
+    ci_high  = .x$ci[2]
+  ))) %>% 
+  unnest(result) %>% 
+  mutate(model = factor(model, levels = c("4_no_pos","16_no_pos","30_no_pos")),
+         metric = factor(metric, levels = c("sensitivity","pos_pred_value","f1"))) %>% 
+  mutate(model = recode(model,
+                        "4_no_pos" = "4 s",
+                        "16_no_pos" = "16 s",
+                        "30_no_pos" = "30 s"),
+         metric = recode(metric,
+                         "sensitivity" = "Sensitivity",
+                         "pos_pred_value" = "Positive Predictive Value",
+                         "f1" = "F1")) %>% 
+  arrange(model, metric) %>% 
+  rename("Window Length" = "model",
+         "Metric" = "metric",
+         "M (Rest.- Unrest.)"="mean_gap",
+         "2.5% CI (Rest.- Unrest.)"="ci_low",
+         "97.5% CI (Rest.- Unrest.)"="ci_high")
 
-#
-tbl_ctgr_30 <- lm_ctgr("30_no_pos")%>% 
-  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
-  relocate("Indices", .before = "Estimate")%>% 
-  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
-  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
-         p = str_remove(p, "^0")) %>% 
-  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
-  add_column(Term = rep(c("Intercept","Prediction of Unrestrained"), 4), .before = "Estimate") %>% 
-  select(-SE)
-
-tbl_ctgr_16 <- lm_ctgr("16_no_pos") %>% 
-  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
-  relocate("Indices", .before = "Estimate")%>% 
-  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
-  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
-         p = str_remove(p, "^0")) %>% 
-  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
-  add_column(Term = rep(c("Int.","Unrest. vs Rest."), 4), .before = "Estimate") %>% 
-  select(-SE)
-
-tbl_ctgr_4 <- lm_ctgr("4_no_pos") %>% 
-  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
-  relocate("Indices", .before = "Estimate")%>% 
-  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
-  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
-         p = str_remove(p, "^0")) %>% 
-  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
-  add_column(Term = rep(c("Int.","Unrest. vs Rest."), 4), .before = "Estimate") %>% 
-  select(-SE)
-
-tbl_ctgr <- cbind(tbl_ctgr_4, tbl_ctgr_16 %>% select(-Indices,-Term) %>% rename(Estimate_16=Estimate,t_16=t,p_16=p)) %>% 
-  cbind(tbl_ctgr_30 %>% select(-Indices, -Term) %>% rename(Estimate_30=Estimate,t_30=t,p_30=p))
-
-tbl_flex_ctgr <- tbl_ctgr %>% 
+table_flex_btw_ctg <- table_btw_ctg %>% 
   flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
-  set_header_labels(Estimate_16 = "Estimate", t_16 = "t", p_16 = "p",
-                    Estimate_30 = "Estimate", t_30 = "t", p_30 = "p") %>%
-  merge_v(j = "Indices") %>% fix_border_issues() %>% 
-  hline(i = c(2,4,6,8)) %>% 
+  merge_v(j = "Window Length") %>% fix_border_issues() %>% 
+  hline(i = c(3,6,9)) %>% 
   padding(padding.top = 2, padding.bottom = 2) %>% 
   set_table_properties(layout = "autofit", width = .99) %>% 
-  align(align = "center", part = "all") %>% 
-  add_header_row(values = c("", "4s Model", "16s Model", "30s Model"), colwidths = c(2,3,3,3))
-tbl_flex_ctgr
+  align(align = "center", part = "all")
+table_flex_btw_ctg
+```
+
+```{r fig-cibtwctg}
+#| fig-cap: Compare between categories
+#| fig-width: 7
+#| fig-height: 5
+#| apa-note: XX
+p_ci <- ggplot(table_btw_ctg, 
+       aes(x = Metric, 
+           y = `M (Rest.- Unrest.)`, 
+           ymin = `2.5% CI (Rest.- Unrest.)`, 
+           ymax = `97.5% CI (Rest.- Unrest.)`, 
+           color = `Window Length`)) +
+  geom_pointrange(position = position_dodge(width = 0.5), size = 0.7) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
+  scale_color_manual(values=model_colors,labels=c("4 s","16 s", "30 s")) +
+  scale_y_continuous(
+    limits = c(-0.15, 0.1),  
+    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1) 
+  )+
+  theme_update() +
+  labs(x = NULL,y = "CI of metric difference between restrained and unrestrained",
+    color = "Window Length")
+
+p_ci
+# ggsave("figure/CI.png",width=7, height=5)
+# ggsave("figure/CI.eps",width=7, height=5)
+
+```
+
+```{r tbl-prevalencegap}
+#| apa-note: Diiference of Restraint Prevalence Between Model and Human
+#| tbl-cap: XX
+#| ft.align: left
+
+df_prev <- matrix_comb %>% 
+  filter(positive_class=="r") %>% 
+  select(model, id, session, detection_prevalence, prevalence)
+
+paired_t <- function(df){
+  tt<-t.test(df$detection_prevalence, df$prevalence, paired=T)
+  list(
+    t = tt$statistic,
+    p = tt$p.value
+  )
+}
+
+table_prev <- df_prev %>% 
+  group_by(model) %>% 
+  summarise(actual_prev = mean(detection_prevalence),
+            prev = mean(prevalence),
+            .groups = "drop") %>% 
+  mutate(result=map(model_labels, ~paired_t(df_prev %>% filter(model==.x))))%>% 
+  mutate(result = map(result, ~tibble(
+    t = .x$t,
+    p = .x$p
+  ))) %>% 
+  unnest(result) %>% 
+  mutate(model=recode(model,
+                      "4_no_pos" = "4 s",
+                       "16_no_pos" = "16 s",
+                       "30_no_pos" = "30 s")) %>% 
+  rename("Window Length" = "model",
+         "Human Annotation Prevalence" = "actual_prev",
+         "Model Prediction Prevalence" = "prev")
+
+table_flex_prev <- table_prev %>% 
+  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
+  fix_border_issues() %>% 
+  hline(i = c(1,2,3)) %>% 
+  padding(padding.top = 2, padding.bottom = 2) %>% 
+  set_table_properties(layout = "autofit", width = .99) %>% 
+  align(align = "center", part = "all")
+table_flex_prev
+
+```
+
+```{r fig-prevalence}
+#| fig-cap: Prevalence Correlation
+#| fig-width: 7
+#| fig-height: 5
+#| apa-note: XX
+p_prev <- ggplot(data=matrix_comb %>% filter(positive_class=="r") %>%
+                   mutate(model=recode(model,
+                                       "4_no_pos" = "4 s",
+                                       "16_no_pos" = "16 s",
+                                       "30_no_pos" = "30 s")),
+                 aes(x=prevalence, y=detection_prevalence, color=model))+
+  geom_point(shape = 16,size = 5,alpha=.3) + 
+  geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "black") +
+  geom_smooth(method = "lm", se = FALSE, size = 1) +  
+  scale_color_manual(values=model_colors, labels=c("4 s", "16 s","30 s")) +
+  scale_x_continuous(limits = c(0,1), breaks=c(0,.5,1)) +
+  scale_y_continuous(limits = c(0,1), breaks=c(0,.5,1))+
+  labs(x="Prevalence of restraint by human annotation", y="Prevalence of restraint by model prediction", color="Window Length")+
+  theme_update()
+p_prev
+
+```
+
+```{r fig-prevalencegap}
+#| fig-cap: Prevalence Gap 
+#| fig-width: 7
+#| fig-height: 5
+#| apa-note: XX
+p_prev_gap <- ggplot(data=matrix_comb %>% filter(positive_class=="r") %>%
+                   mutate(model=recode(model,
+                                       "4_no_pos" = "4 s",
+                                       "16_no_pos" = "16 s",
+                                       "30_no_pos" = "30 s")),
+                 aes(x=model, y=detection_prevalence-prevalence, color=model))+
+  geom_violin(trim = FALSE, alpha = 0.4) +
+  geom_quasirandom(aes(color = model),
+                   dodge.width = 0.7, varwidth = TRUE, alpha = 0.6, size = 2) +
+  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white") + 
+  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
+  scale_color_manual(values=model_colors, labels=c("4 s", "16 s","30 s")) +
+  labs(x="", y="Detection prevalence - true prevalence of restraint")+
+  guides(color = "none")+
+  theme_update()
+p_prev_gap
 ```

### 2025-10-05 manuscript_restrained_classification validation.qmd
Updated results

--- 2025-10-01 manuscript_restrained_classification validation.qmd
@@
 model_shapes <- c(5,1,2) %>% set_names(model_labels)
+model_colors <- c("#0072B2","#E69F00", "#009E73") %>% set_names(c("4 s", "16 s", "30 s"))
 class_labels <- c("r", "u")
 class_shapes <- c(5,1) %>% set_names(class_labels)
-class_colors <- c("#0072B2","#E69F00") %>% set_names(class_labels)
+# class_colors <- c("#0072B2","#E69F00") %>% set_names(class_labels)
 
-indx <- c("balanced_accuracy","kappa", "sensitivity","pos_pred_value","f1")
-indices <- c("Sensitivity", "Positive Predictive Value", "F1", "Prevalence Difference")
+indx <- c("accuracy","kappa", "sensitivity","pos_pred_value","f1")
+per_class_indx <- c("sensitivity","pos_pred_value","f1")
 
@@
 ```
+
+[NEED TO BE INCORPORATED WITH WORD DOC]
 
@@
 
-## Data Sharing
+# Results
+We first compared the model’s performance to the human annotation of the first 1.5 hours across all 142 sessions using leave-one-out cross-validation approach.
+We reported performance metrics including accuracy, kappa, sensitivity, positive predictive value, F1, and the correlation of overall estimate between the modal and human annotation.
+We also examined prediction biases of the model by looking at class-specific metrics--sensitivity and positive predictive value--and the prediction error of the overall estimate.
+Second, we applied the model to process infants’ full day sensor data and examined the model’s convergent validity by comparing its estimate and age-related change to previous works.
 
-NOT SURE HOW WE ARE GOING TO SHARE THE DATA
-
-# Results
-
-We first compared the model’s performance to the human annotation of the first 1.5 hours across all 142 sessions.
-We then compared the model’s estimate of infants' full day restrained time and its developmental change to previous works.
-
-## Model Performance Compared to Human Annotation
+## Model Prediction Compared to Human Annotation
 
@@
 ```{r tbl-indexbetweenmodels}
-#| apa-note: Age in months served as a continuous predictor and was centered within each age group; skill was a categorical predictor with non-sitters/non-walkers serving as the reference categories
-#| tbl-cap: Performance indices and ANOVA comparison across models
+#| apa-note: 
+#| tbl-cap: Performance metrics for each window length and ANOVA comparison across window lengths
 #| ft.align: left
@@
   summarise(
-    mean_acc = mean(balanced_accuracy, na.rm = TRUE),
+    mean_acc = mean(accuracy, na.rm = TRUE),
     mean_kappa = mean(kappa, na.rm = TRUE),
@@
   mutate(Metric = recode(Metric,
-                        "mean_acc" = "Balanced Accuracy",
+                        "mean_acc" = "Accuracy",
                         "mean_kappa" = "Kappa",
                         "mean_sen" = "Sensitivity",
-                        "mean_ppv" = "Positive Predictive Value",
-                        "mean_f1" = "F1")) %>%
+                        "mean_ppv" = "PPV",
+                        "mean_f1" = "F1"),
+         positive_class = recode (positive_class,
+                                  "r" = "Restrained",
+                                  "u" = "Unrestrained")) %>%
   relocate("Metric", .before = "positive_class") %>% 
-  mutate(Metric = factor(Metric, levels=c("Balanced Accuracy", 
+  mutate(Metric = factor(Metric, levels=c("Accuracy", 
                                             "Kappa",
                                             "Sensitivity", 
-                                            "Positive Predictive Value", 
+                                            "PPV", 
                                             "F1"))) %>% 
@@
   mutate(Metric = recode(Metric,
-                        "balanced_accuracy" = "Balanced Accuracy",
+                        "accuracy" = "Accuracy",
                         "kappa" = "Kappa",
                         "sensitivity" = "Sensitivity",
-                        "pos_pred_value" = "Positive Predictive Value",
+                        "pos_pred_value" = "PPV",
                         "f1" = "F1")) %>%
-  mutate(Metric = factor(Metric, levels=c("Balanced Accuracy", 
+  mutate(Metric = factor(Metric, levels=c("Accuracy", 
                                             "Kappa",
                                             "Sensitivity", 
-                                            "Positive Predictive Value", 
+                                            "PPV", 
                                             "F1"))) %>% 
@@
 
-```{r tbl-indexbetweenctgr}
-#| apa-note: Age in months served as a continuous predictor and was centered within each age group; skill was a categorical predictor with non-sitters/non-walkers serving as the reference categories
-#| tbl-cap: POSTPONE-Models' prediction performance for unrestrained periods compared to restrained periods
+```{r tbl-cibtwctg}
+#| apa-note: Differences were calculated as restrained – unrestrained metric value.
+M = mean difference; CI = bootstrapped confidence interval.
+All window lengths had lower sensitivity and higher PPV for restrained period detection.
+#| tbl-cap: Differences in sensitivity and positive predictive value (PPV) between restrained and unrestrained prediction across window lengths
 #| ft.align: left
+# For per-class metrics (i.e., sens, ppv, f1), calculate the CI of the metric difference btw classes (i.e., r and u)
 
-# Compare between categories while controlling for models
-lm_ctgr <- function(model_select){
-  temp_ds <- matrix_comb %>% filter(model==model_select)
+# Function to bootstrap CI of gap for each model and each metric
+bootstrap_gap <- function(df, metric, B = 2000, gap_cal = F) {
+  if (gap_cal == F){
+    df_wide <- df %>%
+    select(session, id, positive_class, !!sym(metric)) %>%
+    pivot_wider(names_from = positive_class, values_from = !!sym(metric)) %>%
+    mutate(gap = r - u,
+           id_session = paste(id, session, sep = "_"))
+  } else{
+    df_wide <- df %>% 
+      mutate(id_session = paste(id, session, sep = "_"))
+  } # if gap column is calculated or not
+
+  sessions <- unique(df_wide$id_session)
+  n <- length(sessions)
+  res <- numeric(B)
+
+  for (b in seq_len(B)) {
+    sampled <- sample(sessions, size = n, replace = TRUE)
+    df_b <- df_wide %>% filter(id_session %in% sampled)
+    res[b] <- mean(df_b$gap, na.rm = TRUE)
+  }
+
+  ci <- quantile(res, c(0.025, 0.975), na.rm = TRUE)
+
   list(
-    sen = summary(lm(sensitivity ~ positive_class, data=temp_ds))$coefficients,
-    ppv = summary(lm(pos_pred_value ~ positive_class, data=temp_ds))$coefficients,
-    f1  = summary(lm(f1 ~ positive_class, data=temp_ds))$coefficients,
-    prev_df = summary(lm(prevalence_df ~ positive_class, data=temp_ds))$coefficients
+    metric = metric,
+    mean_gap = mean(df_wide$gap, na.rm = TRUE),
+    ci = ci
   )
@@
 
-#
-t <- matrix_comb %>% 
-  filter (model=="30_no_pos") %>% 
-  select(id, session, positive_class, sensitivity) %>% 
-  pivot_wider(names_from = positive_class, values_from = sensitivity) %>% 
-  mutate(df = u-r)
-summary(lm(df ~ 1, data = t))
+table_btw_ctg <- crossing(model = model_labels, metric = c("sensitivity","pos_pred_value")) %>%   # all combinations
+  mutate(result = map2(model, metric, 
+                       ~ bootstrap_gap(df = matrix_comb %>% filter(model == .x), 
+                                       metric = .y))) %>%
+  mutate(result = map(result, ~ tibble(
+    mean_gap = .x$mean_gap,
+    ci_low   = .x$ci[1],
+    ci_high  = .x$ci[2]
+  ))) %>% 
+  unnest(result) %>% 
+  mutate(model = factor(model, levels = c("4_no_pos","16_no_pos","30_no_pos")),
+         metric = factor(metric, levels = c("sensitivity","pos_pred_value"))) %>% 
+  mutate(model = recode(model,
+                        "4_no_pos" = "4 s",
+                        "16_no_pos" = "16 s",
+                        "30_no_pos" = "30 s"),
+         metric = recode(metric,
+                         "sensitivity" = "Sensitivity",
+                         "pos_pred_value" = "PPV")) %>% 
+  arrange(model, metric) %>% 
+  rename("Window Length" = "model",
+         "Metric" = "metric",
+         "M"="mean_gap",
+         "95%CI lower bound"="ci_low",
+         "95%CI upper bound"="ci_high")
 
-#
-tbl_ctgr_30 <- lm_ctgr("30_no_pos")%>% 
-  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
-  relocate("Indices", .before = "Estimate")%>% 
-  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
-  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
-         p = str_remove(p, "^0")) %>% 
-  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
-  add_column(Term = rep(c("Intercept","Prediction of Unrestrained"), 4), .before = "Estimate") %>% 
-  select(-SE)
-
-tbl_ctgr_16 <- lm_ctgr("16_no_pos") %>% 
-  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
-  relocate("Indices", .before = "Estimate")%>% 
-  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
-  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
-         p = str_remove(p, "^0")) %>% 
-  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
-  add_column(Term = rep(c("Int.","Unrest. vs Rest."), 4), .before = "Estimate") %>% 
-  select(-SE)
-
-tbl_ctgr_4 <- lm_ctgr("4_no_pos") %>% 
-  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
-  relocate("Indices", .before = "Estimate")%>% 
-  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
-  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
-         p = str_remove(p, "^0")) %>% 
-  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
-  add_column(Term = rep(c("Int.","Unrest. vs Rest."), 4), .before = "Estimate") %>% 
-  select(-SE)
-
-tbl_ctgr <- cbind(tbl_ctgr_4, tbl_ctgr_16 %>% select(-Indices,-Term) %>% rename(Estimate_16=Estimate,t_16=t,p_16=p)) %>% 
-  cbind(tbl_ctgr_30 %>% select(-Indices, -Term) %>% rename(Estimate_30=Estimate,t_30=t,p_30=p))
-
-tbl_flex_ctgr <- tbl_ctgr %>% 
-  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>%
-  set_header_labels(Estimate_16 = "Estimate", t_16 = "t", p_16 = "p",
-                    Estimate_30 = "Estimate", t_30 = "t", p_30 = "p") %>%
-  merge_v(j = "Indices") %>% fix_border_issues() %>% 
-  hline(i = c(2,4,6,8)) %>% 
+table_flex_btw_ctg <- table_btw_ctg %>% 
+  flextable() %>% flextable::theme_apa() %>% 
+  line_spacing(part = "all") %>%
+  merge_v(j = "Window Length") %>% fix_border_issues() %>% 
+  add_header_row(values = c("","", "Difference (rest.- unrest.)"), colwidths = c(1,1,3)) %>% 
+  hline(i = c(2,4,6), part = "body") %>%
   padding(padding.top = 2, padding.bottom = 2) %>% 
   set_table_properties(layout = "autofit", width = .99) %>% 
-  align(align = "center", part = "all") %>% 
-  add_header_row(values = c("", "4s Model", "16s Model", "30s Model"), colwidths = c(2,3,3,3))
-tbl_flex_ctgr
+  align(align = "center", part = "all")
+table_flex_btw_ctg
 ```
 
-```{r fig-indexcomparison}
-#| fig-cap: Performance indices across models 
-#| fig-width: 13
-#| fig-height: 8
-#| apa-note: Each point shows the percent of awake time spent in a body position at each session.
-Younger infants (4-7 months) are plotted separately for non-sitters (light blue) compared to sitters (dark blue) and older infants (11-14 months) are plotted separately for non-walkers (orange) versus walkers (brown).
-Linear best-fit lines are shown within each age and skill group.
+```{r fig-cibtwctg}
+#| fig-cap: Differences in sensitivity and (PPV) between restrained and unrestrained prediction across window lengths
+#| fig-width: 8
+#| fig-height: 5
+#| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction.
+The dot on each line denotes the mean of the difference.
+All window lengths had higher sensitivity for unrestrained period prediction and higher ppv for restrained period prediction.
 
-# Sensitivity
-sig_sen <- tibble(x = c(1,2,3),
-                  y = c(1.05,1.05,1.05))
+p_ci <- ggplot(table_btw_ctg, 
+       aes(x = Metric, 
+           y = `M`, 
+           ymin = `95%CI lower bound`, 
+           ymax = `95%CI upper bound`, 
+           color = `Window Length`)) +
+  geom_pointrange(position = position_dodge(width = 0.5), size = 0.7) +
+  geom_segment(aes(x = 0, xend = 3, y = 0, yend = 0),
+             linetype = "dashed", color = "gray")+
+  scale_color_manual(values=model_colors,labels=c("4 s","16 s", "30 s")) +
+  scale_x_discrete(expand = expansion(add = c(0,2)))+
+  scale_y_continuous(
+    limits = c(-0.15, 0.1),  
+    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1) 
+  )+
+  theme_update() +
+  theme(legend.position = c(0.89,0.1))+ 
+  labs(x = "",y = "Metric value difference (rest.- unrest.)",
+    color = "Window Length")+
+  annotate("text", x = 3.5, y = -0.05, label="Unrestrained prediction \n has higher \n metric's value",color = "black")+
+  annotate("text", x = 3.5, y = 0.05, label="Restrained prediction \n has higher \n metric's value",color = "black")
 
-p_sen <- ggplot(data=matrix_comb, aes(x=model, y=sensitivity, color=positive_class, shape=model))+
-  geom_point(position = position_jitterdodge(
-              jitter.width = 0.4,
-              jitter.height = 0,
-              dodge.width = 0.6),
-             size = 5)+
-  scale_color_manual(values=class_colors,labels=c("restrained","unrestrained")) +
-  scale_shape_manual(values=model_shapes) +
-  stat_summary(fun = mean,
-               geom = "crossbar",
-               aes(group = positive_class),  
-               position = position_dodge(width = 0.6), 
-               width = 0.4,
-               color = "black") +
-  geom_text(data = sig_sen, mapping = aes(x = x, y = y, label = "*"), 
-            inherit.aes = FALSE, size = 15, color="red") +
-  labs(x="",
-       y="Sensitivity",
-       color="Category")+
-  scale_x_discrete(labels = c("4s","16s","30s"))+
-  scale_y_continuous(
-    limits = c(0, 1.1),  
-    breaks = c(0, 0.25, 0.5, 0.75, 1) 
-  ) +
-  guides(color = "none",shape="none")+
-  theme_update()
+p_ci
+# ggsave("figure/CI.png",width=7, height=5)
+# ggsave("figure/CI.eps",width=7, height=5)
 
-# Positive Predictive Value
-sig_ppv <- tibble(x = c(1),
-                  y = c(1.05))
-p_ppv <- ggplot(data=matrix_comb, aes(x=model, y=pos_pred_value, color=positive_class, shape=model))+
-  geom_point(position = position_jitterdodge(
-    jitter.width = 0.4,
-    jitter.height = 0,
-    dodge.width = 0.6),
-    size = 5)+
-  scale_color_manual(values=class_colors,labels=c("restrained","unrestrained")) +
-  scale_shape_manual(values=model_shapes) +
-  stat_summary(fun = mean,
-               geom = "crossbar",
-               aes(group = positive_class),  
-               position = position_dodge(width = 0.6), 
-               width = 0.4,
-               color = "black") +
-  geom_text(data = sig_ppv, mapping = aes(x = x, y = y, label = "*"), 
-            inherit.aes = FALSE, size = 15, color="red") +
-  labs(x="",
-       y="Positive Predictive Value",
-       color="Category")+
-  scale_x_discrete(labels = c("4s","16s","30s"))+
-  scale_y_continuous(
-    limits = c(0, 1.1),  
-    breaks = c(0, 0.25, 0.5, 0.75, 1) 
-  ) +
-  guides(shape="none")+
-  theme_update()
-
-# F1 score
-p_f1 <- ggplot(data=matrix_comb, aes(x=model, y=f1, shape=model, color=positive_class))+
-  geom_point(position = position_jitterdodge(
-    jitter.width = 0.4,
-    jitter.height = 0,
-    dodge.width = 0.6),
-    size = 5)+
-  scale_color_manual(values=class_colors,labels=c("restrained","unrestrained")) +
-  scale_shape_manual(values=model_shapes) +
-  stat_summary(fun = mean,
-               geom = "crossbar",
-               aes(group = positive_class),  
-               position = position_dodge(width = 0.6), 
-               width = 0.4,
-               color = "black") +
-  labs(x="Models' Window Size",
-       y="F1",
-       color="Category")+
-  scale_x_discrete(labels = c("4s","16s","30s"))+
-  scale_y_continuous(
-    limits = c(0, 1),  
-    breaks = c(0, 0.25, 0.5, 0.75, 1) 
-  ) +
-  guides(color = "none",shape="none")+
-  theme_update()
-
-# Prevalence Difference
-sig_prev <- tibble(x = c(1, 1.35, 2, 2.35, 3),
-                  y = c(0.75,0,0.75,0,0.75))
-p_prev_df <- ggplot(data=matrix_comb, aes(x=model, y=prevalence_df, shape=model, color=positive_class))+
-  geom_point(position = position_jitterdodge(
-    jitter.width = 0.4,
-    jitter.height = 0,
-    dodge.width = 0.6),
-    size = 5)+
-  geom_hline(yintercept = 0, linetype = 'dashed')+
-  scale_color_manual(values=class_colors,labels=c("restrained","unrestrained")) +
-  scale_shape_manual(values=model_shapes) +
-  stat_summary(fun = mean,
-               geom = "crossbar",
-               aes(group = positive_class),  
-               position = position_dodge(width = 0.6), 
-               width = 0.4,
-               color = "black") +
-  geom_text(data = sig_prev, mapping = aes(x = x, y = y, label = "*"), 
-            inherit.aes = FALSE, size = 15, color="red") +
-  labs(x="Models' Window Size",
-       y="Prevalence Difference",
-       color="Category")+
-  scale_x_discrete(labels = c("4s","16s","30s"))+
-  scale_y_continuous(
-    limits = c(-.8, .8),
-    breaks = c(-.8, -.4, 0, .4, .8)
-  ) +
-  guides(shape = "none")+
-  theme_update()
-
-(p_sen + p_ppv)/(p_f1 + p_prev_df)
-# figure arrangement
-# ggpubr::ggarrange(p_sen, p_ppv, p_f1, p_prev_df, ncol = 2, nrow = 2, 
-#                   labels = c("A", "B", "C", "D"), 
-#                   label.x = 0, 
-#                   label.y = 1,
-#                   font.label = list(size = 16, face = "bold"),
-#                   widths=c(6,7), heights=c(4,4))
-# ggsave("figure/indices.png", width=13, height=8)
-# ggsave("figure/indices.eps", width=13, height=8)
 ```
 
-@tbl-indexbetweenmodels shows that all three models reached high accuracy in predicting restrained periods (`r apa_num(table_btw_mdl[1,3]*100)`% for 4s model, `r apa_num(table_btw_mdl[1,4]*100)`% for 16s model, and `r apa_num(table_btw_mdl[1,5]*100)`% for 30s model).
-All three models showed high sensitivity (ranging from `r apa_num(min(table_btw_mdl[2,3:5]))` to `r apa_num(max(table_btw_mdl[2,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[3,3:5]))` to `r apa_num(max(table_btw_mdl[3,3:5]))` for unrestrained prediction), high positive predictive value (ranging from `r apa_num(min(table_btw_mdl[4,3:5]))` to `r apa_num(max(table_btw_mdl[4,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[5,3:5]))` to `r apa_num(max(table_btw_mdl[5,3:5]))` for unrestrained prediction), and high F1 score (ranging from `r apa_num(min(table_btw_mdl[6,3:5]))` to `r apa_num(max(table_btw_mdl[6,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[7,3:5]))` to `r apa_num(max(table_btw_mdl[7,3:5]))` for unrestrained prediction).
-Moreover, the three models do not have significant differences in either of the performance indices for each prediction category.
+```{r tbl-prevalencecor}
+#| apa-note: Prevalence was calculated as the proportion of restrained periods among all window samples for each session.
+Prediction error was calculated as model prediction prevalence - human annotation prevalence.
+A negative prediction error means an underestimate from model prediction.
+M = mean difference; CI = bootstrapped confidence interval.
+#| tbl-cap: Model predicted and human annotated prevalence of restrained periods, their correlation, and differences
+#| ft.align: left
 
-Despite the overall good performance, the models presented predicting bias towards unrestrained category.
-@tbl-indexbetweenctgr and @fig-indexcomparison A shows that all three models were more sensitive predicting unrestrained periods (*p*s \<= `r apa_num(tbl_ctgr[2,11])`).
-It means the models tended to successfully detect an unrestrained period when it actually happened, but could not successfully detect a true restrained period.
-@fig-indexcomparison B shows that the 4s model had a significantly lower positive predictive value for unrestrained periods compared to restrained periods (*p* = `r apa_num(tbl_ctgr[4,5])`), while the 16s model and 30s model presented an insignificant tendency.
-It means the models, especially 4s model, tended to unreliably report an unrestrained period when it did not happen.
-@fig-indexcomparison D shows the prevalence difference of models, defined as the model’s predicted percentage of time of a certain category minus the human’s predicted percentage of time—a positive value means the model predicts more time for certain category than human annotation.
-All three models tended to estimate more unrestrained periods than restrained periods (*p*s \<= `r apa_num(tbl_ctgr[8,11])`).
-Moreover, the 4s model and 16s model's predicted prevalence for unrestrained periods was significantly higher than human annotated prevalence (*p*s \<= `r apa_num(tbl_ctgr[7,8])`).
-The 30s model did not show a significant bias for overpredicting unrestrained periods.
-The possible reasons for performance differences are explained in discussion section.
-Additionally, all three models presented comparable F1 values for both categories (*p*s \<= `r apa_num(tbl_ctgr[6,5])`).
+df_prev <- matrix_comb %>%
+  filter(positive_class=="r") %>%
+  select(model, id, session, detection_prevalence, prevalence) %>% 
+  mutate(gap = detection_prevalence-prevalence)
 
-With all indices taken into consideration, the 30s model turns out to be a relatively good model out of three, with a high accuracy, a balanced performance across categories, and a minimal offset in overall prevalence compared to human annotation.
-We then applied the 30s model to process the full-day sensor data.
+prev_gap <- map(model_labels, ~ bootstrap_gap(df = df_prev %>% filter(model == .x), metric = "gap", gap_cal = T))
+
+table_prev <- df_prev %>%
+  group_by(model) %>%
+  summarise(actual_prev = mean(prevalence),
+            prev = mean(detection_prevalence),
+            cor = cor (detection_prevalence, prevalence),
+            .groups = "drop") %>% 
+  mutate(gap = map(prev_gap, ~ tibble(
+    mean_gap = .x$mean_gap,
+    ci_low   = .x$ci[1],
+    ci_high  = .x$ci[2]))) %>% 
+  unnest(gap) %>% 
+  mutate(model=recode(model,
+                      "4_no_pos" = "4 s",
+                       "16_no_pos" = "16 s",
+                       "30_no_pos" = "30 s"))
+
+table_prev_100 <- table_prev %>% 
+  mutate(actual_prev=actual_prev*100,
+         prev = prev*100, 
+         mean_gap = mean_gap *100,
+         ci_low=ci_low * 100,
+         ci_high=ci_high*100) %>% 
+  rename("Window Length" = "model",
+         "Human Annotation Prevalence" = "actual_prev",
+         "Model Prediction Prevalence" = "prev",
+         "Correlation" = "cor",
+         "M" = "mean_gap",
+         "95%CI lower bound" = "ci_low",
+         "95%CI upper bound" = "ci_high")
+
+table_flex_prev <- table_prev_100 %>% 
+  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
+  fix_border_issues() %>% 
+  add_header_row(values = c("","Prevalence (%)", "Prediction Error (%)"), colwidths = c(1,3,3)) %>% 
+  hline(i = c(1,2,3)) %>% 
+  padding(padding.top = 2, padding.bottom = 2) %>% 
+  set_table_properties(layout = "autofit", width = .99) %>% 
+  align(align = "center", part = "all")
+table_flex_prev
+```
+
+```{r fig-prevalencegap}
+#| fig-cap: Prediction error between model prediction and human annotation
+#| fig-width: 6
+#| fig-height: 5
+#| apa-note: Prediction error is the difference in the overall restrained time between the model estimate and the human annotation.
+A negative prediction error indicates the model underestimate restrained time overall compared to human annotation.
+Each point denotes one session.
+The violin shape denotes the distribution of prediction error of each window length.
+The box denotes the first (Q1) quartile, the median, and the third (Q3) quartile of prediction error.
+
+p_prev_gap <- ggplot(data=matrix_comb %>% filter(positive_class=="r") %>%
+                   mutate(model=recode(model,
+                                       "4_no_pos" = "4 s",
+                                       "16_no_pos" = "16 s",
+                                       "30_no_pos" = "30 s")),
+                 aes(x=model, y=detection_prevalence-prevalence, color=model))+
+  geom_violin(trim = FALSE, alpha = 0.4) +
+  geom_quasirandom(aes(color = model),
+                   dodge.width = 0.7, varwidth = TRUE, alpha = 0.6, size = 2) +
+  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white") + 
+  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
+  scale_color_manual(values=model_colors, labels=c("4 s", "16 s","30 s")) +
+  labs(x="Window Length", y="Prediction Error (prediction-human annotation)")+
+  guides(color = "none")+
+  theme_update()
+p_prev_gap
+```
+
+### Overall Model Performance
+
+@tbl-indexbetweenmodels shows that all window lengths reached high accuracy (`r apa_num(table_btw_mdl[1,3])` for 4 s, `r apa_num(table_btw_mdl[1,4])` for 16 s, and `r apa_num(table_btw_mdl[1,5])` for 30 s), and good kappa value (`r apa_num(table_btw_mdl[2,3])` for 4 s, `r apa_num(table_btw_mdl[2,4])` for 16 s, and `r apa_num(table_btw_mdl[2,5])` for 30 s).
+All the window lengths also presented high sensitivity (ranging from `r apa_num(min(table_btw_mdl[3,3:5]))` to `r apa_num(max(table_btw_mdl[3,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[4,3:5]))` to `r apa_num(max(table_btw_mdl[4,3:5]))` for unrestrained prediction), high positive predictive value (ranging from `r apa_num(min(table_btw_mdl[5,3:5]))` to `r apa_num(max(table_btw_mdl[5,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[6,3:5]))` to `r apa_num(max(table_btw_mdl[6,3:5]))` for unrestrained prediction), and high F1 score (ranging from `r apa_num(min(table_btw_mdl[7,3:5]))` to `r apa_num(max(table_btw_mdl[7,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[8,3:5]))` to `r apa_num(max(table_btw_mdl[8,3:5]))` for unrestrained prediction).
+Moreover, the three window lengths did not present significant differences in any of the metrics (*p*s >= `r apa_num(min(table_btw_mdl[1:8,8]))`).
+
+Meanwhile, @tbl-prevalencecor shows that all window lengths provided reliable overall estimate of restrained time.
+The prevalence of restrained periods the models predicted was strongly correlated with the prevalence annotated by human coders, the correlation coefficient ranging from `r apa_num(table_prev[1,4])` for 4 s and `r apa_num(table_prev[3,4])` for 30 s.
+
+### Bias Towards Unrestrained Class Across Window Lengths
+
+Despite the overall good performance, all window lengths presented prediction bias towards unrestrained class.
+@tbl-cibtwctg and @fig-cibtwctg show that all window lengths had lower sensitivity and higher positive predictive value (PPV) for restrained compared to unrestrained periods, indicated by the negative difference of sensitivity and positive difference of PPV.
+The sensitivity difference for 4 s window was `r apa_num(table_btw_ctg[1,3])`, 95% bootstrapped CI [`r apa_num(table_btw_ctg[1,4])`, `r apa_num(table_btw_ctg[1,5])`], for 16 s window was `r apa_num(table_btw_ctg[3,3])`, 95% CI [`r apa_num(table_btw_ctg[3,4])`, `r apa_num(table_btw_ctg[3,5])`], and for 30 s window was `r apa_num(table_btw_ctg[5,3])`, 95% CI [`r apa_num(table_btw_ctg[5,4])`, `r apa_num(table_btw_ctg[5,5])`].
+The PPV difference for 4 s window was `r apa_num(table_btw_ctg[2,3])`, 95% CI [`r apa_num(table_btw_ctg[2,4])`, `r apa_num(table_btw_ctg[2,5])`], for 16 s window was `r apa_num(table_btw_ctg[4,3])`, 95% CI [`r apa_num(table_btw_ctg[4,4])`, `r apa_num(table_btw_ctg[4,5])`], and for 30 s window was `r apa_num(table_btw_ctg[6,3])`, 95% CI [`r apa_num(table_btw_ctg[6,4])`, `r apa_num(table_btw_ctg[6,5])`].
+The differences indicate that the model tended to under-detect restrained periods but was reliable when labeling a period as restrained.
+Conversely, it overestimated unrestrained periods, with lower reliability in unrestrained prediction.
+While none of the window lengths were free from bias, the 30 s window presented a more balanced detection between restrained and unrestrained periods.
+
+The prediction bias was also evident when examining the model’s overall estimate of restrained time and its difference from human annotation.
+@tbl-prevalencecor and @fig-prevalencegap show a consistent underestimate of restrained periods despite the high correlation of the overall estimate.
+The 4 s window missed `r apa_num(abs(table_prev[1,5])*100)`% of time that were annotated as restrained (95% CI [`r apa_num(abs(table_prev[1,6])*100)`%, `r apa_num(abs(table_prev[1,7])*100)`%]), the 16 s window missed `r apa_num(abs(table_prev[2,5])*100)`% (95% CI [`r apa_num(abs(table_prev[2,6])*100)`%, `r apa_num(abs(table_prev[2,7])*100)`%]]), and the 30 s window missed the least amount of `r apa_num(abs(table_prev[3,5])*100)`% (95% CI [`r apa_num(abs(table_prev[3,6])*100)`%, `r apa_num(abs(table_prev[3,7])*100)`%]).
+Taken together, 30 s turned out to be the best window length out of the three, which brought about a balanced performance across classes.
+We then applied the 30 s model to process the full-day sensor data.
 
@@
 ```{r fig-agechange}
-#| fig-cap: Restrained time proportion by age 
+#| fig-cap: Infants' full day restrained time proportion with age
 #| fig-width: 4
 #| fig-height: 3
-#| apa-note: Each point shows the percent of awake time spent in a body position at each session.
-Younger infants (4-7 months) are plotted separately for non-sitters (light blue) compared to sitters (dark blue) and older infants (11-14 months) are plotted separately for non-walkers (orange) versus walkers (brown).
-Linear best-fit lines are shown within each age and skill group.
+#| apa-note: Each point denotes one session's restrained time proportion processed by the 30 s model.
+The box denotes the first (Q1) quartile, the median, and the third (Q3) quartile of prediction error.
 
-t <- summary(lm(restrained_prop ~ age_group, data = ds_sum_excl))$coefficients # t test
+t_age <- summary(lm(restrained_prop ~ age_group, data = ds_sum_excl))$coefficients # t test
 
-p_age <- ggplot(ds_sum_excl) +
-  geom_point(aes(x = agemo, y = restrained_prop),shape = 1,size = 3) + 
-  geom_smooth(aes(x = agemo, y = restrained_prop), method = "lm", se = T) + 
-  scale_x_continuous(name = "Age (mo)", breaks = c(4,7,11,14), limits = c(3,15)) +
+ds_sum_excl <- ds_sum_excl %>% 
+  mutate(age_group_num=ifelse(age_group=="Younger",5.5, 12.5))
+
+p_age <- ggplot(ds_sum_excl, aes(x=agemo, y=restrained_prop)) +
+  geom_point(shape = 1,size = 3) + 
+  geom_boxplot(
+    aes(x = age_group_num, y = restrained_prop, group = age_group_num),
+    inherit.aes = FALSE,
+    width = 3,
+    alpha = 0.2,
+    outlier.shape = NA,
+    color = "blue"
+  ) +
+  scale_x_continuous(name = "Age Group", breaks = c(5.5, 12.5), labels=c("Young", "Old"), limits = c(3,15)) +
   scale_y_continuous(name = "Restrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
-  theme_update()
+  theme_update()+
+  labs(x="Age Group")
+
 p_age
@@
 
-In order to compare the model estimate of infants' full day experience with previous studies, we excluded eleven sessions whose total recording time was less than 3 hours.
+In order to compare the model estimate of infants’ full day experience with previous studies, we excluded eleven sessions whose total recording time was less than 3 hours, which were useful for training and validating the models but insufficient for measuring day-long experience.
 We also excluded another two sessions where the family took the garment off for a substantial portion of time but did not log the time stamp of removal.
-The rest 129 sessions have an average recording time of `r apa_num(mean(ds_sum_excl$total_time))` hours (*SD* = `r apa_num(sd(ds_sum_excl$total_time))`).
+The remaining 129 sessions have an average recording time of `r apa_num(mean(ds_sum_excl$total_time))` hours (*SD* = `r apa_num(sd(ds_sum_excl$total_time))`).
 
 The model prediction on the 129 full-day sessions converged with previous studies [@Carson2022n; @Franchak2019o; @Franchak2024l].
-Younger infants aged from 4 to 7 months spent *M* = `r apa_num(stats_fullday[1,2]*100)`% (*SD* = `r apa_num(stats_fullday[1,3]*100)`%) of awake time restrained.
-Older infants aged from 11 to 14 months spent *M* = `r apa_num(stats_fullday[2,2]*100)`% (*SD* = `r apa_num(stats_fullday[2,3]*100)`%) of awake time restrained.
-Recall the previous EMA study [@Franchak2024l] that found infants aged from 10 to 13 months spent around 46% of time restrained by both caregiver and devices.
-The current study found a smaller portion of time in infants slightly older using wearable sensors.
-The possible reason behind the discrepancy is explained in the discussion section.
-Moreover, @fig-agechange shows a significant decrease in restrained time in older infants (*t* = `r apa_num(t[2,3])`, *p* = `r apa_num(t[2,4])`), corroborating the convergent validity of the current system.
+Younger infants aged from 4 to 7 months spent *M* = `r apa_num(stats_fullday[1,2]*100)`% (*SD* = `r apa_num(stats_fullday[1,3]*100)`%) of awake time restrained, and older infants aged from 11 to 14 months spent *M* = `r apa_num(stats_fullday[2,2]*100)`% (*SD* = `r apa_num(stats_fullday[2,3]*100)`%) of awake time restrained, which is comparable to the 46% restrained time for infants aged from 10 to 13 months in the previous EMA study [@Franchak2024l].
+Moreover, @fig-agechange shows a significant group difference in restrained time between older infants and younger infants (*t* = `r apa_num(t_age[2,3])`, *p* `r apa_num(ifelse(t_age[2,4] < .001, "<.001", t_age[2,4]))`), supporting the convergent validity of the current system.
 

### 2025-10-08 manuscript_restrained_classification validation.qmd
Updated discussion

@@
 library(ggbeeswarm)
+library(DescTools)
 ```
@@
 
+```{r}
+# check if the data set is imbalanced 
+matrix_comb %>% 
+  filter(positive_class=="r") %>% 
+  group_by(model) %>% 
+  summarize(mean_prev = mean(prevalence),
+            sd_prev = sd(prevalence),
+            .groups="drop")
+
+ggplot(matrix_comb %>% filter(positive_class=="r"), aes(x = prevalence, color=model)) +
+  geom_density(alpha = 0.4)
+
+#slightly imbalanced: less unrestrained periods, the ratio (~40%) is similar to the total estimated proportion
+```
+
+```{r}
+# check accuracy outliers 
+outlier_4 <- matrix_comb %>%
+  filter(model == "4_no_pos", positive_class=="r") %>%
+  mutate(
+    q1 = quantile(accuracy, 0.25),
+    q3 = quantile(accuracy, 0.75),
+    iqr = q3 - q1,
+    is_outlier = accuracy < (q1 - 1.5 * iqr) | accuracy > (q3 + 1.5 * iqr)
+  ) %>%  filter(is_outlier)
+outlier_16 <- matrix_comb %>%
+  filter(model == "16_no_pos", positive_class=="r") %>%
+  mutate(
+    q1 = quantile(accuracy, 0.25),
+    q3 = quantile(accuracy, 0.75),
+    iqr = q3 - q1,
+    is_outlier = accuracy < (q1 - 1.5 * iqr) | accuracy > (q3 + 1.5 * iqr)
+  ) %>%  filter(is_outlier)
+outlier_30 <- matrix_comb %>%
+  filter(model == "30_no_pos", positive_class=="r") %>%
+  mutate(
+    q1 = quantile(accuracy, 0.25),
+    q3 = quantile(accuracy, 0.75),
+    iqr = q3 - q1,
+    is_outlier = accuracy < (q1 - 1.5 * iqr) | accuracy > (q3 + 1.5 * iqr)
+  ) %>%  filter(is_outlier)
+
+outliers <- rbind(outlier_4,outlier_16, outlier_30)
+
+ggplot(data=outliers, aes(x=session, y=id, color=model))+
+  geom_jitter()+
+  geom_text(aes(label = paste0("(", id, ", ", session, ")")),
+            vjust = -0.5,   # move label a bit above point
+            size = 3) 
+```
+
 ### Overall Model Performance
@@
 The model prediction on the 129 full-day sessions converged with previous studies [@Carson2022n; @Franchak2019o; @Franchak2024l].
-Younger infants aged from 4 to 7 months spent *M* = `r apa_num(stats_fullday[1,2]*100)`% (*SD* = `r apa_num(stats_fullday[1,3]*100)`%) of awake time restrained, and older infants aged from 11 to 14 months spent *M* = `r apa_num(stats_fullday[2,2]*100)`% (*SD* = `r apa_num(stats_fullday[2,3]*100)`%) of awake time restrained, which is comparable to the 46% restrained time for infants aged from 10 to 13 months in the previous EMA study [@Franchak2024l].
+Younger infants aged from 4 to 7 months spent *M* = `r apa_num(stats_fullday[1,2]*100)`% (*SD* = `r apa_num(stats_fullday[
+1,3]*100)`%) of awake time restrained, and older infants aged from 11 to 14 months spent *M* = `r apa_num(stats_fullday[2,2]*100)`% (*SD* = `r apa_num(stats_fullday[2,3]*100)`%) of awake time restrained, which is comparable to the 46% restrained time for infants aged from 10 to 13 months in the previous EMA study [@Franchak2024l].
 Moreover, @fig-agechange shows a significant group difference in restrained time between older infants and younger infants (*t* = `r apa_num(t_age[2,3])`, *p* `r apa_num(ifelse(t_age[2,4] < .001, "<.001", t_age[2,4]))`), supporting the convergent validity of the current system.
@@
 # Discussions
+We validated a system that combines wearable sensors and machine learning models to quantify infants’ restrained experiences in the home environment.
+The model presented high accuracy and substantial kappa agreement with human-annotated ground truth.
+The model also presented convergent validity in overall estimate of restraint time and age-related change compared with previous studies.
+
+We also examined the optimal window length for the classification (4 s, 16 s, and 30 s).
+All window lengths showed a slight prediction bias towards unrestrained class, which might be due to the inherent imbalance in the training data set—unrestrained periods were more prevalent than restrained periods.
+However, the 30 s window offered the most balanced detection compared to shorter windows.
+This advantage might be due to the contextual nature of restraint: infants may exhibit similar body movement whether restrained or not, and a restrained bout usually lasts minutes rather than seconds.
+A longer window can provide richer contextual information for distinguishing restrained from unrestrained states, whereas shorter windows only capture momentary movements that may appear similar across restrained and unrestrained periods.
+
+A key contribution of this work is the extension of wearable-sensor methodology to detect infant restraint in naturalistic settings.
+Wearable sensors have been established as a reliable and unobtrusive approach for capturing full-day, time-series movement data.
+Previous works validated the use of wearable sensors in classifying infant positions.
+Our work shows the configuration of four sensors on thighs and ankles and random forest algorithms can also provide sufficiently accurate classification of infant restraint.
+Although the system cannot be as accurate as human annotation, the trade-off between accuracy and scale favors the sensor-based system, particularly when research questions is oriented at infants’ experiences fluctuating throughout a day or across multiple days.
+
+With the validated system, researchers can now study infant restrained experiences in a longer time scale and a greater temporal precision.
+Research questions about the temporal structure of restrained periods can be addressed—for example, how frequently are infants switched between restrained and unrestrained states, or how restrained bouts are distributed across the day or multiple days?
+Moreover, the system enables investigations linking restraint to concurrent learning experiences: How does the temporal profile of restraint relate to motor development?
+How does adult language input or infants’ own vocalization patterns shift when they are restrained versus unrestrained?
+
+Moving forward, we can refine the model to differentiate between caregiver restraint and device restraint, as these contexts may differ in the perceptual, social, and cognitive input they afford.
+Nonetheless, quantifying infants’ general restrained experiences already represents a significant advance.
+This system provides a powerful tool for capturing infants’ physical contexts and how the contexts shape their daily learning opportunities.
 

### 2025-10-12 manuscript_restrained_classification validation.qmd
updated tables and figure 1

@@
 ```{r tbl-indexbetweenmodels}
-#| apa-note: 
-#| tbl-cap: Performance metrics for each window length and ANOVA comparison across window lengths
+#| tbl-cap: Performance metrics according to window lengths
 #| ft.align: left
+#| apa-note: F, df, and p value indicate one-way ANOVA on each metric comparing across the three window lengths
 
@@
 ```{r tbl-cibtwctg}
-#| apa-note: Differences were calculated as restrained – unrestrained metric value.
-M = mean difference; CI = bootstrapped confidence interval.
-All window lengths had lower sensitivity and higher PPV for restrained period detection.
-#| tbl-cap: Differences in sensitivity and positive predictive value (PPV) between restrained and unrestrained prediction across window lengths
-#| ft.align: left
 # For per-class metrics (i.e., sens, ppv, f1), calculate the CI of the metric difference btw classes (i.e., r and u)
-
+# Don't need an APA table for this, but need the table to call values from
 # Function to bootstrap CI of gap for each model and each metric
@@
          "95%CI upper bound"="ci_high")
-
-table_flex_btw_ctg <- table_btw_ctg %>% 
-  flextable() %>% flextable::theme_apa() %>% 
-  line_spacing(part = "all") %>%
-  merge_v(j = "Window Length") %>% fix_border_issues() %>% 
-  add_header_row(values = c("","", "Difference (rest.- unrest.)"), colwidths = c(1,1,3)) %>% 
-  hline(i = c(2,4,6), part = "body") %>%
-  padding(padding.top = 2, padding.bottom = 2) %>% 
-  set_table_properties(layout = "autofit", width = .99) %>% 
-  align(align = "center", part = "all")
-table_flex_btw_ctg
+# 
+# table_flex_btw_ctg <- table_btw_ctg %>% 
+#   flextable() %>% flextable::theme_apa() %>% 
+#   line_spacing(part = "all") %>%
+#   merge_v(j = "Window Length") %>% fix_border_issues() %>% 
+#   add_header_row(values = c("","", "Difference (rest.- unrest.)"), colwidths = c(1,1,3)) %>% 
+#   hline(i = c(2,4,6), part = "body") %>%
+#   padding(padding.top = 2, padding.bottom = 2) %>% 
+#   set_table_properties(layout = "autofit", width = .99) %>% 
+#   align(align = "center", part = "all")
+# table_flex_btw_ctg
 ```
@@
 ```{r fig-cibtwctg}
-#| fig-cap: Differences in sensitivity and (PPV) between restrained and unrestrained prediction across window lengths
-#| fig-width: 8
+#| fig-cap: Differences in sensitivity and PPV between restrained and unrestrained prediction according to window lengths
+#| fig-width: 11
 #| fig-height: 5
-#| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction.
-The dot on each line denotes the mean of the difference.
-All window lengths had higher sensitivity for unrestrained period prediction and higher ppv for restrained period prediction.
+#| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction. The dot on each line denotes the mean of the difference.
 
-p_ci <- ggplot(table_btw_ctg, 
-       aes(x = Metric, 
+p_ci_sen <- ggplot (table_btw_ctg %>% 
+                      filter(Metric=="Sensitivity"),
+                    aes(x=`Window Length`, 
                         y = `M`,
@@
                         ymax = `95%CI upper bound`, 
-           color = `Window Length`)) +
+                        color = `Window Length`))+
   geom_pointrange(position = position_dodge(width = 0.5), size = 0.7) +
-  geom_segment(aes(x = 0, xend = 3, y = 0, yend = 0),
-             linetype = "dashed", color = "gray")+
+  geom_segment(aes(x = 0, xend = 3.5, y = 0, yend = 0),
+             linetype = "dashed", color = "black")+
   scale_color_manual(values=model_colors,labels=c("4 s","16 s", "30 s")) +
-  scale_x_discrete(expand = expansion(add = c(0,2)))+
   scale_y_continuous(
-    limits = c(-0.15, 0.1),  
-    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1) 
+    limits = c(-0.15, 0.05),
+    breaks = c(-0.15,-0.1, -0.05, 0, 0.05)
   )+
   theme_update() +
-  theme(legend.position = c(0.89,0.1))+ 
-  labs(x = "",y = "Metric value difference (rest.- unrest.)",
-    color = "Window Length")+
-  annotate("text", x = 3.5, y = -0.05, label="Unrestrained prediction \n has higher \n metric's value",color = "black")+
-  annotate("text", x = 3.5, y = 0.05, label="Restrained prediction \n has higher \n metric's value",color = "black")
+  labs(x = "",y = "Metric value of rest. - unrest.")+
+  guides(color="none")+
+  ggtitle("A. Sensitivity value difference")+
+  annotate("text", x = 2.6, y = -0.13, label="The model is more \n sensitive in detecting \n unrestrained bouts",color = "black")
+p_ci_sen
 
-p_ci
-# ggsave("figure/CI.png",width=7, height=5)
-# ggsave("figure/CI.eps",width=7, height=5)
+p_ci_ppv <-ggplot (table_btw_ctg %>% 
+                      filter(Metric=="PPV"),
+                    aes(x=`Window Length`, 
+                        y = `M`,
+                        ymin = `95%CI lower bound`, 
+                        ymax = `95%CI upper bound`, 
+                        color = `Window Length`))+
+  geom_pointrange(position = position_dodge(width = 0.5), size = 0.7) +
+  geom_segment(aes(x = 0, xend = 3.5, y = 0, yend = 0),
+             linetype = "dashed", color = "black")+
+  scale_color_manual(values=model_colors,labels=c("4 s","16 s", "30 s")) +
+  scale_y_continuous(
+    limits = c(-0.05, 0.1),
+    breaks = c(-0.05, 0, 0.05, 0.1)
+  )+
+  theme_update() +
+  labs(x = "",y = "Metric value of rest. - unrest.")+
+  guides(color="none")+
+  ggtitle("B. PPV value difference")+
+  annotate("text", x = 2.6, y = 0.085, label="The model's prediction \n of restrained bouts \n is more reliable",color = "black")
+p_ci_ppv
+
+p_ci_sen | p_ci_ppv
+# p_ci <- ggplot(table_btw_ctg, 
+#        aes(x = Metric, 
+#            y = `M`, 
+#            ymin = `95%CI lower bound`, 
+#            ymax = `95%CI upper bound`, 
+#            color = `Window Length`)) +
+#   geom_pointrange(position = position_dodge(width = 0.5), size = 0.7) +
+#   geom_segment(aes(x = 0, xend = 3, y = 0, yend = 0),
+#              linetype = "dashed", color = "gray")+
+#   scale_color_manual(values=model_colors,labels=c("4 s","16 s", "30 s")) +
+#   scale_x_discrete(expand = expansion(add = c(0,2)))+
+#   scale_y_continuous(
+#     limits = c(-0.15, 0.1),  
+#     breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1) 
+#   )+
+#   theme_update() +
+#   theme(legend.position = c(0.89,0.1))+ 
+#   labs(x = "",y = "Metric value difference (rest.- unrest.)",
+#     color = "Window Length")+
+#   annotate("text", x = 3.5, y = -0.05, label="Unrestrained prediction \n has higher \n metric's value",color = "black")+
+#   annotate("text", x = 3.5, y = 0.05, label="Restrained prediction \n has higher \n metric's value",color = "black")
 
@@
          ci_high=ci_high*100) %>% 
+  mutate(ci = paste0("[",round(ci_low, 2),", ", round(ci_high, 2), "]" )) %>% 
+  select(-ci_low, -ci_high) %>% 
   rename("Window Length" = "model",
-         "Human Annotation Prevalence" = "actual_prev",
-         "Model Prediction Prevalence" = "prev",
-         "Correlation" = "cor",
+         "Human Annotated" = "actual_prev",
+         "Model Predicted" = "prev",
+         "Correlation (r)" = "cor",
          "M" = "mean_gap",
-         "95%CI lower bound" = "ci_low",
-         "95%CI upper bound" = "ci_high")
+         "95%CI" = "ci")
 
@@
   fix_border_issues() %>% 
-  add_header_row(values = c("","Prevalence (%)", "Prediction Error (%)"), colwidths = c(1,3,3)) %>% 
+  add_header_row(values = c("","Prevalence (%)", "","Prediction Error (%)"), colwidths = c(1,2,1,2)) %>% 
   hline(i = c(1,2,3)) %>% 
@@
 ```{r fig-prevalencegap}
-#| fig-cap: Prediction error between model prediction and human annotation
-#| fig-width: 6
+#| fig-cap: XX
+#| fig-width:11
 #| fig-height: 5
-#| apa-note: Prediction error is the difference in the overall restrained time between the model estimate and the human annotation.
-A negative prediction error indicates the model underestimate restrained time overall compared to human annotation.
-Each point denotes one session.
-The violin shape denotes the distribution of prediction error of each window length.
-The box denotes the first (Q1) quartile, the median, and the third (Q3) quartile of prediction error.
+#| apa-note: XX
+
+p_prev_cor <- ggplot(data=)
 
@@
 ```{r}
+#| eval: false
+#| include: false
 # check if the data set is imbalanced 
@@
 ```{r}
+#| eval: false
+#| include: false
 # check accuracy outliers 

### 2025-10-22 manuscript_restrained_classification validation.qmd
updated figures and tables

@@
 
-```{r tbl-cibtwctg}
+```{r}
 # For per-class metrics (i.e., sens, ppv, f1), calculate the CI of the metric difference btw classes (i.e., r and u)
@@
 #| fig-width: 11
-#| fig-height: 5
+#| fig-height: 7
 #| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction. The dot on each line denotes the mean of the difference.
@@
   scale_y_continuous(
-    limits = c(-0.15, 0.05),
-    breaks = c(-0.15,-0.1, -0.05, 0, 0.05)
+    limits = c(-0.15, 0.1),
+    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1, 0.1)
   )+
   theme_update() +
-  labs(x = "",y = "Metric value of rest. - unrest.")+
+  labs(x = "",y = "Metric value difference (rest. - unrest.)")+
   guides(color="none")+
@@
   annotate("text", x = 2.6, y = -0.13, label="The model is more \n sensitive in detecting \n unrestrained bouts",color = "black")
-p_ci_sen
 
@@
   scale_y_continuous(
-    limits = c(-0.05, 0.1),
-    breaks = c(-0.05, 0, 0.05, 0.1)
+    limits = c(-0.15, 0.1),
+    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1, 0.1)
   )+
   theme_update() +
-  labs(x = "",y = "Metric value of rest. - unrest.")+
+  labs(x = "",y = "")+
   guides(color="none")+
   ggtitle("B. PPV value difference")+
-  annotate("text", x = 2.6, y = 0.085, label="The model's prediction \n of restrained bouts \n is more reliable",color = "black")
-p_ci_ppv
+  annotate("text", x = 2.6, y = 0.085, label="The model's prediction \n of restrained bouts \n is more credible",color = "black")
 
-p_ci_sen | p_ci_ppv
-# p_ci <- ggplot(table_btw_ctg, 
-#        aes(x = Metric, 
-#            y = `M`, 
-#            ymin = `95%CI lower bound`, 
-#            ymax = `95%CI upper bound`, 
-#            color = `Window Length`)) +
-#   geom_pointrange(position = position_dodge(width = 0.5), size = 0.7) +
-#   geom_segment(aes(x = 0, xend = 3, y = 0, yend = 0),
-#              linetype = "dashed", color = "gray")+
-#   scale_color_manual(values=model_colors,labels=c("4 s","16 s", "30 s")) +
-#   scale_x_discrete(expand = expansion(add = c(0,2)))+
-#   scale_y_continuous(
-#     limits = c(-0.15, 0.1),  
-#     breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1) 
-#   )+
-#   theme_update() +
-#   theme(legend.position = c(0.89,0.1))+ 
-#   labs(x = "",y = "Metric value difference (rest.- unrest.)",
-#     color = "Window Length")+
-#   annotate("text", x = 3.5, y = -0.05, label="Unrestrained prediction \n has higher \n metric's value",color = "black")+
-#   annotate("text", x = 3.5, y = 0.05, label="Restrained prediction \n has higher \n metric's value",color = "black")
-
+(p_ci_sen + p_ci_ppv)
 ```
@@
 #| apa-note: Prevalence was calculated as the proportion of restrained periods among all window samples for each session.
-Prediction error was calculated as model prediction prevalence - human annotation prevalence.
-A negative prediction error means an underestimate from model prediction.
+Prediction error was calculated as model predicted prevalence minus human annotated prevalence.
+A negative prediction error means model prediction underestimates the restrained time compared to human annotation.
 M = mean difference; CI = bootstrapped confidence interval.
@@
   fix_border_issues() %>% 
-  add_header_row(values = c("","Prevalence (%)", "","Prediction Error (%)"), colwidths = c(1,2,1,2)) %>% 
+  add_header_row(values = c("","Mean of Prevalence (%)", "","Prediction Error (%)"), colwidths = c(1,2,1,2)) %>% 
   hline(i = c(1,2,3)) %>% 
@@
 ```{r fig-prevalencegap}
-#| fig-cap: XX
-#| fig-width:11
+#| fig-cap: Restraint prevalence 
+#| fig-width: 11
 #| fig-height: 5
-#| apa-note: XX
+#| apa-note: Prevalence is defined as the proportion of windows predicted (or annotated) as restrained among all windows.
+(A) shows the correlation of prevalence between model prediction and human annotation.
+The black dashed line denotes the reference of y=x.
+Each dot is a session for a specific window length.
+(B) shows the difference of prevalence between human annotation and model prediction (i.e., prediction error).
+A negative prediction error indicates the model estimated less amount of time infants were restrained compared to human annotation.
+The line in the box denotes the median; the boarder lines of the box denotes the first (Q1) quartile and the third (Q3) quartile; and the lines extended denotes the min and max data points within the range of 1.5 IQR (i.e., Q3-Q1).
 
-p_prev_cor <- ggplot(data=)
+p_prev_cor <- ggplot(data=matrix_comb %>% filter(positive_class=="r") %>%
+                   mutate(model=recode(model,
+                                       "4_no_pos" = "4 s",
+                                       "16_no_pos" = "16 s",
+                                       "30_no_pos" = "30 s")),
+                 aes(x=prevalence, y=detection_prevalence,
+                     color=model))+
+  geom_point(shape = 16,size = 5, alpha=.4) + 
+  geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "black") +
+  geom_smooth(method = "lm", se = FALSE) +  
+  geom_text(data = table_prev,
+            aes(x = c(1.15,1.15,1.15),
+                y = c(.7,.8,.9),
+                label = paste0("r = ", round(cor,2)),
+                color = model), 
+            show.legend = FALSE) +
+  scale_color_manual(values=model_colors, labels=c("4 s", "16 s","30 s")) +
+  scale_x_continuous(limits = c(0,1.3), breaks=c(0,.5,1)) +
+  scale_y_continuous(limits = c(0,1), breaks=c(0,.5,1))+
+  labs(x = "Human annotated prevalence", y = "Model predicted prevalence", color="Window Length")+
+  theme_update()+
+  theme(legend.position = c(0.8, 0.2))+
+  ggtitle("A. Prevalence Correlation")
 
@@
                                        "30_no_pos" = "30 s")),
-                 aes(x=model, y=detection_prevalence-prevalence, color=model))+
-  geom_violin(trim = FALSE, alpha = 0.4) +
-  geom_quasirandom(aes(color = model),
-                   dodge.width = 0.7, varwidth = TRUE, alpha = 0.6, size = 2) +
+                 aes(x=model, y=(detection_prevalence-prevalence)*100, color=model))+
+  # geom_violin(trim = FALSE, alpha = 0.4) +
+  # geom_quasirandom(aes(color = model),
+  #                  dodge.width = 0.7, varwidth = TRUE, alpha = 0.6, size = 2) +
   geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white") + 
@@
   scale_color_manual(values=model_colors, labels=c("4 s", "16 s","30 s")) +
-  labs(x="Window Length", y="Prediction Error (prediction-human annotation)")+
+  scale_y_continuous(
+    limits = c(-20,20),
+    breaks = c(-20, -10, 0, 10, 20)
+  )+
+  labs(x="Window Length", y="Prediction Error (%)")+
   guides(color = "none")+
-  theme_update()
-p_prev_gap
+  theme_update()+
+  ggtitle("B. Prevalence Offset")
+
+(p_prev_cor+p_prev_gap)
 ```
@@
 Despite the overall good performance, all window lengths presented prediction bias towards unrestrained class.
-@tbl-cibtwctg and @fig-cibtwctg show that all window lengths had lower sensitivity and higher positive predictive value (PPV) for restrained compared to unrestrained periods, indicated by the negative difference of sensitivity and positive difference of PPV.
+@fig-cibtwctg show that all window lengths had lower sensitivity and higher positive predictive value (PPV) for restrained compared to unrestrained periods, indicated by the negative difference of sensitivity and positive difference of PPV.
 The sensitivity difference for 4 s window was `r apa_num(table_btw_ctg[1,3])`, 95% bootstrapped CI [`r apa_num(table_btw_ctg[1,4])`, `r apa_num(table_btw_ctg[1,5])`], for 16 s window was `r apa_num(table_btw_ctg[3,3])`, 95% CI [`r apa_num(table_btw_ctg[3,4])`, `r apa_num(table_btw_ctg[3,5])`], and for 30 s window was `r apa_num(table_btw_ctg[5,3])`, 95% CI [`r apa_num(table_btw_ctg[5,4])`, `r apa_num(table_btw_ctg[5,5])`].
@@
 #| fig-height: 3
-#| apa-note: Each point denotes one session's restrained time proportion processed by the 30 s model.
-The box denotes the first (Q1) quartile, the median, and the third (Q3) quartile of prediction error.
+#| apa-note: Each point denotes one session. The line denotes the group mean of restrained time (%)
 
@@
 
-ds_sum_excl <- ds_sum_excl %>% 
-  mutate(age_group_num=ifelse(age_group=="Younger",5.5, 12.5))
-
-p_age <- ggplot(ds_sum_excl, aes(x=agemo, y=restrained_prop)) +
-  geom_point(shape = 1,size = 3) + 
-  geom_boxplot(
-    aes(x = age_group_num, y = restrained_prop, group = age_group_num),
-    inherit.aes = FALSE,
-    width = 3,
-    alpha = 0.2,
-    outlier.shape = NA,
-    color = "blue"
-  ) +
-  scale_x_continuous(name = "Age Group", breaks = c(5.5, 12.5), labels=c("Young", "Old"), limits = c(3,15)) +
+p_age <- ggplot(ds_sum_excl, aes(x=age_group, y=restrained_prop)) +
+  geom_jitter(width=.2, shape = 1,size = 3, color="black") + 
+  stat_summary(fun = mean, geom = "crossbar",
+               width = 0.4, color = "red") +
   scale_y_continuous(name = "Restrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
@@
 p_age
-# ggsave("figure/age_related_change.png",width=6, height=4)
-# ggsave("figure/age_related_change.eps",width=6, height=4)
 

### 2025-10-24 manuscript_restrained_classification validation.qmd
updated figures and tables

@@
 
-theme_update(text = element_text(size = 11),
-             axis.text.x = element_text(size = 11, color = "black"),
-             axis.title.x = element_text(size = 12),
-             axis.text.y = element_text(size = 11,  color = "black"),
-             axis.title.y = element_text(size = 12),
+theme_update(text = element_text(size = 12),
+             axis.text.x = element_text(size = 12, color = "black"), 
+             axis.title.x = element_text(size = 14),
+             axis.text.y = element_text(size = 12,  color = "black"), 
+             axis.title.y = element_text(size = 14), 
              panel.background = element_blank(),panel.border = element_blank(), 
@@
              panel.grid.minor = element_blank(), axis.line = element_blank(), 
+             axis.ticks.length=unit(.25, "cm"), 
              legend.key = element_rect(fill = "white"))
@@
 ```{r tbl-indexbetweenmodels}
-#| tbl-cap: Performance metrics according to window lengths
+#| tbl-cap: Performance metrics according to window length
 #| ft.align: left
@@
 #| fig-cap: Differences in sensitivity and PPV between restrained and unrestrained prediction according to window lengths
-#| fig-width: 11
-#| fig-height: 7
-#| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction. The dot on each line denotes the mean of the difference.
+#| fig-width: 8
+#| fig-height: 5
+#| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction.
+The dot on each line denotes the mean of the difference (restrained - unrestrained).
 
@@
   geom_pointrange(position = position_dodge(width = 0.5), size = 0.7) +
-  geom_segment(aes(x = 0, xend = 3.5, y = 0, yend = 0),
+  geom_segment(aes(x = 0.5, xend = 4, y = 0, yend = 0),
              linetype = "dashed", color = "black")+
@@
   scale_y_continuous(
-    limits = c(-0.15, 0.1),
-    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1, 0.1)
+    limits = c(-0.15, 0.15),
+    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1, 0.15)
   )+
-  theme_update() +
   labs(x = "",y = "Metric value difference (rest. - unrest.)")+
@@
   ggtitle("A. Sensitivity value difference")+
-  annotate("text", x = 2.6, y = -0.13, label="The model is more \n sensitive in detecting \n unrestrained bouts",color = "black")
+  annotate("text", x = 3.8, y = -0.12, label="More sensitive \n to unrestrained",color = "black", hjust=1, size = 4.5)+
+  annotate("segment",
+         x = 3.9, xend = 3.9,  # arrow start → end
+         y = -0.1,  yend = -0.14,
+         arrow = arrow(length = unit(0.2, "cm")),
+         size = 1, color = "black")+
+  annotate("text", x = 3.8, y = 0.12, label="More sensitive \n to restrained",color = "black", hjust=1, size = 4.5)+
+  annotate("segment",
+           x = 3.9, xend = 3.9,  # arrow start → end
+           y = 0.1,  yend = 0.14,
+           arrow = arrow(length = unit(0.2, "cm")),
+           size = 1, color = "black")+
+  theme_update()
 
@@
   scale_y_continuous(
-    limits = c(-0.15, 0.1),
-    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1, 0.1)
+    limits = c(-0.15, 0.15),
+    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1, 0.15)
   )+
-  theme_update() +
   labs(x = "",y = "")+
@@
   ggtitle("B. PPV value difference")+
-  annotate("text", x = 2.6, y = 0.085, label="The model's prediction \n of restrained bouts \n is more credible",color = "black")
+  annotate("text", x = 3.8, y = -0.12, label="More credible \n unrestrained predictions",color = "black", hjust=1, size = 4.5)+
+  annotate("segment",
+         x = 3.9, xend = 3.9,  # arrow start → end
+         y = -0.1,  yend = -0.14,
+         arrow = arrow(length = unit(0.2, "cm")),
+         size = 1, color = "black")+
+  annotate("text", x = 3.8, y = 0.12, label="More credible \n restrained predictions",color = "black", hjust=1, size = 4.5)+
+  annotate("segment",
+           x = 3.9, xend = 3.9,  # arrow start → end
+           y = 0.1,  yend = 0.14,
+           arrow = arrow(length = unit(0.2, "cm")),
+           size = 1, color = "black")+
+  theme_update()
 
@@
 
-```{r tbl-prevalencecor}
-#| apa-note: Prevalence was calculated as the proportion of restrained periods among all window samples for each session.
-Prediction error was calculated as model predicted prevalence minus human annotated prevalence.
-A negative prediction error means model prediction underestimates the restrained time compared to human annotation.
-M = mean difference; CI = bootstrapped confidence interval.
-#| tbl-cap: Model predicted and human annotated prevalence of restrained periods, their correlation, and differences
-#| ft.align: left
-
+```{r}
+# Table. Model predicted and human annotated prevalence of restrained periods, their correlation, and differences
 df_prev <- matrix_comb %>%
@@
 
-table_prev_100 <- table_prev %>% 
-  mutate(actual_prev=actual_prev*100,
-         prev = prev*100, 
-         mean_gap = mean_gap *100,
-         ci_low=ci_low * 100,
-         ci_high=ci_high*100) %>% 
-  mutate(ci = paste0("[",round(ci_low, 2),", ", round(ci_high, 2), "]" )) %>% 
-  select(-ci_low, -ci_high) %>% 
-  rename("Window Length" = "model",
-         "Human Annotated" = "actual_prev",
-         "Model Predicted" = "prev",
-         "Correlation (r)" = "cor",
-         "M" = "mean_gap",
-         "95%CI" = "ci")
-
-table_flex_prev <- table_prev_100 %>% 
-  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
-  fix_border_issues() %>% 
-  add_header_row(values = c("","Mean of Prevalence (%)", "","Prediction Error (%)"), colwidths = c(1,2,1,2)) %>% 
-  hline(i = c(1,2,3)) %>% 
-  padding(padding.top = 2, padding.bottom = 2) %>% 
-  set_table_properties(layout = "autofit", width = .99) %>% 
-  align(align = "center", part = "all")
-table_flex_prev
+# table_prev_100 <- table_prev %>% 
+#   mutate(actual_prev=actual_prev*100,
+#          prev = prev*100, 
+#          mean_gap = mean_gap *100,
+#          ci_low=ci_low * 100,
+#          ci_high=ci_high*100) %>% 
+#   mutate(ci = paste0("[",round(ci_low, 2),", ", round(ci_high, 2), "]" )) %>% 
+#   select(-ci_low, -ci_high, -actual_prev, -prev) %>% 
+#   rename("Window Length" = "model",
+#          "Correlation (r)" = "cor",
+#          "M" = "mean_gap",
+#          "95%CI" = "ci")
+# 
+# table_flex_prev <- table_prev_100 %>% 
+#   flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
+#   fix_border_issues() %>% 
+#   add_header_row(values = c("","Prediction Error (%)"), colwidths = c(2,2)) %>% 
+#   hline(i = c(1,2,3)) %>% 
+#   padding(padding.top = 2, padding.bottom = 2) %>% 
+#   set_table_properties(layout = "autofit", width = .99) %>% 
+#   align(align = "center", part = "all")
+# table_flex_prev
 ```
@@
 #| fig-cap: Restraint prevalence 
-#| fig-width: 11
-#| fig-height: 5
+#| fig-width: 8
+#| fig-height: 4
 #| apa-note: Prevalence is defined as the proportion of windows predicted (or annotated) as restrained among all windows.
 (A) shows the correlation of prevalence between model prediction and human annotation.
-The black dashed line denotes the reference of y=x.
-Each dot is a session for a specific window length.
+The black dashed identity line indicates perfect agreement.
+Each circular symbol represents one session for a specific window length (denoted by color).
 (B) shows the difference of prevalence between human annotation and model prediction (i.e., prediction error).
 A negative prediction error indicates the model estimated less amount of time infants were restrained compared to human annotation.
-The line in the box denotes the median; the boarder lines of the box denotes the first (Q1) quartile and the third (Q3) quartile; and the lines extended denotes the min and max data points within the range of 1.5 IQR (i.e., Q3-Q1).
+The horizontal line in the box denotes the median; the boarder lines of the box denotes the first (Q1) quartile and the third (Q3) quartile; and the lines extended denotes the min and max data points within the range of 1.5 IQR (i.e., Q3-Q1).
 
@@
                      color=model))+
-  geom_point(shape = 16,size = 5, alpha=.4) + 
+  geom_point(shape = 16,size = 3, alpha=.3) + 
   geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "black") +
@@
   geom_text(data = table_prev,
-            aes(x = c(1.15,1.15,1.15),
+            aes(x = c(1.16,1.16,1.16),
                 y = c(.7,.8,.9),
                 label = paste0("r = ", round(cor,2)),
-                color = model), 
+                color = model, size = 12), 
             show.legend = FALSE) +
   scale_color_manual(values=model_colors, labels=c("4 s", "16 s","30 s")) +
-  scale_x_continuous(limits = c(0,1.3), breaks=c(0,.5,1)) +
+  scale_x_continuous(limits = c(0,1.2), breaks=c(0,.5,1)) +
   scale_y_continuous(limits = c(0,1), breaks=c(0,.5,1))+
-  labs(x = "Human annotated prevalence", y = "Model predicted prevalence", color="Window Length")+
+  coord_cartesian(clip = "off")+
+  labs(x = "Human annotated prevalence", y = "Model predicted prevalence")+
   theme_update()+
-  theme(legend.position = c(0.8, 0.2))+
+  theme(axis.title.x = element_text(hjust = 0.37))+
+  guides(color = "none")+
   ggtitle("A. Prevalence Correlation")
@@
                  aes(x=model, y=(detection_prevalence-prevalence)*100, color=model))+
-  # geom_violin(trim = FALSE, alpha = 0.4) +
-  # geom_quasirandom(aes(color = model),
-  #                  dodge.width = 0.7, varwidth = TRUE, alpha = 0.6, size = 2) +
   geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white") + 
@@
   )+
-  labs(x="Window Length", y="Prediction Error (%)")+
-  guides(color = "none")+
+  labs(x="Window Length", y="Prediction Error (%)", color="Window Length")+
   theme_update()+
+  theme(legend.position = c(0.5,0.9),legend.direction = "horizontal")+
   ggtitle("B. Prevalence Offset")
@@
 (p_prev_cor+p_prev_gap)
+
 ```
@@
 ```{r fig-agechange}
-#| fig-cap: Infants' full day restrained time proportion with age
-#| fig-width: 4
-#| fig-height: 3
-#| apa-note: Each point denotes one session. The line denotes the group mean of restrained time (%)
+#| fig-cap: Younger versus older infants’ full day restrained time (as a proportion of recording time) 
+#| fig-width: 5
+#| fig-height: 4
+#| apa-note: Each point denotes one session. The horizontal red line denotes the group mean of restrained time proportion.
 
@@
                width = 0.4, color = "red") +
-  scale_y_continuous(name = "Restrained time (%)", breaks = seq(0, 1, .25), limits = c(0,1))+
+  scale_y_continuous(name = "Restrained time proportion of awake time", breaks = seq(0, 1, .25), limits = c(0,1))+
   theme_update()+

### 2025-11-09 manuscript_restrained_classification validation.qmd
Add discussion to the manuscript
Add discussion to the manuscript

@@
 blank-lines-above-title: 2
-shorttitle: "Quantify infant restrained time with sensors"
+shorttitle: "Quantify infant restraint with sensors"
 fig-dpi: 300
@@
   - name: Hailey N. Rousey
-    orcid: XX
-    email: XX
+    orcid: 0009-0000-9828-0693
+    email: hailey.rousey@email.ucr.edu
     affiliations:
@@
     authorship-agreements: null
-abstract: "XXX"
+abstract: "Physical restraint—including being held, carried, and restrained in devices—is a common feature of infants’ everyday lives.
+This study developed and validated a machine learning model to detect infants’ restraint using wearable sensors movement data derived from an existing data set (Franchak, 2023).
+The model achieved high accuracy (89%) and substantial kappa agreement (0.73) compared with human-coded ground truth.
+The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when using a longer data-aggregating window length.
+The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infant restraint time.
+In short, the current study demonstrated the feasibility of using wearable sensors to quantify infants’ real-world restraint experiences, offering a new tool for studying how daily restraint influences early development."
 keywords: [infant, restraint, everyday experiences, wearable sensors, machine learning]
@@
 load("whole-model-prediction.RData")
-rm(ds, ds_sum_excl, ds_sum_excl0, id_demo, id_info)
+
 ds_sum <- ds_sum %>% 
@@
   filter(unique_id != "185/3")
-demo <- read.csv("lena-imu-compiled.csv")
-demo[demo$id=="180", "infant_sex"] <- "male"
+id_info <- id_info %>% 
+  filter(unique_id != "185/3")
+id_demo <- id_demo %>% 
+  filter(unique_id != "185/3")
 
-temp <- ds_sum
-temp <- temp %>% 
-  separate(unique_id, into=c("id","sess"),sep="/",convert = TRUE) %>% 
-  distinct(id, .keep_all = TRUE)
-
-temp <- left_join(temp, demo %>% select("id","infant_sex","infant_race","infant_ethnicity") %>% distinct(id, .keep_all = TRUE), by="id", unmatched="drop")
-
-age<-temp %>% 
+age <- id_demo %>% 
   group_by(age_group) %>% 
@@
 
-sex<-temp %>% 
+sex<-id_demo %>% 
   group_by(infant_sex) %>% 
@@
             .groups="drop")
-race<-temp %>% 
+
+race<-id_demo %>% 
   group_by(infant_race) %>% 
@@
 
-[NEED TO BE INCORPORATED WITH WORD DOC]
+Physical restraint is a common feature in infants’ everyday lives.
+Infants are routinely held or carried by caregivers, and infants are often restrained in devices such as highchairs, walkers, and strollers [@Birken2015s].
+As an externally regulated aspect of infants’ daily life, restraint shapes infants’ experiences based on caregivers’ decisions.
+Restraint limits infants’ physical activities and therefore hinders infants’ motor development [e.g., @Pin2007z].
+Restraint also frames infants’ accessible physical environment by altering experiences that are relevant for perceptual and cognitive development.
+For example, restraint is associated with fewer object holding experiences [@Franchak2024l], reduced amount and quality of language input [@Malachowski2023a], and increased opportunities for joint attention [@Karasik2025h].
+Therefore, characterizing infants’ restrained experiences helps profile infants’ daily learning contexts and situate their motor, perceptual, and cognitive development within everyday experiences.
 
-Physical restraint is a common feature in infants’ everyday life.
-Infants are routinely held or carried by caregivers, or restrained in devices like highchairs, walkers, and strollers for the sake of comforting, safety, transportation, etc [@Birken2015s].
-Restraint influences infants’ motor development by constraining their body movement [e.g., @Pin2007z].
-Restraint also frames infants’ learning experiences for perceptual and cognitive development.
-For example, the frequency of object holding experiences differs between restrained and unrestrained times [@Franchak2024l], the amount and quality of language infants receive decrease as infants spend more time restrained in devices [@Malachowski2023a], yet the frequency of joint attention and
+Previous studies have used a variety of methods to document infants’ restrained time in the home.
+Some used parent surveys, including questionnaires, diaries, or ecological momentary assessment [e.g., @Abbott2001u; @Karasik2018t; @Franchak2019o].
+Some used video recording and human annotation [e.g., @Springfield2025p; @Wang2025v].
+Although surveys provide an overall profile of infants’ daily life, and video recordings provide continuous time series data, no method so far can simultaneously provide full-day and continuous data to account for infants’ restraint experiences.
+In the current study, we propose that wearable sensors can be a solution for recording full-day, continuous movement behavior, and machine learning algorithms can be an accurate and efficient way to automatically classify wearable sensor data.
 
-Holding and carrying restraint practices may facilitate infants' social cognitive development [e.g., @Berecz2020y].
-Restraint, as an externally regulated aspect of infants’ daily life, decides the scope of input and experiences infants have access to.
-Therefore, characterizing infants restrained experiences helps profile infants’ daily learning contexts and situate their motor, perceptual, and cognitive development within everyday experiences.
+Moreover, previous studies did not define restraint consistently.
+Some definitions only focused on caregiver restraint such as holding and carrying [e.g., @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q], some only focused on seating devices [e.g., @Callahan1997s], some looked at different surfaces (e.g., lap, cradle, and seat) infants are on [e.g., @Graciosa2024o], some focused broadly on containment equipment [e.g., @Bartlett2003w; @Carson2022n; @Hesketh2015i; @Karasik2022a], some combined holding and carrying with seating equipment [@Springfield2025p], and others covered a wider range of restraint like furniture placement, seating devices and holding [@Franchak2024l; @Malachowski2023a; @Siddicky2021c].
+The inconsistency in restraint definitions makes it difficult to compare results across studies.
+In the current study, we take a broader scope and consider any practices that limit infants’ freedom of movement as restraint.
+We do not pre-define a list of restraint practices but focus on infants’ real-time opportunities for movement by classifying restraint whenever infants cannot change body position or locomote freely.
+Quantifying both caregiver restraint (i.e., holding and carrying) and device restraint, we can get a whole picture of the external constraints that infants experience in daily life.
 
-Previous studies made several attempts to document infant restrained time at home.
-Some used parental surveys, including questionnaires, diaries, or ecological momentary assessment [e.g., @Abbott2001u; @Karasik2018t; @Franchak2019o].
-Some used video recording and human annotation [e.g., @Wang2025v].
-However, these methods cannot provide a continuous and prolonged account of infants’ experiences.
-In the current study, we propose that wearable sensors can be a solution for recording continuous, full-day movement data.
-Moreover, previous studies had different emphases on restraint practices.
-Some only focused on caregiver restraint like holding and carrying [e.g., @Airaksinen2024z], some only focused on seating devices use [e.g., @Callahan1997s], some focused broadly on containment equipment [e.g., @Bartlett2003w; @Carson2022n; @Hesketh2015i; @Karasik2022a], and others covered a wider range of restraint like furniture placement, seating devices and holding [e.g., @Siddicky2020f; @Franchak2024l; @Malachowski2023a].
-In the current study, we combine different aspects of restraint that have been separately examined into one investigation.
+## Previous Measurements of Restraint
 
-## Previous Measurements of Restrained Experiences
+Despite the inconsistency in definitions, prior work converges to show that caregiver and device restraints are common in infant daily life, and the time infants spend restrained changes over development.
+In terms of caregiver restraint, 2- to 6-month-old infants spent around 40% of their awake time being held or carried [@Malachowski2023a; @Siddicky2020f].
+Older 10- to 13-month-olds spent around 19% of their awake time being held or carried [@Franchak2024l].
+In terms of device restraint, 4- to 6-month-olds spent around 20% of awake time in seating devices, and 35% in placement with physical support [@Callahan1997s; @Malachowski2023a].
+Older 10- to 13-month-olds spent around 27% of awake time in restraining furniture or devices [@Franchak2024l].
+Additionally, a study that compared across ages also confirmed that infants’ overall restrained time decreased from 2 to 6 months of age [@Carson2022n].
 
-Despite different emphases on infant restrained experiences, past research shows that both caregiver restraint and device restraint are common in infant daily life, and the time infants spend restrained changes over development.
-In terms of caregiver restraint, infants aged from 2 to 6 months spent around 40% of their awake time being held or carried [@Malachowski2023a; @Siddicky2020f].
-Older infants aged from 10 to 13 months spent around 19% of their awake time being held or carries [@Franchak2024l].
-In terms of device restraint, infants aged from 4 to 6 months spent around 20% of awake time in seating devices, and 35% in placement with physical support [@Callahan1997s; @Malachowski2023a].
-Older infants aged from 10 to 13 months spent around 27% of awake time in restraining furniture or devices [@Franchak2024l].
+## Restraint Shapes Infant Learning Opportunities
 
-Based on the statistics above, it appears that caregiver restraint and device restraint decrease with age.
-However, other work that compared across ages within a single study presented contradictory results.
-@Airaksinen2024z, using wearable sensors and machine learning modeling, found that infants from Finland experienced increased holding and carrying time from 4 to 18 months of age.
-On the contrary, @Carson2022n, using parental retrospective report, found that infants from Canada experienced decreased overall restrained time from 2 to 6 months of age.
-@Franchak2019o, using ecological momentary assessment (EMA), also found that infants from the US experienced decreased time held by caregivers from 3 to 12 months.
-The author and their colleagues further expanded the findings to 10- to 13-month-olds, showing that infants experienced decreased held time and increased overall unrestrained time with age [@Franchak2024l].
-The discrepancies may be attributed to study design (i.e., longitudinal versus cross-sectional), cultural differences, as well as research methods (i.e., wearable sensors versus parental reports).
-The current study, which uses wearable sensors on U.S.
-infants aged from 4 to 14 months, might help clarify the inconsistencies found in prior research.
+In addition to measuring the total duration of infants’ restrained time, prior research has also examined associations between restraint and developmental outcomes.
+Motor development appears to be the domain most directly associated with restraint.
+Various studies have shown that restraint is related to less proficiency in infant motor skills and temporary delay in milestone achievement [@Abbott2001u; @Bartlett2003w; @Carson2022n; @Karasik2023d; @Pin2007z].
+Physiological studies tracking infants’ body movements (e.g., spine, legs, hips) in real time suggest that reduced movement opportunities may underlie these effects [@Jiang2016z; @Siddicky2020f; @Siddicky2021c].
+Restraint may also influence other aspects of infant development by shaping infants’ learning opportunities in real time.
+Infants produce fewer object interactions when restrained compared to when they are unrestrained [@Franchak2024l].
+Moreover, infants hear less amount of and less consistent adult language when restrained compared to unrestrained [@Malachowski2023a].
 
-## Restrained Practices Shape Infant Learning Opportunities in Real Time
-
-Beyond measuring total duration of infants’ restrained time, prior research also examined how real-time restrained experiences may constrain or support infant learning.
-On the one hand, restraint practices limit infants’ opportunities for learning.
-First, restraint practices prevent infants from moving their body parts (e.g., spine, legs, hip) and thus may cause temporary delay in motor development [@Abbott2001u; @Bartlett2003w; @Carson2022n; @Karasik2023d; @Pin2007z; @Siddicky2020f].
-Second, infants with relatively more restrained time may produce fewer object interactions, with longest caregiver restrained periods related to the least object interaction [@Franchak2024l].
-Third, infants with relatively more device-restrained time may hear less amount of and less consistent adult language [@Malachowski2023a].
-
-On the other hand, restrained periods can enrich certain kinds of learning, especially for younger infants, through stabilizing their body posture.
-Consider young infants who cannot sit up independently yet and thus cannot free up their hands to explore objects [@Soska2010t].
-Restraint devices allow these pre-sitting infants to experience the advantages of a sitting posture by steadying their bodies and enabling eye-hand coordination.
-Moreover, infants in sitting or upright positions have better opportunities to look at faces compared to infants in prone position [@Franchak2018r].
-Restraint devices or carriers that rescue infants from prone position can elevate their viewpoint and afford richer visual input.
-These examples demonstrate that restraint practices might provide infants with learning experiences that they would not otherwise obtain due to their body constraints, especially for younger infants.
-These examples also underscore that real-time observation of restraint is essential for understanding how context shapes infants’ multimodal learning opportunities.
+Restraint is not always detrimental.
+Restraint can facilitate certain kinds of learning through stabilizing infants’ body posture or changing their placement in space.
+For example, restraint devices that support sitting can steady non-sitters’ bodies and facilitate spontaneous object exploration [@Soska2014u].
+Moreover, some restraint devices (e.g., high chairs) or carriers put infants in sitting or upright positions high off the ground, in which infants have richer visual input of faces, distant toys, and elevated portions of their environment compared to when prone on the ground [@Kretch2015p; @Kretch2014r; @Franchak2018r; @Luo2020u].
+Together, these examples underscore that real-time observation of restraint is essential for understanding how restraint shapes infants’ multimodal learning opportunities from moment to moment.
 
@@
 
-Researchers have used survey methods including traditional retrospective report and ecological momentary assessment (EMA) to quantify infants’ restrained time, but both methods have similar limitations.
-Retrospective parental report, which was most commonly used, asks parents to estimate their infant’s average daily restrained time over the past few weeks or month [@Abbott2001u; @Bartlett2003w; @Callahan1997s; @Carson2022n; @Hesketh2015i; @Karasik2018t; @Siddicky2020f].
-Despite its widespread use, this method may not be precise enough because it highly relies on parents’ memory—it is difficult to recall and sum every bout of restraint practices and then report an overall time estimation [@Bradburn1987g].
+Researchers have used survey methods, including traditional retrospective report and ecological momentary assessment (EMA), to quantify infants’ restrained time, but survey methods have limitations.
+Retrospective parent report asks parents to estimate their infant’s average daily restrained time over the past few weeks or month [@Abbott2001u; @Bartlett2003w; @Callahan1997s; @Carson2022n; @Hesketh2015i; @Karasik2018t; @Siddicky2020f].
+Despite the widespread use of retrospective reports, they lack precision by relying on parents’ memory—it is difficult to recall and sum every bout of restraint practices and then report an overall time estimation [@Bradburn1987g].
 There are exceptions where caregivers were asked to complete structured daily diaries reflecting on the previous day’s restraint practices in short intervals (e.g., every five minutes) [@Karasik2018t; @Majnemer2005j].
-However, this method might still be sensitive to memory loss and thus not reliable in providing accurate data.
+However, this method might still be sensitive to memory loss and thus not as accurate in providing data.
+EMA survey techniques reduce caregivers’ memory burden by making momentary phone-survey prompts to caregivers multiple times (e.g., 5, 10, or 12 times) randomly throughout a day [@Franchak2019o; @Franchak2024l; @Malachowski2023a].
+Despite the benefits of gathering immediate observations rather than retrospective reports, EMA cannot provide continuous time-series data.
+Thus, a method is needed that provides time-sequence information without relying on parent report.
 
-The recently developed method EMA reduces caregivers’ memory burden by making momentary phone-survey prompts to caregivers multiple times (e.g., 5, 10, or 12 times) randomly throughout a day [@Franchak2019o; @Franchak2024l; @Malachowski2023a].
-However, same limitations still remain.
-First, both methods assume that the respondent is consistently present to observe the infant throughout the day.
-The report might not be fully representative of full-day activities if the infant happens to be restrained more frequently when the respondent is absent (e.g., the caregiver restrains the infant at home while they goes out on errands).
-Second, neither method provides continuous time-series data.
-They cannot answer granular questions such as: When throughout the day does restraint happen most frequently?
-Are restrained bouts usually prolonged or brief?
-Thus, a method not relying on parental report and providing time-sequence information is needed.
-
-Video recording and human annotation, as another method, overcomes the above drawbacks of survey-based methods.
-However, one of the limitations of this method is being obtrusive.
+Video recording and human annotation overcomes the above drawbacks of survey-based methods.
+However, one of the limitations of video observation is being obtrusive.
 The presence of a researcher in the room may affect the caregiver’s and infant’s behaviors.
-Another limitation is that this approach is labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies.
-Indeed, these studies usually do not have long duration of recording (e.g., one hour in @Wang2025v).
-Within the short duration of recording, researchers would tend to record the period when the infants are awake and active.
-In contrast, the periods when the infants are restrained, like seated in front of TV or in highchairs, would likely be skipped.
-Thus, this method is sensitive to sampling bias and might not be representative of infants’ whole-day experiences.
+Another limitation is that video recording and video annotation are both labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies.
+Indeed, prior work that videotaped and manually annotated infants’ restraint had short durations of recording (e.g., half an hour in @Springfield2025p, and one hour in @Wang2025v).
+Within the short duration of recording, researchers tend to record the period when infants are awake and active.
+In contrast, periods when the infants are restrained, like seated in front of TV or in highchairs, would likely be skipped.
+Thus, video recording is subject to sampling bias and might not be representative of infants’ whole-day experiences.
 
@@
 
-A system that uses wearable sensors (i.e., inertial movement units) to record real-time movement data and machine learning models to classify movement categories can overcome all the limitations listed above.
-First, by collecting infants’ movement data through sensors, the system can objectively capture infants’ position, not relying on parental report.
-Second, the sensors can provide moment-to-moment observation thatd is continuous and rich in time-sequence information.
-Third, the sensors are embedded in baby garments, making the system less obtrusive.
-Fourth, the system can record and process full-day or multiple-day data at a time due to the sensors’ sufficient battery life and automative classification of machine learning models.
+A system that uses wearable sensors (i.e., inertial movement units) to record real-time movement data and machine learning models to classify movement categories can overcome the limitations listed above.
+First, by collecting infants’ movement data through sensors, the system can objectively capture infants’ motion without relying on a respondent or a video annotator’s interpretation.
+Second, the sensors can provide moment-to-moment observation that is continuous and rich in time-sequence information.
+Third, sensors embedded in baby garments are mobile and unobtrusive—an experimenter does not need to follow the infant with a camera to record movements.
+Fourth, the system can record and process full-day or multiple-day data at a time due to the sensors’ sufficient battery life and automated classification of machine learning models.
 
-The past decade has witnessed an uptake of using wearable sensors to record and classify movement categories, but challenges remain.
+The past decade has witnessed an uptick of using wearable sensors to record and classify movement categories to enjoy these advantages.
 The application of wearable sensors progressed from adults [@Arif2015u; @Preece2009n], to children [@Nam2013y; @Ren2016-px; @Stewart2018h], and then to infants [@Airaksinen2020j; @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q].
-For infants, there have been established models that classify infants’ body postures [@Airaksinen2020j; @Franchak2021k; @Franchak2024e], and models that detect holding and carrying periods [@Airaksinen2024z; @Yao2019q].
-Prior works concluded valuable experience, such as using multiple sensors, and including gyroscope data together with accelerator data [@Airaksinen2025l].
+For infants, there have been established models that classify infants’ body postures in general—including holding and carrying [@Airaksinen2020j; @Franchak2021k; @Franchak2024e], and models that specifically detect holding and carrying periods [@Airaksinen2024z; @Yao2019q].
+Insights from prior studies suggest effective practices, such as using multiple sensors and using sensors that record both gyroscope and accelerator data [@Airaksinen2025l; @Franchak2024e].
 
 However, one of the challenges researchers continue to face is the reduced model performance in naturalistic home settings [@Airaksinen2024z; @Franchak2024e; @Yao2019q], compared to more controlled laboratory environments [@Airaksinen2020j; @Franchak2021k].
-Indeed, infants may exhibit more complex and fluid body positions at home versus in lab, which requires the models to be able to generalize across scenarios.
-Moreover, since no prior work has used wearable sensors to detect infants’ overall restrained periods beyond caregiver restraint, the current study faces a unique challenge to detect this new behavior category.
+Indeed, infants may exhibit more complex and unexpected movements at home versus in lab, which requires the models to be able to generalize to scenarios that go beyond the models’ training data set.
+Moreover, since no prior work has used wearable sensors to detect infants’ overall restrained periods, the current study addresses a unique detection problem.
 
-Detecting restrained periods would be even more challenging than classifying body postures or detecting holding periods only.
-First, infants may show similar body movements whether restrained or not.
-For example, infants would kick legs when they are lying without restraint and also when they are strapped in a swing.
-The differences in body movement between restrained and unrestrained periods may be subtle to capture.
+Detecting restrained periods might be more challenging than classifying body postures.
+First, infants may show similar body movements whether restrained or not, making restrained and unrestrained classes more difficult to differentiate.
+For example, infants might kick legs when they are lying without restraint and also when they are strapped in a highchair.
 Possible distinctions might be manifested in reduced range of motion or repetitive motion due to constraints, which requires high-precision sensors and sensitive models to detect.
-Second, infants may present completely different body movements across various restrained scenarios.
-Holding is relatively easy to detect because it produces distinguishable patterns such as infants getting lifted or put down at an unusual speed, and infants moving at an adult’s pace with minimal leg movement.
-However, other kinds of restraint have diverse movement patterns.
-For example, infants seated in highchairs would be in a sitting position with occasional limbs swinging, whereas infants positioned in walkers would be in an upright position with extended legs and slow walking.
-Accordingly, the wearable sensors system must identify the underlying kinematic signatures common to various restraint types while accounting for their contextual variability.
+Second, there are lots of possible restraint forms infants can present—caregiver holding can be holding the infant up straight or holding in arms breastfeeding; seating restraint can happen in countless infant seats, which vary in affordances of moving [@Alghamdi2024d].
+Accordingly, the wearable sensor system should be robust enough to generalize across heterogeneous restraint scenarios.
+Thus, the dual requirement of sensitivity and robustness is the key challenge for developing a reliable model to detect infant restraint in the home.
 
-A promising strategy to address this untested classification lies in the flexibility of modeling approaches.
-One strategy is adjusting the window size within which movement data are aggregated.
-With a fixed sampling rate, larger windows provide more data for the model to leverage to predict the current category.
-Prior studies have used window sizes ranging from 2.3 to 4 seconds for infant posture category—a multiclass classification [@Airaksinen2020j; @Franchak2024e], and from 1.15 to 10 seconds for infant holding and carrying detection—a binary task [@Airaksinen2024z].
-Considering that transitions between restraint states are typically slower than transitions between postures, we set window sizes longer than 4 seconds.
-Theoretically, shorter windows have higher temporal resolution and are sensitive to abrupt changes but may not be robust enough to make holistic judgement.
-In contrast, longer windows can smooth noises and are more informative but may not be sensitive enough or may introduce irrelevant data from heterogeneous activities.
-Determining the optimal window size is therefore a key step for the model to reliably detect infants’ restrained periods.
+Strategies to address this untested classification lie in the flexibility of modeling approaches.
+One strategy is adjusting the window length within which movement data are aggregated.
+With a fixed sampling rate, larger windows provide more information for the model to leverage to predict the current category.
+Prior studies have used window lengths ranging from 2.3 to 4 seconds for infant posture category [@Airaksinen2020j; @Franchak2024e], and from 1.15 to 10 seconds for infant holding and carrying detection [@Airaksinen2024z].
+Considering that transitions between restraint states are typically slower than transitions between postures, we set window lengths longer than 4 seconds.
+Theoretically, shorter windows have higher temporal resolution and are sensitive to abrupt changes, but less data might mean the model lacks context to make an accurate classification.
+In contrast, longer windows can incorporate more data but may lack temporal sensitivity or may introduce irrelevant data from heterogeneous movement within an event.
+Determining the optimal window length is therefore a key step for the model to reliably detect infants’ restrained periods.
 
@@
 
-In short, we propose that wearable sensors offer a more objective, unobtrusive, information-rich, and efficient approach to capturing infants’ full-day experiences in natural settings compared to prior methods.
-However, unique challenges remain in our new effort to detect infants’ restrained periods.
-We used an existing data set with infants aged from 4 to 14 months.
-Infants wore a special garment with four sensors tucked in pockets—two on each side of thighs and two on each side of ankles.
-In each visit, the family received the equipment dropped off at their doorstep in the morning.
-The caregiver was instructed to put the garment on the infant to record their whole-day movement and set up a third-person camera to record the first 1.5 hours of the study as video ground truth for human annotation.
-The current study tested whether a supervised machine learning model trained on human annotated labels can accurately classify infants’ restrained status.
-We varied the window size as 4s, 16s, and 30s and tested the model performance.
+We propose that wearable sensors offer a more objective, unobtrusive, information-rich, and efficient approach to capturing infants’ full-day restraint experiences in natural settings compared to prior methods.
+To test the approach, we used an existing data set with infants aged from 4 to 14 months (Franchak, 2023).
+The data set consists of 142 sessions of full-day wearable sensor recording and 1.5-hour video recording of infants’ movement in the home.
+We coded infants’ restrained status from the videos to provide ground truth labels.
+We then trained a supervised machine learning model based on the human annotated labels.
+We varied the window lengths as 4 s, 16 s, and 30 s to examine the impact of window length on model performance.
 
-We tested the model in two ways.
-First, we compared how accurately model predictions matched human annotation for the 1.5 hours of the video-recorded session.
-We defined \>85% agreement as a good accuracy, based on prior studies reporting 85–96% accuracy in classifying infants’ movement at home.
+We validated the model in two ways.
+First, we compared how accurately model predictions matched human annotation for the 1.5 hours of each video-recorded session.
+We predicted that accuracy would be good, exceeding \~85% agreement based on prior studies reporting 78—98% accuracy [@Airaksinen2020j; @Airaksinen2024z; @Franchak2024e; @Yao2019q] in classifying infants’ movement at home.
+We also compared across window lengths according to overall accuracy and prediction bias to decide which window length was most optimal.
 Second, we applied the model to full-day data to estimate the proportion of time each infant was restrained.
-We tested the convergent validity of the model against past research to determine whether restrained time decreased with age [@Carson2022n; @Franchak2019o; @Franchak2024l].
+We examined the convergent validity of the model against past research by testing whether restrained time decreased with age [@Carson2022n].
 
@@
 
-The data set was published on Databrary (Franchak, 2023) and used in other empirical works (XX).
+The data set is hosted on Databrary (Franchak, 2023) and has been used in other publications.
 The data set consists of two cohorts of infants: `r apa_num(age[1,2])` younger infants aged from 4 to 7 months of age, and `r apa_num(age[2,2])` older infants aged from 11 to 14 months of age.
@@
 Caregivers reported the infants’ race as White (*n* = 23), Hispanic or Latinx (*n* = 8), African American (*n* = 2), Asian (*n* = 2), more than one (*n* = 15), or not fitting into the above races or chose not to answer (*n* = 15).
-Families were recruited from Southern California via social media and community recruitment events and received \$30 for compensation for each visit.
-Caregivers provided informed consent before each session of the study.
-The whole procedures were approved by the University of California, Riverside Institutional Review Board.
 
@@
 
-As shown in figure XX, a special infant garment was customized to secure four lightweight wearable sensors (MC10 Biostamp) at four specific locations (i.e., right thigh, left thigh, right ankle, and left ankle).
+In the data set, a special infant garment was customized to secure four lightweight wearable sensors (MC10 Biostamp) at four specific locations (i.e., right thigh, left thigh, right ankle, and left ankle).
 Each sensor collects raw data as accelerometer signals (in x, y, and z orientations) and gyroscope signals (in roll, pitch, and yaw orientations).
-The sampling rate of 62.5 Hz allows for millisecond-level granularity, and the sufficient battery life and built-in storage allows for whole-day-long recording.
-An action camera fixed on a mini tripod (GoPro HERO9 Black) was used to record a third-person video of the first 1.5 hours into the session as ground truth.
-The camera and sensors were synchronized at the beginning of each session.
+The sampling rate of 62.5 Hz allows for high resolution motion recording, and the sufficient battery life and built-in storage allow for whole-day-long recording.
+An action camera fixed on a mini tripod (GoPro HERO9 Black) was used to record synchronized third-person video of the first 1.5 hours into the session as ground truth.
 
@@
 
-On the day of a session, a researcher arrived at the family’s house in the morning with a bucket of all equipment.
-The researcher turned on the camera and synchronized it with the sensors by dropping the garment in front of the camera—the action would leave a distinctive timestamp on both the camera and the sensors.
-The researcher then called the caregiver to pick up the bucket and instructed the caregiver to set up the equipment.
-The caregiver was told to place the camera to where it could capture the main room and to move the camera if needed to ensure the infant was always in view.
-The caregiver was instructed to put the garment on the infant and position the infant in various postures (including supine, prone, sitting, crawling, standing, walking, being held, and being restrained in seating devices) for at least one minute each for model training purpose.
-Next, the caregiver played with the infant as usual in front of the camera for 10 minutes.
-Afterwards, the family went about their day with the garment on the infant.
-The caregiver logged the time when the infant napped or that the garment was removed for reasons such as diaper changes, baths, going out on errands, etc.
-The researcher picked up the equipment the next day.
+Data collection took place in the family’s home and began with synchronizing the video camera and wearable sensors.
+The caregiver was instructed to set up the camera to capture the main room and to move it as needed to ensure the infant remained in view.
+The caregiver put the garment on the infant and positioned the infant in various postures (including holding and sitting in a restraint device) for at least one minute each for model training purpose.
+Afterwards, the infant continued wearing the garment during their daily routines.
+The caregiver logged the times when the infant was napping or when the garment was removed (e.g., for diaper changes, baths, or errands).
 
@@
 
-Trained human coders watched the 1.5-hour videos and annotated the periods when the infant was restrained or unrestrained using Datavyu software (https://datavyu.org/).
-Restrained and unrestrained is defined as a binary category, meaning that the infant could only be in either of the two statuses unless they were out of view.
-We define restrained periods as when infants cannot initiate or control their own movements, which includes being held or carried by caregivers, or being restrained in devices that constrain them from changing body position.
-Unrestrained periods, on the contrary, are when infants could freely change their body position or locomote.
-Each video file was coded by two coders: one primary coder annotated the whole 1.5-hour video, and an independent reliability coder annotated the first 30 minutes of the video.
-The inter-rater reliability was calculated as the proportion of video frames where the two coders coded the same label.
+Using the 1.5-hour videos, our team annotated the periods when the infant was restrained or unrestrained using Datavyu software (https://datavyu.org/).
+Restrained and unrestrained was defined as a binary category, meaning that all times were coded as one of the two categories.
+A period was coded as missing if the infant went out of view and was excluded from further analysis.
+Restraint was defined as when infants could not initiate or control their own movements, which included being held or carried by caregivers, or being restrained in devices that constrain them from changing body position.
+Unrestrained periods, in contrast, were when infants could freely change their body position or locomote.
+For example, an infant was considered restrained when strapped into a high chair seat, but not when they were placed in a crib and free to change position.
+Each video file was coded by two coders: a primary coder annotated the whole 1.5-hour video, and an independent reliability coder annotated the first 30 minutes of the video.
+Inter-rater reliability was calculated as the proportion of video frames where the two coders coded the same label.
 The human annotation reached a high agreement with an overall accuracy of `r apa_num(stats_human[1,1]*100)`%, ranging from `r apa_num(stats_human[1,2]*100)`% to `r apa_num(stats_human[1,3]*100)`%, and an average kappa of `r apa_num(stats_human[1,4])`, ranging from `r apa_num(stats_human[1,5])` to `r apa_num(stats_human[1,6])`.
@@
 
-The machine-learning classification process followed procedures similar to [@Franchak2021k; @Franchak2024e].
+The supervised machine-learning classification process followed procedures similar to prior work [@Franchak2021k; @Franchak2024e; @Nam2013y].
 First, we aligned the time series of human annotations and sensor data by matching the synchronization time stamp in the video and sensors.
@@
 
-Building on previous methods, we focused on short temporal windows to capture the dynamic nature of movement.
-Accordingly, we aggregated raw IMU data into motion features within sliding windows of n seconds.
+We aggregated raw IMU data into motion features within sliding windows of *n* seconds, where the window length *n* varied from 4 s to 16 s and 30 s.
 For each window, we computed statistical features—mean, median, minimum, 25th percentile, 75th percentile, maximum, skewness, kurtosis, standard deviation, and sum—for all six signals across the four sensors, resulting in 240 features.
@@
 
-We started the windowing process from the syncing point and slide the window every *n*/2 seconds.
+We started the windowing process from the synchronization point and slid the window every *n*/2 seconds.
 Consequently, each time point spaced every *n*/2 seconds contains 436 motion features, aggregated from the *n*/2 seconds before and after that point.
 For model training and testing, we included only time windows where the infant remained in a single status, so that each window could be assigned one unambiguous label.
-The window size *n* varied from 4s, to 16s and 30s.
 
@@
 We applied the random forest algorithm (Breiman, 2001) in modeling using the *DecisionTree* package (Sadeghi et al., 2022) in Julia.
-We used a “training” set of the windowed data set and validated the models by comparing their predictions on a separate “testing” set to the corresponding human annotations.
-Specifically, we used a leave-one-out cross-validation approach—each session was held out once and only as the testing set—to evaluate how well the model generalized to unseen data.
-After validation, an optimal window size was picked and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.
+We used a leave-one-out cross-validation approach, where we held out one session at each iteration as testing set and trained the model using the other sessions,to evaluate how well the model generalized to unseen data.
+After validation, an optimal window length was picked[,]{.underline} and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.
+
+## Data Sharing
+
+Machine-learning process was conducted in Julia version 1.11.1 using the DecisionTree package (Sadeghi et al., 2022).
+Model evaluation, statistical analyses, and this reproducible manuscript were conducted in RStudio environment (R version 4.5.0; RStudio version 2025.5.0.496; RStudioTeam, 2024) using the R packages tidyverse [@tidyverse], flextable [@flextable], and papaja [@papaja], as well as the apa quarto template (Schneider, n.d.).
+All human annotation results and processing scripts are publicly shared on GitHub (XXX).
 
 # Results
-We first compared the model’s performance to the human annotation of the first 1.5 hours across all 142 sessions using leave-one-out cross-validation approach.
-We reported performance metrics including accuracy, kappa, sensitivity, positive predictive value, F1, and the correlation of overall estimate between the modal and human annotation.
-We also examined prediction biases of the model by looking at class-specific metrics--sensitivity and positive predictive value--and the prediction error of the overall estimate.
-Second, we applied the model to process infants’ full day sensor data and examined the model’s convergent validity by comparing its estimate and age-related change to previous works.
+
+We first compared the model’s performance to the human annotation of the 1.5-hour video recorded period across all 142 sessions using leave-one-out cross-validation approach.
+We reported performance metrics including accuracy, kappa, sensitivity, positive predictive value, F1, and the correlation of overall estimate time of restraint between the model and human annotation.
+We also examined prediction biases model and how they vary across window lengths.
+Second, we applied the model with the optimal window length to process infants’ full-day sensor data and examined the model’s convergent validity by replicating a previously-reported effect: that restraint time decreases with age [@Carson2022n].
 
@@
 #| ft.align: left
-#| apa-note: F, df, and p value indicate one-way ANOVA on each metric comparing across the three window lengths
+#| apa-note: F, df, and p value indicate one-way ANOVA on each metric comparing across the three window lengths.
 
@@
 
-@tbl-indexbetweenmodels shows that all window lengths reached high accuracy (`r apa_num(table_btw_mdl[1,3])` for 4 s, `r apa_num(table_btw_mdl[1,4])` for 16 s, and `r apa_num(table_btw_mdl[1,5])` for 30 s), and good kappa value (`r apa_num(table_btw_mdl[2,3])` for 4 s, `r apa_num(table_btw_mdl[2,4])` for 16 s, and `r apa_num(table_btw_mdl[2,5])` for 30 s).
-All the window lengths also presented high sensitivity (ranging from `r apa_num(min(table_btw_mdl[3,3:5]))` to `r apa_num(max(table_btw_mdl[3,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[4,3:5]))` to `r apa_num(max(table_btw_mdl[4,3:5]))` for unrestrained prediction), high positive predictive value (ranging from `r apa_num(min(table_btw_mdl[5,3:5]))` to `r apa_num(max(table_btw_mdl[5,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[6,3:5]))` to `r apa_num(max(table_btw_mdl[6,3:5]))` for unrestrained prediction), and high F1 score (ranging from `r apa_num(min(table_btw_mdl[7,3:5]))` to `r apa_num(max(table_btw_mdl[7,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[8,3:5]))` to `r apa_num(max(table_btw_mdl[8,3:5]))` for unrestrained prediction).
-Moreover, the three window lengths did not present significant differences in any of the metrics (*p*s >= `r apa_num(min(table_btw_mdl[1:8,8]))`).
+@tbl-indexbetweenmodels shows that all window lengths reached high accuracy (`r apa_num(table_btw_mdl[1,3])` for 4 s, `r apa_num(table_btw_mdl[1,4])` for 16 s, and `r apa_num(table_btw_mdl[1,5])` for 30 s), and good kappa values (`r apa_num(table_btw_mdl[2,3])` for 4 s, `r apa_num(table_btw_mdl[2,4])` for 16 s, and `r apa_num(table_btw_mdl[2,5])` for 30 s).
+All window lengths presented high sensitivity—how often the model correctly detected when the infant was actually restrained or unrestrained—ranging from `r apa_num(min(table_btw_mdl[3,3:5]))` to `r apa_num(max(table_btw_mdl[3,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[4,3:5]))` to `r apa_num(max(table_btw_mdl[4,3:5]))` for unrestrained prediction.
+All window lengths had high positive predictive value (PPV)—how often the model’s predictions of being restrained or unrestrained were correct—ranging from `r apa_num(min(table_btw_mdl[5,3:5]))` to `r apa_num(max(table_btw_mdl[5,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[6,3:5]))` to `r apa_num(max(table_btw_mdl[6,3:5]))` for unrestrained prediction.
+All window lengths also had high F1 scores—a balanced measure that combines sensitivity and PPV—ranging from `r apa_num(min(table_btw_mdl[7,3:5]))` to `r apa_num(max(table_btw_mdl[7,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[8,3:5]))` to `r apa_num(max(table_btw_mdl[8,3:5]))` for unrestrained prediction.
+Moreover, the three window lengths did not significantly differ for any of the metrics (*p*s \>= `r apa_num(min(table_btw_mdl[1:8,8]))`).
 
-Meanwhile, @tbl-prevalencecor shows that all window lengths provided reliable overall estimate of restrained time.
-The prevalence of restrained periods the models predicted was strongly correlated with the prevalence annotated by human coders, the correlation coefficient ranging from `r apa_num(table_prev[1,4])` for 4 s and `r apa_num(table_prev[3,4])` for 30 s.
+@fig-prevalencegap A shows that, at session level, all window lengths provided reliable estimate of restrained time.
+We defined prevalence as the proportion of windows classified as restrained among all windows for each session.
+The prevalence predicted by the model was highly correlated with the prevalence annotated by human coders (*r* = `r apa_num(table_prev[1,4])` for 4 s, *r* = `r apa_num(table_prev[2,4])` for 16 s, and *r* = `r apa_num(table_prev[3,4])` for 30 s).
 
-### Bias Towards Unrestrained Class Across Window Lengths
+### Models were biased towards predicted unrestrained
 
-Despite the overall good performance, all window lengths presented prediction bias towards unrestrained class.
-@fig-cibtwctg show that all window lengths had lower sensitivity and higher positive predictive value (PPV) for restrained compared to unrestrained periods, indicated by the negative difference of sensitivity and positive difference of PPV.
-The sensitivity difference for 4 s window was `r apa_num(table_btw_ctg[1,3])`, 95% bootstrapped CI [`r apa_num(table_btw_ctg[1,4])`, `r apa_num(table_btw_ctg[1,5])`], for 16 s window was `r apa_num(table_btw_ctg[3,3])`, 95% CI [`r apa_num(table_btw_ctg[3,4])`, `r apa_num(table_btw_ctg[3,5])`], and for 30 s window was `r apa_num(table_btw_ctg[5,3])`, 95% CI [`r apa_num(table_btw_ctg[5,4])`, `r apa_num(table_btw_ctg[5,5])`].
-The PPV difference for 4 s window was `r apa_num(table_btw_ctg[2,3])`, 95% CI [`r apa_num(table_btw_ctg[2,4])`, `r apa_num(table_btw_ctg[2,5])`], for 16 s window was `r apa_num(table_btw_ctg[4,3])`, 95% CI [`r apa_num(table_btw_ctg[4,4])`, `r apa_num(table_btw_ctg[4,5])`], and for 30 s window was `r apa_num(table_btw_ctg[6,3])`, 95% CI [`r apa_num(table_btw_ctg[6,4])`, `r apa_num(table_btw_ctg[6,5])`].
-The differences indicate that the model tended to under-detect restrained periods but was reliable when labeling a period as restrained.
-Conversely, it overestimated unrestrained periods, with lower reliability in unrestrained prediction.
-While none of the window lengths were free from bias, the 30 s window presented a more balanced detection between restrained and unrestrained periods.
+Despite the overall good accuracy, models using all three window lengths were biased towards predicted that infants were unrestrained rather than restrained.
+@fig-cibtwctg shows the difference of a metric’s value (e.g., sensitivity, PPV) between restrained class and unrestrained class within a single window length.
+Larger difference scores (away from 0) indicate larger asymmetry in predicting one category rather than the other.
+@fig-cibtwctg A shows that the sensitivity difference for 4 s window was `r apa_num(table_btw_ctg[1,3])`, 95% bootstrapped CI \[`r apa_num(table_btw_ctg[1,4])`, `r apa_num(table_btw_ctg[1,5])`\], for 16 s window was `r apa_num(table_btw_ctg[3,3])`, 95% CI \[`r apa_num(table_btw_ctg[3,4])`, `r apa_num(table_btw_ctg[3,5])`\], and for 30 s window was `r apa_num(table_btw_ctg[5,3])`, 95% CI \[`r apa_num(table_btw_ctg[5,4])`, `r apa_num(table_btw_ctg[5,5])`\].
+The negative differences of sensitivity across all window lengths indicate that the model was more sensitive in detecting the unrestrained class compared to restrained class regardless of window lengths.
+@fig-cibtwctg B shows that the PPV difference for 4 s window was `r apa_num(table_btw_ctg[2,3])`, 95% CI \[`r apa_num(table_btw_ctg[2,4])`, `r apa_num(table_btw_ctg[2,5])`\], for 16 s window was `r apa_num(table_btw_ctg[4,3])`, 95% CI \[`r apa_num(table_btw_ctg[4,4])`, `r apa_num(table_btw_ctg[4,5])`\], and for 30 s window was `r apa_num(table_btw_ctg[6,3])`, 95% CI \[`r apa_num(table_btw_ctg[6,4])`, `r apa_num(table_btw_ctg[6,5])`\].
+The positive differences of PPV across all window lengths indicate that the model’s prediction of restrained periods was more credible compared to its prediction of unrestrained periods.
+Notably, while none of the window lengths were free from bias, the 30 s window presented smaller offsets of metric values between restrained and unrestrained detection.
 
 The prediction bias was also evident when examining the model’s overall estimate of restrained time and its difference from human annotation.
-@tbl-prevalencecor and @fig-prevalencegap show a consistent underestimate of restrained periods despite the high correlation of the overall estimate.
-The 4 s window missed `r apa_num(abs(table_prev[1,5])*100)`% of time that were annotated as restrained (95% CI [`r apa_num(abs(table_prev[1,6])*100)`%, `r apa_num(abs(table_prev[1,7])*100)`%]), the 16 s window missed `r apa_num(abs(table_prev[2,5])*100)`% (95% CI [`r apa_num(abs(table_prev[2,6])*100)`%, `r apa_num(abs(table_prev[2,7])*100)`%]]), and the 30 s window missed the least amount of `r apa_num(abs(table_prev[3,5])*100)`% (95% CI [`r apa_num(abs(table_prev[3,6])*100)`%, `r apa_num(abs(table_prev[3,7])*100)`%]).
-Taken together, 30 s turned out to be the best window length out of the three, which brought about a balanced performance across classes.
-We then applied the 30 s model to process the full-day sensor data.
+@fig-prevalencegap B shows that the model predictions consistently underestimated overall restrained time compared to human annotation, despite the high correlation between the two.
+We defined prediction error as the model predicted prevalence (in % units) subtracted by the human annotated prevalence of restrained periods (in % units).
+The 4 s window had a prediction error of `r apa_num(table_prev[1,5]*100)`%, 95% CI \[`r apa_num(table_prev[1,6]*100)`%, `r apa_num(table_prev[1,7]*100)`%\], the 16 s window of `r apa_num(table_prev[2,5])*100`%, 95% CI \[`r apa_num(table_prev[2,6]*100)`%, `r apa_num(table_prev[2,7]*100)`%\], and the 30 s window of `r apa_num(table_prev[3,5]*100)`%, 95% CI \[`r apa_num(table_prev[3,6]*100)`%, `r apa_num(table_prev[3,7]*100)`%\].
+
+Across the metrics of sensitivity, PPV, and prediction error, 30 s turned out to be the best window length out of the three, which brought about a balanced performance across classes.
 
@@
 
-In order to compare the model estimate of infants’ full day experience with previous studies, we excluded eleven sessions whose total recording time was less than 3 hours, which were useful for training and validating the models but insufficient for measuring day-long experience.
+We applied the best model based on the analyses in the previous (30-s window length) to create full-day predictions for all sessions that included full-day experiences.
+We excluded eleven sessions whose total recording time was less than 3 hours, which were useful for training and validating the models but insufficient for measuring day-long experience.
 We also excluded another two sessions where the family took the garment off for a substantial portion of time but did not log the time stamp of removal.
-The remaining 129 sessions have an average recording time of `r apa_num(mean(ds_sum_excl$total_time))` hours (*SD* = `r apa_num(sd(ds_sum_excl$total_time))`).
+The remaining 129 sessions had an average recording time of `r apa_num(mean(ds_sum_excl$total_time))` hours (*SD* = `r apa_num(sd(ds_sum_excl$total_time))`).
 
-The model prediction on the 129 full-day sessions converged with previous studies [@Carson2022n; @Franchak2019o; @Franchak2024l].
-Younger infants aged from 4 to 7 months spent *M* = `r apa_num(stats_fullday[1,2]*100)`% (*SD* = `r apa_num(stats_fullday[
-1,3]*100)`%) of awake time restrained, and older infants aged from 11 to 14 months spent *M* = `r apa_num(stats_fullday[2,2]*100)`% (*SD* = `r apa_num(stats_fullday[2,3]*100)`%) of awake time restrained, which is comparable to the 46% restrained time for infants aged from 10 to 13 months in the previous EMA study [@Franchak2024l].
+Model predicted restrained and unrestrained time converged with previous studies [@Carson2022n].
+Younger infants aged from 4 to 7 months spent *M* = `r apa_num(stats_fullday[1,2]*100)`% (*SD* = `r apa_num(stats_fullday[ 1,3]*100)`%) of awake time restrained, and older infants aged from 11 to 14 months spent *M* = `r apa_num(stats_fullday[2,2]*100)`% (*SD* = `r apa_num(stats_fullday[2,3]*100)`%) of awake time restrained.
 Moreover, @fig-agechange shows a significant group difference in restrained time between older infants and younger infants (*t* = `r apa_num(t_age[2,3])`, *p* `r apa_num(ifelse(t_age[2,4] < .001, "<.001", t_age[2,4]))`), supporting the convergent validity of the current system.
 
-# Discussions
-We validated a system that combines wearable sensors and machine learning models to quantify infants’ restrained experiences in the home environment.
-The model presented high accuracy and substantial kappa agreement with human-annotated ground truth.
-The model also presented convergent validity in overall estimate of restraint time and age-related change compared with previous studies.
+# Discussion
 
-We also examined the optimal window length for the classification (4 s, 16 s, and 30 s).
-All window lengths showed a slight prediction bias towards unrestrained class, which might be due to the inherent imbalance in the training data set—unrestrained periods were more prevalent than restrained periods.
-However, the 30 s window offered the most balanced detection compared to shorter windows.
-This advantage might be due to the contextual nature of restraint: infants may exhibit similar body movement whether restrained or not, and a restrained bout usually lasts minutes rather than seconds.
-A longer window can provide richer contextual information for distinguishing restrained from unrestrained states, whereas shorter windows only capture momentary movements that may appear similar across restrained and unrestrained periods.
+In the current study, we developed and validated a machine learning algorithm to detect infants’ restrained experiences from wearable sensors during home life.
+The model presented high accuracy and substantial kappa agreement compared to human-annotated ground truth, which are comparable to other wearable sensor algorithms classifying other infant movement categories in the home [@Airaksinen2020j; @Airaksinen2024z; @Franchak2024e; @Yao2019q].
+The model presented a slight bias towards predicted unrestrained—the model tended to overestimate unrestrained periods but was not as credible when labeling a period as unrestrained.
+The bias might be due to the inherent imbalance in the training data set where unrestrained periods were more prevalent than restrained periods, indicated by human annotated prevalence.
+However, using a longer window length (i.e., 30 s, rather than 16 s or 4 s) mitigated the prediction bias.
+Ultimately, the model with 30 s window length presented convergent validity in overall estimate of restraint time and age-related change compared with previous findings [@Carson2022n].
 
-A key contribution of this work is the extension of wearable-sensor methodology to detect infant restraint in naturalistic settings.
-Wearable sensors have been established as a reliable and unobtrusive approach for capturing full-day, time-series movement data.
-Previous works validated the use of wearable sensors in classifying infant positions.
-Our work shows the configuration of four sensors on thighs and ankles and random forest algorithms can also provide sufficiently accurate classification of infant restraint.
-Although the system cannot be as accurate as human annotation, the trade-off between accuracy and scale favors the sensor-based system, particularly when research questions is oriented at infants’ experiences fluctuating throughout a day or across multiple days.
+Why could a longer window length (i.e., 30 s) provide more balanced detection?
+For one thing, a longer window length can better capture the contextual information of restraint.
+Consider the contextual nature of restraint: infants may exhibit similar body movement whether restrained or not for most of the time, but one subtle occurrence of distinctive movement can signal restraint.
+For example, an infant can be sitting on the floor unrestrained, on caregiver’s lap restrained, or strapped in a seat restrained.
+It is hard to distinguish if the infant is restrained while sitting still, but when the infant moves their limbs, restraint may present differences, such as a stronger damping in the motion decay because the support (i.e., caregiver’s laps, or seats) absorbs the motion.
+These differences only present themselves in the moments of movement, not throughout the whole period.
+Thus, a longer window length can better capture temporal information, whereas shorter windows only capture momentary movements that may appear similar across restrained and unrestrained periods.
+For another, a longer window length can improve generalization across a variety of restraint forms.
+Infants can be restrained in various ways—such as in stationary seats, in jumpers, in swings, or in caregivers’ arms, either still or while being carried around.
+Different restraint forms are presented by different motion features.
+By encompassing a broader range of movement patterns, a longer window can provide more informative and generalizable features for the model to account for variability within the restrained class.
+Given a restrained bout usually lasts minutes rather than seconds, a 30 s window length presents a good balance.
 
-With the validated system, researchers can now study infant restrained experiences in a longer time scale and a greater temporal precision.
-Research questions about the temporal structure of restrained periods can be addressed—for example, how frequently are infants switched between restrained and unrestrained states, or how restrained bouts are distributed across the day or multiple days?
-Moreover, the system enables investigations linking restraint to concurrent learning experiences: How does the temporal profile of restraint relate to motor development?
-How does adult language input or infants’ own vocalization patterns shift when they are restrained versus unrestrained?
+Our analysis of window length informs future modeling works to consider temporal resolution as a behavior-informed decision.
+In infants’ everyday activities, behaviors unfold across various timescales—from second-by-second movement (e.g., locomotion) to longer periods of states (e.g., on the ground vs.
+lifted).
+Our optimal window length for restraint detection turns out to be longer than the window lengths in other models for body position classification [@Airaksinen2020j; @Franchak2021k; @Franchak2024e; @Yao2019q], corresponding to the fact that restraint bouts last longer than position bouts.
+We recommend selecting window lengths that align with typical duration and transition of the targeted behavior for machine learning modeling.
 
-Moving forward, we can refine the model to differentiate between caregiver restraint and device restraint, as these contexts may differ in the perceptual, social, and cognitive input they afford.
-Nonetheless, quantifying infants’ general restrained experiences already represents a significant advance.
-This system provides a powerful tool for capturing infants’ physical contexts and how the contexts shape their daily learning opportunities.
+## Advances in Methodology
+
+The key contribution of this work is introducing a novel approach for measuring infant restraint in the home environment using wearable sensors.
+Wearable sensors offer several advantages over previous methods when quantifying day-long movement in naturalistic settings.
+Compared with survey methods, wearable sensors provide objective, continuous measurements with fine-grained temporal resolution.
+Compared with video recording and human coding method, wearable sensors are less obtrusive and enable substantially longer recordings.
+With the machine learning model established, researchers can extend their investigation timescales to a full day or even multiple days at low cost.
+Although the current machine learning model achieves slightly lower accuracy (89%) than human annotation, the trade-off between accuracy and human labor favors our sensor approach—especially when the target behavior fluctuates over longer timescales beyond the practical limits of human coding.
+Restraint is such a behavior that does not recur repeatedly every hour but follows the rhythm of everyday routines and activities.
+Therefore, a wearable sensor system capable of generating restraint profiles across full days offers greater scalability for measuring infants’ everyday experiences.
+
+The current work also demonstrates that wearable sensors have the potential to characterize other categories of infant movement that might interest researchers in a variety of domains.
+Compared with previous works that classify infants’ positions [@Airaksinen2020j; @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q], restraint might be the most challenging one out of all categories, because the motion between classes can be very similar sometimes—such as sitting on the ground versus sitting on a caregiver’s lap.
+In contrast, other categories (e.g., positions) usually have distinctive motion features between classes—such as the legs’ moving direction in supine position versus upright position.
+Our results, however, showed that even a context-dependent category as restraint can be successfully classified by machine learning models using sensor data.
+Conceivably, this system might perform comparably well in classifying other context-independent categories, such as locomotion, steps, falling, etc.
+Future research needs to further develop the potential of wearable sensors.
+If proven feasible, the new models will substantially reduce the need for human labor and enlarge the data scale.
+
+Beyond expanding the range of classifiable behaviors, future works should also broaden the scenarios in which wearable sensors are applied.
+To date, most applications of wearable sensors have been limited to laboratory or home environments.
+Infants’ outdoor experiences remain largely unexplored, whereas outdoor environments provide infants with rich and unique opportunities for learning and exploration.
+How much time infants spend restrained in stroller or car seat?
+How positions are they frequently in?
+How many steps they take?
+Validating sensor-based models in outdoor scenarios can open new avenues for investigation.
+
+Moving forward, future works can also integrate different categories of model predictions to generate a multifaceted description of infant real-time experience.
+For example, by combing the model prediction of infant positions (i.e., supine, prone, sitting, upright, holding and carrying) with the current model prediction of overall restraint, researchers can easily know what positions infants were in while restrained—such as supine restrained, sitting restrained, or being held—without spending extra efforts in human coding and model training.
+
+## Advances in Understanding Full-day Real-time Restraint
+
+The advances in methodology can facilitate a deeper understanding of infant full-day real-time restraint experiences.
+Restraint, as an external factor imposed by caregivers, plays a critical role in infant real-time experiences—influencing when and how often infants have the opportunities for movement, which portions of the environment they can perceive, and what objects and people they can interact with.
+Prior works provided valuable insights into parental attitudes of restraint and physical activity.
+For example, parental concerns over floor play, beliefs about physical activeness, and practices of co-play are associated with restraint frequency and infants’ physical activity level [@Hanisch2025j; @Hnatiuk2013t; @Prioreschi2017c; @Hesketh2015i; @Karasik2025h].
+However, few studies have examined how each restraint decision is formed in real time and how infants’ sensorimotor experiences are affected in real time.
+
+Having a full-day continuous profile of infant restraint is thus essential in identifying the factors influencing restraint decisions in real time.
+Restraint is believed to serve practical purposes such as safety, transportation, comforting, etc.
+[@Birken2015s], but these functions alone cannot account for all the variability in when caregivers choose to restrain an infant.
+To what extent are such decisions influenced by these intra-individual factors, such as daily routines, momentary tasks, or fluctuations in emotional states?
+To what extent are such decisions influenced by inter-individual factors, such as infant motor skill, attachment style, caregivers’ beliefs, or parental stress?
+And to what extent are the decisions influenced by socio-cultural factors, such as cultural norms [@Karasik2015i; @Karasik2018t], or socioeconomic status?
+The current system’s capability of locating each individual bout of restraint enables more fine-grained analyses of caregiver behavior and provides new insights into the dynamics of everyday caregiving practices.
+
+Having a full-day continuous profile of infant restraint is also essential in understanding infant real-time sensorimotor experiences.
+An overall estimate of restraint time is not enough to examine the mechanisms between restraint practices and development outcomes.
+Take motor development as an example.
+@Karasik2023d found that Tajik infants fell behind US infants in achieving motor milestones despite the comparable overall restrained time between cultural groups, might due to the prolonged individual restrained bouts in Tajik infants [@Karasik2022a].
+The current system allows researchers to characterize the temporal structure of restraint and examine the relationship between restraint and other experiences.
+A prior study [@Malachowski2023a] found device restraint is associated with reduced adult language input using ecological momentary assessment.
+However, a continuous, full-day recording of restraint captures the temporal distribution rather than snapshots.
+Questions can be answered such as: Does adult word count vary based on the duration of each restraint bout?
+How do specific language inputs (e.g., referential words) distribute throughout a day based on restraint states and restraint types?
+Additionally, real-time restraint can also be linked to infants’ self-initiated explorations.
+For example, how do infants’ own vocalizations, attention, locomotion, or object interactions differ across restrained and unrestrained?
+In short, continuous, full-day quantification of infant restraint offers a foundation for linking caregiver practices to infants’ moment-to-moment sensorimotor experiences, thereby deepening our understanding of how daily contexts accumulate shape developmental outcomes.
+
+## Conclusions
+
+To conclude, the current study provided a powerful tool for quantifying infants’ everyday restraint experiences in the home using wearable sensors and machine learning modeling.
+The model achieved high accuracy and strong agreement with human annotations, as well as convergent validity by showing an age-related decrease in infant restraint time.
+We explored the window length in the data aggregating procedure of modeling and proved that a longer window enhanced the model’s ability to capture the contextual heterogeneous nature of restraint.
+Together, the current study advanced the method for measuring infant full-day, real-time movement, and opened avenues for research on how external constraints from caregivers shape early development.
 

### 2025-11-09 manuscript_restrained_classification validation.qmd
Fixed typo

@@
   disclosures:
+    # Acknowledge and cite data/materials to be shared.
+    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
     # Example: This study was registered at ClinicalTrials.gov (Identifier NTC998877).
     study-registration: null
-    # Acknowledge and cite data/materials to be shared.
     data-sharing: null
@@
     gratitude: null
-    # Example. Because the authors are equal contributors, order of authorship was determined by a fair coin toss.
     authorship-agreements: null
@@
     # for now, tables and figures do not render properly in jou mode. 
+    # can be 10pt, 11pt, or 12pt
+    # Integrate tables and figures in text
+    # a4 paper if true, letter size if false
+    # suppresses loading of the lmodern font package
+    # Suppresses loading of the fontenc package
+    # Suppress the title above the introduction
+    # In jou mode, use times or pslatex instead of txfonts
+    # In jou mode, use Computer Modern font instead of times
+    # In jou mode, cancels automatic stretching of tables to column width 
+    # Uses Helvetica font in stu and man mode
+    # In man and stu mode, neutralizes the \helvetica command
+    # In man and stu mode, uses typewriter font
+    # Puts draft watermark on first page
+    # Puts draft watermark on all pages
+    # Masks references that are marked as the author's own
+    # Hides correspondence text
     documentmode: man
-    # can be 10pt, 11pt, or 12pt
     fontsize: 12pt
-    # Integrate tables and figures in text
     floatsintext: false
-    # a4 paper if true, letter size if false
     a4paper: false
-    # suppresses loading of the lmodern font package
     nolmodern: false
-    # Suppresses loading of the fontenc package
     nofontenc: false
-    # Suppress the title above the introduction
     donotrepeattitle: false
-    # In jou mode, use times or pslatex instead of txfonts
     notxfonts: false
-    # In jou mode, use Computer Modern font instead of times
     notimes: false
-    # In jou mode, cancels automatic stretching of tables to column width 
     notab: false
-    # Uses Helvetica font in stu and man mode
     helv: false
-    # In man and stu mode, neutralizes the \helvetica command
     nosf: false
-    # In man and stu mode, uses typewriter font
     tt: false
-    # Puts draft watermark on first page
     draftfirst: false
-    # Puts draft watermark on all pages
     draftall: false
-    # Masks references that are marked as the author's own
     mask: false
@@
     duedate: null
-    # Hides correspondence text
     nocorrespondence: false
@@
 ```{r fig-cibtwctg}
-#| fig-cap: Differences in sensitivity and PPV between restrained and unrestrained prediction according to window lengths
+#| fig-cap: Differences in sensitivity and PPV between restrained and unrestrained prediction according to window length
 #| fig-width: 8
@@
 We defined prediction error as the model predicted prevalence (in % units) subtracted by the human annotated prevalence of restrained periods (in % units).
-The 4 s window had a prediction error of `r apa_num(table_prev[1,5]*100)`%, 95% CI \[`r apa_num(table_prev[1,6]*100)`%, `r apa_num(table_prev[1,7]*100)`%\], the 16 s window of `r apa_num(table_prev[2,5])*100`%, 95% CI \[`r apa_num(table_prev[2,6]*100)`%, `r apa_num(table_prev[2,7]*100)`%\], and the 30 s window of `r apa_num(table_prev[3,5]*100)`%, 95% CI \[`r apa_num(table_prev[3,6]*100)`%, `r apa_num(table_prev[3,7]*100)`%\].
+The 4 s window had a prediction error of `r apa_num(table_prev[1,5]*100)`%, 95% CI \[`r apa_num(table_prev[1,6]*100)`%, `r apa_num(table_prev[1,7]*100)`%\], the 16 s window of `r apa_num(table_prev[2,5]*100)`%, 95% CI \[`r apa_num(table_prev[2,6]*100)`%, `r apa_num(table_prev[2,7]*100)`%\], and the 30 s window of `r apa_num(table_prev[3,5]*100)`%, 95% CI \[`r apa_num(table_prev[3,6]*100)`%, `r apa_num(table_prev[3,7]*100)`%\].
 

### 2025-11-09 train_group_loocv_restrained.jl
Update train_group_loocv_restrained.jl
Cleaned the code

@@ -4,8 +4,9 @@ using CSV, DataFrames, DataFrameMacros, Chain, Statistics, RData
 
 ############# BEFORE RUNNING DIFFERENT BATCHES, MODIFY:
 # Load compiled data from R
+# objs = RData.load("data_compiled/compiled_data_restrained_4s.RData") 
 objs = RData.load("data_compiled/compiled_data_restrained_16s.RData") 
-# objs = RData.load("data_compiled/compiled_data_restrained.RData") 
+# objs = RData.load("data_compiled/compiled_data_restrained.RData") #30s compiled data
 ds = objs["slide_filt"]
 
 # Decide which folder to store individual predictions
@@ -26,13 +27,13 @@ window = 16
 # save_file = "metrics_restrained_30_no_pos.csv"
 # save_file = "metrics_restrained_4_pos.csv"
 # save_file = "metrics_restrained_4_no_pos.csv"
-save_file = "metrics_restrained_16_no_pos.csv"
+# save_file = "metrics_restrained_16_no_pos.csv"
 ####################################################
 
 # Set seed for random forest
 Random.seed!(2025)
 
-# Delete pilot study, subject 104
+# Exclude pilot study, subject 104
 ds = filter!(row -> row.id != "104/1", ds) 
 
 # Pull the ids to set up the LOOCV analysis
@@ -85,6 +86,7 @@ Threads.@threads for i in eachindex(ids)
     # N_trees has a big influence on runtime; 150 is a reasonable estimate, but models tend to work better with 500
     model = RandomForestClassifier(n_trees=150)
     DecisionTree.fit!(model,X,y)
+    impurity_importance(model)
 
     # Predict testing data codes from resulting model
     code_predicted = MLJBase.coerce(DecisionTree.predict(model, Matrix(testing[:,Not(["code", "id"])])), Multiclass) 
@@ -97,18 +99,18 @@ Threads.@threads for i in eachindex(ids)
     # Calculate accuracy and store in the dataframe
     # accuracy = MLJBase.accuracy(code_predicted, code_actual) -- This code does not work: UndefVarError: `accuracy` not defined in `MLJBase`
     # accuracy = sum(code_predicted .== code_actual) / length(code_actual)
-    accuracy = StatisticalMeasures.accuracy(code_predicted, code_actual)
-    sensitivity = StatisticalMeasures.sensitivity(code_predicted, code_actual)
[…]

### 2025-11-09 manuscript_restrained_classification validation.qmd
fixed typo

@@
 Model evaluation, statistical analyses, and this reproducible manuscript were conducted in RStudio environment (R version 4.5.0; RStudio version 2025.5.0.496; RStudioTeam, 2024) using the R packages tidyverse [@tidyverse], flextable [@flextable], and papaja [@papaja], as well as the apa quarto template (Schneider, n.d.).
-All human annotation results and processing scripts are publicly shared on GitHub (XXX).
+All human annotation results and processing scripts are publicly shared on GitHub (https://github.com/Hanzhi-W/ms_restrained_classification-validation.git).
 
@@
 Compared with previous works that classify infants’ positions [@Airaksinen2020j; @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q], restraint might be the most challenging one out of all categories, because the motion between classes can be very similar sometimes—such as sitting on the ground versus sitting on a caregiver’s lap.
-In contrast, other categories (e.g., positions) usually have distinctive motion features between classes—such as the legs’ moving direction in supine position versus upright position.
+In contrast, positions usually have distinctive motion features between classes—such as the legs’ moving direction in supine position versus upright position.
 Our results, however, showed that even a context-dependent category as restraint can be successfully classified by machine learning models using sensor data.
-Conceivably, this system might perform comparably well in classifying other context-independent categories, such as locomotion, steps, falling, etc.
+Conceivably, this system might perform comparably well or better in classifying other context-independent categories, such as locomotion, steps, falling, etc.
 Future research needs to further develop the potential of wearable sensors.
-If proven feasible, the new models will substantially reduce the need for human labor and enlarge the data scale.
+If proven feasible, the new models will substantially reduce the need for human labor in coding and enlarge the data scale.
 
 Beyond expanding the range of classifiable behaviors, future works should also broaden the scenarios in which wearable sensors are applied.
-To date, most applications of wearable sensors have been limited to laboratory or home environments.
-Infants’ outdoor experiences remain largely unexplored, whereas outdoor environments provide infants with rich and unique opportunities for learning and exploration.
+To date, most applications of wearable sensors have been limited to laboratory or in-home environments.
+Infants’ outdoor experiences remain largely underexplored, whereas outdoor environments provide infants with rich and unique opportunities for learning and exploration [@Schneider2025d; @Herrington1998o; @Fjortoft2000o].
 How much time infants spend restrained in stroller or car seat?
-How positions are they frequently in?
+Which positions are they frequently in?
 How many steps they take?
@@
 
-Moving forward, future works can also integrate different categories of model predictions to generate a multifaceted description of infant real-time experience.
+Moving forward, future works can also integrate different categories of model predictions to generate a multifaceted description of infant real-time experiences.
 For example, by combing the model prediction of infant positions (i.e., supine, prone, sitting, upright, holding and carrying) with the current model prediction of overall restraint, researchers can easily know what positions infants were in while restrained—such as supine restrained, sitting restrained, or being held—without spending extra efforts in human coding and model training.
@@
 Prior works provided valuable insights into parental attitudes of restraint and physical activity.
-For example, parental concerns over floor play, beliefs about physical activeness, and practices of co-play are associated with restraint frequency and infants’ physical activity level [@Hanisch2025j; @Hnatiuk2013t; @Prioreschi2017c; @Hesketh2015i; @Karasik2025h].
-However, few studies have examined how each restraint decision is formed in real time and how infants’ sensorimotor experiences are affected in real time.
+For example, parental concerns over floor play, beliefs about physical activeness, and practices of co-play are found to be associated with restraint frequency and infants’ physical activity level [@Hanisch2025j; @Hnatiuk2013t; @Prioreschi2017c; @Hesketh2015i; @Karasik2025h].
+However, few studies have examined how each restraint decision is formed in real time, and how infants’ sensorimotor experiences are affected in real time.
 
@@
 [@Birken2015s], but these functions alone cannot account for all the variability in when caregivers choose to restrain an infant.
-To what extent are such decisions influenced by these intra-individual factors, such as daily routines, momentary tasks, or fluctuations in emotional states?
+To what extent are such decisions influenced by these intra-individual factors, such as daily routines, or momentary tasks?
 To what extent are such decisions influenced by inter-individual factors, such as infant motor skill, attachment style, caregivers’ beliefs, or parental stress?
 And to what extent are the decisions influenced by socio-cultural factors, such as cultural norms [@Karasik2015i; @Karasik2018t], or socioeconomic status?
-The current system’s capability of locating each individual bout of restraint enables more fine-grained analyses of caregiver behavior and provides new insights into the dynamics of everyday caregiving practices.
+The current system’s capability of locating each individual bout of restraint enables more fine-grained analyses of caregiver behaviors and provides new insights into the dynamics of everyday caregiving practices.
 
@@
 Take motor development as an example.
-@Karasik2023d found that Tajik infants fell behind US infants in achieving motor milestones despite the comparable overall restrained time between cultural groups, might due to the prolonged individual restrained bouts in Tajik infants [@Karasik2022a].
-The current system allows researchers to characterize the temporal structure of restraint and examine the relationship between restraint and other experiences.
+@Karasik2023d found that Tajik infants fell behind US infants in achieving motor milestones despite the comparable overall restrained time between cultural groups.
+It is likely due to the prolonged individual restrained bouts in Tajik infants [@Karasik2022a].
+The current system allows researchers to characterize the temporal structure of restraint and examine the relationship between restraint and other experiences, such as language, object interaction, and social interaction.
 A prior study [@Malachowski2023a] found device restraint is associated with reduced adult language input using ecological momentary assessment.
-However, a continuous, full-day recording of restraint captures the temporal distribution rather than snapshots.
+However, a continuous, full-day recording of restraint can capture the temporal distribution rather than snapshots.
 Questions can be answered such as: Does adult word count vary based on the duration of each restraint bout?
@@
 For example, how do infants’ own vocalizations, attention, locomotion, or object interactions differ across restrained and unrestrained?
-In short, continuous, full-day quantification of infant restraint offers a foundation for linking caregiver practices to infants’ moment-to-moment sensorimotor experiences, thereby deepening our understanding of how daily contexts accumulate shape developmental outcomes.
+In short, continuous, full-day quantification of infant restraint offers a foundation for linking caregiver practices to infants’ moment-to-moment sensorimotor experiences, thereby shaping infants' developmental outcomes.
 
@@
 The model achieved high accuracy and strong agreement with human annotations, as well as convergent validity by showing an age-related decrease in infant restraint time.
-We explored the window length in the data aggregating procedure of modeling and proved that a longer window enhanced the model’s ability to capture the contextual heterogeneous nature of restraint.
-Together, the current study advanced the method for measuring infant full-day, real-time movement, and opened avenues for research on how external constraints from caregivers shape early development.
+We explored the window length in the data aggregating procedure of modeling and proved that a longer window enhanced the model’s balanced performance.
+Together, the current study advanced the method for measuring infant full-day, real-time movement, and opened avenues for research on how restraint imposed by caregivers shape early development.
 

### 2025-11-16 manuscript_restrained_classification validation.qmd
updated abstract and fixed typo

@@
 abstract: "Physical restraint—including being held, carried, and restrained in devices—is a common feature of infants’ everyday lives.
-This study developed and validated a machine learning model to detect infants’ restraint using wearable sensors movement data derived from an existing data set (Franchak, 2023).
-The model achieved high accuracy (89%) and substantial kappa agreement (0.73) compared with human-coded ground truth.
+However, previous survey-based and video-based methods cannot simultaneously provide continuous, full-day accounts of infant restraint experiences.
+This study developed and validated a machine learning model to quantify infants’ restraint time moment-to-moment across the full day in the home environment using wearable sensors movement data derived from an existing data set.
+The data set includes 142 home-visit sessions from 65 infants, with 29 younger infants aged 4-7 months, and 36 older infants aged 11-14 months.
+We annotated infants’ restrained states in the first 1.5-hour video recordings of each session as ground truth labels.
+The supervised machine-learning model achieved high accuracy (89%) and substantial kappa agreement (κ = .73) compared with human-coded ground truth.
 The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when using a longer data-aggregating window length.
-The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infant restraint time.
+The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infant overall restraint time throughout the day.
 In short, the current study demonstrated the feasibility of using wearable sensors to quantify infants’ real-world restraint experiences, offering a new tool for studying how daily restraint influences early development."
@@
 Infants are routinely held or carried by caregivers, and infants are often restrained in devices such as highchairs, walkers, and strollers [@Birken2015s].
-As an externally regulated aspect of infants’ daily life, restraint shapes infants’ experiences based on caregivers’ decisions.
-Restraint limits infants’ physical activities and therefore hinders infants’ motor development [e.g., @Pin2007z].
-Restraint also frames infants’ accessible physical environment by altering experiences that are relevant for perceptual and cognitive development.
+As an externally regulated aspect of infants’ daily life, restraint shapes infants’ experiences based on caregivers’ decisions--caregivers, not infants, choose when infants are restrained.
+Restraint limits infants’ physical activities and might hinder infants’ motor development [e.g., @Pin2007z].
+Restraint also alters experiences that are relevant for perceptual and cognitive development by determining what is accessible in the physical world.
 For example, restraint is associated with fewer object holding experiences [@Franchak2024l], reduced amount and quality of language input [@Malachowski2023a], and increased opportunities for joint attention [@Karasik2025h].
@@
 Some used video recording and human annotation [e.g., @Springfield2025p; @Wang2025v].
-Although surveys provide an overall profile of infants’ daily life, and video recordings provide continuous time series data, no method so far can simultaneously provide full-day and continuous data to account for infants’ restraint experiences.
+Although surveys provide an overall profile of infants’ daily life and video recordings provide continuous time series data, neither method can simultaneously provide full-day and continuous data to account for infants’ restraint experiences.
 In the current study, we propose that wearable sensors can be a solution for recording full-day, continuous movement behavior, and machine learning algorithms can be an accurate and efficient way to automatically classify wearable sensor data.
-
-Moreover, previous studies did not define restraint consistently.
-Some definitions only focused on caregiver restraint such as holding and carrying [e.g., @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q], some only focused on seating devices [e.g., @Callahan1997s], some looked at different surfaces (e.g., lap, cradle, and seat) infants are on [e.g., @Graciosa2024o], some focused broadly on containment equipment [e.g., @Bartlett2003w; @Carson2022n; @Hesketh2015i; @Karasik2022a], some combined holding and carrying with seating equipment [@Springfield2025p], and others covered a wider range of restraint like furniture placement, seating devices and holding [@Franchak2024l; @Malachowski2023a; @Siddicky2021c].
-The inconsistency in restraint definitions makes it difficult to compare results across studies.
-In the current study, we take a broader scope and consider any practices that limit infants’ freedom of movement as restraint.
-We do not pre-define a list of restraint practices but focus on infants’ real-time opportunities for movement by classifying restraint whenever infants cannot change body position or locomote freely.
-Quantifying both caregiver restraint (i.e., holding and carrying) and device restraint, we can get a whole picture of the external constraints that infants experience in daily life.
 
@@
 
-Despite the inconsistency in definitions, prior work converges to show that caregiver and device restraints are common in infant daily life, and the time infants spend restrained changes over development.
+Despite different approaches, prior works converge to show that restraint is common in infant daily life, and that the time infants spend restrained changes over development.
 In terms of caregiver restraint, 2- to 6-month-old infants spent around 40% of their awake time being held or carried [@Malachowski2023a; @Siddicky2020f].
@@
 Additionally, a study that compared across ages also confirmed that infants’ overall restrained time decreased from 2 to 6 months of age [@Carson2022n].
+
+Many other studies have documented some portion of infant restraint time, but they differed in restraint definitions, making it difficult to compare results across studies.
+Some definitions only focused on caregiver restraint such as holding and carrying [e.g., @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q], some only focused on seating devices [e.g., @Callahan1997s], some looked at different surfaces (e.g., lap, cradle, and seat) infants are on [e.g., @Graciosa2024o], some focused broadly on containment equipment [e.g., @Bartlett2003w; @Carson2022n; @Hesketh2015i; @Karasik2022a], some combined holding and carrying with seating equipment [@Springfield2025p], and others covered a wider range of restraint like furniture placement, seating devices and holding [@Franchak2024l; @Malachowski2023a; @Siddicky2021c].
+
+In the current study, we take a broader scope and consider any practices that limit infants’ freedom of movement as restraint.
+We do not pre-define a list of restraint practices but focus on infants’ real-time opportunities for movement by classifying restraint whenever infants cannot change body position or locomote freely.
+Quantifying both caregiver restraint (i.e., holding and carrying) and device restraint, we can get a whole picture of the external constraints that infants experience in daily life.
 
@@
 
-Video recording and human annotation overcomes the above drawbacks of survey-based methods.
-However, one of the limitations of video observation is being obtrusive.
+Video recording and human annotation overcomes the above drawbacks of survey-based methods, but video is subject to other limitations.
+One disadvantage of video observation is being obtrusive.
 The presence of a researcher in the room may affect the caregiver’s and infant’s behaviors.
@@
 Possible distinctions might be manifested in reduced range of motion or repetitive motion due to constraints, which requires high-precision sensors and sensitive models to detect.
-Second, there are lots of possible restraint forms infants can present—caregiver holding can be holding the infant up straight or holding in arms breastfeeding; seating restraint can happen in countless infant seats, which vary in affordances of moving [@Alghamdi2024d].
+Second, there are lots of possible restraint forms infants can present--a caregiver can hold an infant up straight or in arms while breastfeeding; seating restraint can happen in countless infant seats, which vary in affordances of moving [@Alghamdi2024d].
 Accordingly, the wearable sensor system should be robust enough to generalize across heterogeneous restraint scenarios.
@@
 With a fixed sampling rate, larger windows provide more information for the model to leverage to predict the current category.
-Prior studies have used window lengths ranging from 2.3 to 4 seconds for infant posture category [@Airaksinen2020j; @Franchak2024e], and from 1.15 to 10 seconds for infant holding and carrying detection [@Airaksinen2024z].
+Prior studies have used window lengths ranging from 2.3 to 4 seconds for infant posture detection [@Airaksinen2020j; @Franchak2024e], and from 1.15 to 10 seconds for infant holding and carrying detection [@Airaksinen2024z].
 Considering that transitions between restraint states are typically slower than transitions between postures, we set window lengths longer than 4 seconds.
-Theoretically, shorter windows have higher temporal resolution and are sensitive to abrupt changes, but less data might mean the model lacks context to make an accurate classification.
+Theoretically, shorter windows have higher temporal resolution and are sensitive to abrupt changes, but fewer samples in a shorter window might mean the model lacks context to make an accurate classification.
 In contrast, longer windows can incorporate more data but may lack temporal sensitivity or may introduce irrelevant data from heterogeneous movement within an event.
@@
 
-The data set is hosted on Databrary (Franchak, 2023) and has been used in other publications.
-The data set consists of two cohorts of infants: `r apa_num(age[1,2])` younger infants aged from 4 to 7 months of age, and `r apa_num(age[2,2])` older infants aged from 11 to 14 months of age.
+The data set is hosted on Databrary (Franchak, 2023).
+The data set consists of two cohorts of infants: `r apa_num(age[1,2])` younger infants aged 4 to 7 months, and `r apa_num(age[2,2])` older infants aged 11 to 14 months.
 Among the total of `r apa_num(age[1,2]+age[2,2])` infants, 26 of them participated in the study for only once, while 39 of them participated from 2 to 4 times, resulting in a total of 142 sessions.
@@
 Restrained and unrestrained was defined as a binary category, meaning that all times were coded as one of the two categories.
-A period was coded as missing if the infant went out of view and was excluded from further analysis.
+A period was considered missing if the infant went out of view; missing periods were excluded from further analysis.
 Restraint was defined as when infants could not initiate or control their own movements, which included being held or carried by caregivers, or being restrained in devices that constrain them from changing body position.
@@
 For example, an infant was considered restrained when strapped into a high chair seat, but not when they were placed in a crib and free to change position.
-Each video file was coded by two coders: a primary coder annotated the whole 1.5-hour video, and an independent reliability coder annotated the first 30 minutes of the video.
+Each video file was coded by two coders: a primary coder annotated the whole 1.5-hour video, and an independent reliability coder annotated the first 30 minutes of each video.
 Inter-rater reliability was calculated as the proportion of video frames where the two coders coded the same label.
@@
 We aggregated raw IMU data into motion features within sliding windows of *n* seconds, where the window length *n* varied from 4 s to 16 s and 30 s.
+Motion features are a set of statistics to describe a window of data that serve as predictors for the machine learning model.
 For each window, we computed statistical features—mean, median, minimum, 25th percentile, 75th percentile, maximum, skewness, kurtosis, standard deviation, and sum—for all six signals across the four sensors, resulting in 240 features.
@@
 Consequently, each time point spaced every *n*/2 seconds contains 436 motion features, aggregated from the *n*/2 seconds before and after that point.
-For model training and testing, we included only time windows where the infant remained in a single status, so that each window could be assigned one unambiguous label.
+For model training and testing, we included only windows where the infant remained in a single status for over 90% of the time and labeled the window with that status.
 
@@
 We applied the random forest algorithm (Breiman, 2001) in modeling using the *DecisionTree* package (Sadeghi et al., 2022) in Julia.
-We used a leave-one-out cross-validation approach, where we held out one session at each iteration as testing set and trained the model using the other sessions,to evaluate how well the model generalized to unseen data.
+We used a leave-one-out cross-validation approach, where we held out one session at each iteration as testing set and trained the model using the other sessions, to evaluate how well the model generalized to unseen data.
 After validation, an optimal window length was picked[,]{.underline} and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.
 
-## Data Sharing
+## Reproducibility and Data Sharing
 
 Machine-learning process was conducted in Julia version 1.11.1 using the DecisionTree package (Sadeghi et al., 2022).
-Model evaluation, statistical analyses, and this reproducible manuscript were conducted in RStudio environment (R version 4.5.0; RStudio version 2025.5.0.496; RStudioTeam, 2024) using the R packages tidyverse [@tidyverse], flextable [@flextable], and papaja [@papaja], as well as the apa quarto template (Schneider, n.d.).
-All human annotation results and processing scripts are publicly shared on GitHub (https://github.com/Hanzhi-W/ms_restrained_classification-validation.git).
+This reproducible manuscript, including the model evaluation, figures, and statistical analyses, was made with Quarto version 1.5.57 (Allaire et al., 2024) using R version 4.5.0 (RStudioTeam, 2024) with the R packages tidyverse [@tidyverse], flextable [@flextable], and papaja [@papaja], as well as the apa quarto template (Schneider, n.d.).
+Human annotation results, which were used for model training and validation, together with processing scripts are publicly shared on OSF (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
 
@@
 
-We first compared the model’s performance to the human annotation of the 1.5-hour video recorded period across all 142 sessions using leave-one-out cross-validation approach.
-We reported performance metrics including accuracy, kappa, sensitivity, positive predictive value, F1, and the correlation of overall estimate time of restraint between the model and human annotation.
+We first compared model performance to the human annotations of the 1.5-hour video recorded period across all 142 sessions using leave-one-out cross-validation approach.
+We calculated performance metrics including accuracy, kappa, sensitivity, positive predictive value, F1, and the correlation of overall estimate time of restraint between the model and human annotation.
 We also examined prediction biases model and how they vary across window lengths.
@@
 ```{r tbl-indexbetweenmodels}
-#| tbl-cap: Performance metrics according to window length
+#| tbl-cap: Performance metrics according to window length (4 s, 16 s, and 30 s)
 #| ft.align: left
@@
 ```{r fig-prevalencegap}
-#| fig-cap: Restraint prevalence 
+#| fig-cap: Agreement between model-predicted and human annotated prevalence of restraint
 #| fig-width: 8
@@
 
-@fig-prevalencegap A shows that, at session level, all window lengths provided reliable estimate of restrained time.
+@fig-prevalencegap A shows that, at the session level, all window lengths provided reliable estimate of restrained time.
 We defined prevalence as the proportion of windows classified as restrained among all windows for each session.
 The prevalence predicted by the model was highly correlated with the prevalence annotated by human coders (*r* = `r apa_num(table_prev[1,4])` for 4 s, *r* = `r apa_num(table_prev[2,4])` for 16 s, and *r* = `r apa_num(table_prev[3,4])` for 30 s).
+Thus, model predictions are suitable for capturing individual differences in restrained time between infants.
 
@@
 
-Despite the overall good accuracy, models using all three window lengths were biased towards predicted that infants were unrestrained rather than restrained.
-@fig-cibtwctg shows the difference of a metric’s value (e.g., sensitivity, PPV) between restrained class and unrestrained class within a single window length.
+Despite the overall strong accuracy, models using all three window lengths were biased towards predicting that infants were unrestrained rather than restrained.
+@fig-cibtwctg shows the difference of a sensitivity and PPV between restrained class and unrestrained class within a single window length.
 Larger difference scores (away from 0) indicate larger asymmetry in predicting one category rather than the other.
 @fig-cibtwctg A shows that the sensitivity difference for 4 s window was `r apa_num(table_btw_ctg[1,3])`, 95% bootstrapped CI \[`r apa_num(table_btw_ctg[1,4])`, `r apa_num(table_btw_ctg[1,5])`\], for 16 s window was `r apa_num(table_btw_ctg[3,3])`, 95% CI \[`r apa_num(table_btw_ctg[3,4])`, `r apa_num(table_btw_ctg[3,5])`\], and for 30 s window was `r apa_num(table_btw_ctg[5,3])`, 95% CI \[`r apa_num(table_btw_ctg[5,4])`, `r apa_num(table_btw_ctg[5,5])`\].
-The negative differences of sensitivity across all window lengths indicate that the model was more sensitive in detecting the unrestrained class compared to restrained class regardless of window lengths.
+The negative difference scores for sensitivity across all window lengths indicate that the model was more sensitive in detecting the unrestrained class compared to restrained class regardless of window length.
 @fig-cibtwctg B shows that the PPV difference for 4 s window was `r apa_num(table_btw_ctg[2,3])`, 95% CI \[`r apa_num(table_btw_ctg[2,4])`, `r apa_num(table_btw_ctg[2,5])`\], for 16 s window was `r apa_num(table_btw_ctg[4,3])`, 95% CI \[`r apa_num(table_btw_ctg[4,4])`, `r apa_num(table_btw_ctg[4,5])`\], and for 30 s window was `r apa_num(table_btw_ctg[6,3])`, 95% CI \[`r apa_num(table_btw_ctg[6,4])`, `r apa_num(table_btw_ctg[6,5])`\].
@@
 @fig-prevalencegap B shows that the model predictions consistently underestimated overall restrained time compared to human annotation, despite the high correlation between the two.
-We defined prediction error as the model predicted prevalence (in % units) subtracted by the human annotated prevalence of restrained periods (in % units).
+We defined prediction error as the model predicted prevalence (in units of % of recording time) subtracted by the human annotated prevalence of restrained periods (in units of % of recording time).
 The 4 s window had a prediction error of `r apa_num(table_prev[1,5]*100)`%, 95% CI \[`r apa_num(table_prev[1,6]*100)`%, `r apa_num(table_prev[1,7]*100)`%\], the 16 s window of `r apa_num(table_prev[2,5]*100)`%, 95% CI \[`r apa_num(table_prev[2,6]*100)`%, `r apa_num(table_prev[2,7]*100)`%\], and the 30 s window of `r apa_num(table_prev[3,5]*100)`%, 95% CI \[`r apa_num(table_prev[3,6]*100)`%, `r apa_num(table_prev[3,7]*100)`%\].
 
-Across the metrics of sensitivity, PPV, and prediction error, 30 s turned out to be the best window length out of the three, which brought about a balanced performance across classes.
+Across the metrics of sensitivity, PPV, and prediction error, 30 s turned out to be the best window length out of the three, which brought about the most balanced performance across restrained and unrestrained classes.
 
@@
 ```{r fig-agechange}
-#| fig-cap: Younger versus older infants’ full day restrained time (as a proportion of recording time) 
+#| fig-cap: Younger versus older infants’ full day restrained time (as a proportion of awake time) 
 #| fig-width: 5
@@
 
-We applied the best model based on the analyses in the previous (30-s window length) to create full-day predictions for all sessions that included full-day experiences.
-We excluded eleven sessions whose total recording time was less than 3 hours, which were useful for training and validating the models but insufficient for measuring day-long experience.
+We applied the best model based on the analyses in the previous (30-s window length) to create full-day predictions for all sessions that included full-day recording.
+We excluded eleven sessions whose recording time was less than 3 hours per session, which were useful for training and validating the models but insufficient for measuring day-long experience.
 We also excluded another two sessions where the family took the garment off for a substantial portion of time but did not log the time stamp of removal.
@@
 The bias might be due to the inherent imbalance in the training data set where unrestrained periods were more prevalent than restrained periods, indicated by human annotated prevalence.
-However, using a longer window length (i.e., 30 s, rather than 16 s or 4 s) mitigated the prediction bias.
+Using a longer window length (i.e., 30 s, rather than 16 s or 4 s) mitigated but did not eliminate the prediction bias.
 Ultimately, the model with 30 s window length presented convergent validity in overall estimate of restraint time and age-related change compared with previous findings [@Carson2022n].
@@
 Why could a longer window length (i.e., 30 s) provide more balanced detection?
-For one thing, a longer window length can better capture the contextual information of restraint.
+One explanation is that a longer window length can better capture the contextual information of restraint.
 Consider the contextual nature of restraint: infants may exhibit similar body movement whether restrained or not for most of the time, but one subtle occurrence of distinctive movement can signal restraint.
@@
 It is hard to distinguish if the infant is restrained while sitting still, but when the infant moves their limbs, restraint may present differences, such as a stronger damping in the motion decay because the support (i.e., caregiver’s laps, or seats) absorbs the motion.
-These differences only present themselves in the moments of movement, not throughout the whole period.
-Thus, a longer window length can better capture temporal information, whereas shorter windows only capture momentary movements that may appear similar across restrained and unrestrained periods.
-For another, a longer window length can improve generalization across a variety of restraint forms.
+These differences only present themselves in the episodes of movement, not throughout the whole period.
+A longer window length has a better chance to capture the episodes.
+Another explanation is that a longer window length can improve generalization across a variety of restraint forms.
 Infants can be restrained in various ways—such as in stationary seats, in jumpers, in swings, or in caregivers’ arms, either still or while being carried around.
@@
 
-Our analysis of window length informs future modeling works to consider temporal resolution as a behavior-informed decision.
+Our analysis of window length informs future modeling studies to consider temporal resolution as a behavior-informed decision.
 In infants’ everyday activities, behaviors unfold across various timescales—from second-by-second movement (e.g., locomotion) to longer periods of states (e.g., on the ground vs.
@@
 With the machine learning model established, researchers can extend their investigation timescales to a full day or even multiple days at low cost.
-Although the current machine learning model achieves slightly lower accuracy (89%) than human annotation, the trade-off between accuracy and human labor favors our sensor approach—especially when the target behavior fluctuates over longer timescales beyond the practical limits of human coding.
+Although the current machine learning model achieves slightly lower accuracy (89%) than human annotation (99%), the trade-off between accuracy and human labor favors our sensor approach—especially when the target behavior fluctuates over longer timescales beyond the practical limits of human coding.
 Restraint is such a behavior that does not recur repeatedly every hour but follows the rhythm of everyday routines and activities.
@@
 The current work also demonstrates that wearable sensors have the potential to characterize other categories of infant movement that might interest researchers in a variety of domains.
-Compared with previous works that classify infants’ positions [@Airaksinen2020j; @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q], restraint might be the most challenging one out of all categories, because the motion between classes can be very similar sometimes—such as sitting on the ground versus sitting on a caregiver’s lap.
+Compared with previous works that classify infants’ positions [@Airaksinen2020j; @Airaksinen2024z; @Franchak2021k; @Franchak2024e; @Yao2019q], restraint might be the most challenging one out of all categories, because the motion between classes can be very similar—such as sitting on the ground versus sitting on a caregiver’s lap.
 In contrast, positions usually have distinctive motion features between classes—such as the legs’ moving direction in supine position versus upright position.
-Our results, however, showed that even a context-dependent category as restraint can be successfully classified by machine learning models using sensor data.
+Our results, however, showed that even a context-dependent category such as restraint can be successfully classified by machine learning models using sensor data.
 Conceivably, this system might perform comparably well or better in classifying other context-independent categories, such as locomotion, steps, falling, etc.
@@
 Infants’ outdoor experiences remain largely underexplored, whereas outdoor environments provide infants with rich and unique opportunities for learning and exploration [@Schneider2025d; @Herrington1998o; @Fjortoft2000o].
-How much time infants spend restrained in stroller or car seat?
-Which positions are they frequently in?
-How many steps they take?
+How much time do infants spend restrained in strollers or car seats and while worn by parents in wearable carriers?
+Which positions are they frequently in when restrained out of the house?
 Validating sensor-based models in outdoor scenarios can open new avenues for investigation.
@@
 Prior works provided valuable insights into parental attitudes of restraint and physical activity.
-For example, parental concerns over floor play, beliefs about physical activeness, and practices of co-play are found to be associated with restraint frequency and infants’ physical activity level [@Hanisch2025j; @Hnatiuk2013t; @Prioreschi2017c; @Hesketh2015i; @Karasik2025h].
+For example, parental concerns over floor play, beliefs about physical activity, and practices of co-play are found to be associated with restraint frequency and infants’ physical activity level [@Hanisch2025j; @Hnatiuk2013t; @Prioreschi2017c; @Hesketh2015i; @Karasik2025h].
 However, few studies have examined how each restraint decision is formed in real time, and how infants’ sensorimotor experiences are affected in real time.
@@
 [@Birken2015s], but these functions alone cannot account for all the variability in when caregivers choose to restrain an infant.
-To what extent are such decisions influenced by these intra-individual factors, such as daily routines, or momentary tasks?
 To what extent are such decisions influenced by inter-individual factors, such as infant motor skill, attachment style, caregivers’ beliefs, or parental stress?
@@
 A prior study [@Malachowski2023a] found device restraint is associated with reduced adult language input using ecological momentary assessment.
-However, a continuous, full-day recording of restraint can capture the temporal distribution rather than snapshots.
+However, a continuous, full-day recording of restraint and speech can capture the temporal distribution rather than snapshots.
 Questions can be answered such as: Does adult word count vary based on the duration of each restraint bout?
@@
 
+# Declarations
+## Funding
+This work was supported by National Science Foundation Grant BCS #1941449.
+
+## Conflicts of interest
+The authors report no conflicts of interest related to this article.
+
+## Ethics approval
+All study procedures were reviewed and approved by the Institutional Review Board at the University of California, Riverside (Protocol HS-15-050).
+
+## Consent to participate
+Written informed consent was obtained from all caregivers before the study began.
+
+## Consent for publication
+Additional written consent was obtained from caregivers to allow the sharing of audio and video data.
+
+## Availability of data and materials
+Human annotation results, which were used for model training and validation are publicly shared on OSF (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
+
+## Code availability
+All the processing scripts are also publicly shared on OSF.
+
 # References

### 2025-11-17 manuscript_restrained_classification validation.qmd
fixed typo

@@
 However, previous survey-based and video-based methods cannot simultaneously provide continuous, full-day accounts of infant restraint experiences.
-This study developed and validated a machine learning model to quantify infants’ restraint time moment-to-moment across the full day in the home environment using wearable sensors movement data derived from an existing data set.
+This study developed and validated a machine learning model to quantify infants’ restraint time moment-to-moment across the full day in the home environment using wearable sensor movement data derived from an existing data set.
 The data set includes 142 home-visit sessions from 65 infants, with 29 younger infants aged 4-7 months, and 36 older infants aged 11-14 months.
@@
 The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when using a longer data-aggregating window length.
-The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infant overall restraint time throughout the day.
+The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infants' overall restraint time throughout the day.
 In short, the current study demonstrated the feasibility of using wearable sensors to quantify infants’ real-world restraint experiences, offering a new tool for studying how daily restraint influences early development."
@@
 Restraint also alters experiences that are relevant for perceptual and cognitive development by determining what is accessible in the physical world.
-For example, restraint is associated with fewer object holding experiences [@Franchak2024l], reduced amount and quality of language input [@Malachowski2023a], and increased opportunities for joint attention [@Karasik2025h].
+For example, restraint is associated with fewer object holding experiences [@Franchak2024l], reduced amount of language input [@Malachowski2023a], and increased opportunities for joint attention [@Karasik2025h].
 Therefore, characterizing infants’ restrained experiences helps profile infants’ daily learning contexts and situate their motor, perceptual, and cognitive development within everyday experiences.
@@
 Infants produce fewer object interactions when restrained compared to when they are unrestrained [@Franchak2024l].
-Moreover, infants hear less amount of and less consistent adult language when restrained compared to unrestrained [@Malachowski2023a].
+Moreover, infants hear less frequent and less consistent adult language when restrained compared to unrestrained [@Malachowski2023a].
 
@@
 
-Video recording and human annotation overcomes the above drawbacks of survey-based methods, but video is subject to other limitations.
+Video recording and human annotation overcome the above drawbacks of survey-based methods, but video is subject to other limitations.
 One disadvantage of video observation is being obtrusive.
@@
 Within the short duration of recording, researchers tend to record the period when infants are awake and active.
-In contrast, periods when the infants are restrained, like seated in front of TV or in highchairs, would likely be skipped.
+In contrast, periods when the infants are restrained, like seated in front of the TV or seated in highchairs, would likely be skipped.
 Thus, video recording is subject to sampling bias and might not be representative of infants’ whole-day experiences.
@@
 Third, sensors embedded in baby garments are mobile and unobtrusive—an experimenter does not need to follow the infant with a camera to record movements.
-Fourth, the system can record and process full-day or multiple-day data at a time due to the sensors’ sufficient battery life and automated classification of machine learning models.
+Fourth, the system can record and process full-day and multiple-day data due to the sensors’ sufficient battery life and automated classification of machine learning models.
 
@@
 For example, infants might kick legs when they are lying without restraint and also when they are strapped in a highchair.
-Possible distinctions might be manifested in reduced range of motion or repetitive motion due to constraints, which requires high-precision sensors and sensitive models to detect.
+Possible distinctions might be manifested in a reduced range of motion or repetitive motion due to constraints, which requires high-precision sensors and sensitive models to detect.
 Second, there are lots of possible restraint forms infants can present--a caregiver can hold an infant up straight or in arms while breastfeeding; seating restraint can happen in countless infant seats, which vary in affordances of moving [@Alghamdi2024d].
@@
 The data set consists of two cohorts of infants: `r apa_num(age[1,2])` younger infants aged 4 to 7 months, and `r apa_num(age[2,2])` older infants aged 11 to 14 months.
-Among the total of `r apa_num(age[1,2]+age[2,2])` infants, 26 of them participated in the study for only once, while 39 of them participated from 2 to 4 times, resulting in a total of 142 sessions.
+Among the total of `r apa_num(age[1,2]+age[2,2])` infants, 26 of them participated in the study only once, while 39 of them participated from 2 to 4 times, resulting in a total of 142 sessions.
 Thirty-four of the infants are female and thirty-one are male.
@@
 Each sensor collects raw data as accelerometer signals (in x, y, and z orientations) and gyroscope signals (in roll, pitch, and yaw orientations).
-The sampling rate of 62.5 Hz allows for high resolution motion recording, and the sufficient battery life and built-in storage allow for whole-day-long recording.
+The sampling rate of 62.5 Hz allows for high resolution motion recording, and the sufficient battery life and built-in storage allow for daylong recording.
 An action camera fixed on a mini tripod (GoPro HERO9 Black) was used to record synchronized third-person video of the first 1.5 hours into the session as ground truth.
@@
 Using the 1.5-hour videos, our team annotated the periods when the infant was restrained or unrestrained using Datavyu software (https://datavyu.org/).
-Restrained and unrestrained was defined as a binary category, meaning that all times were coded as one of the two categories.
+Restrained and unrestrained were defined as a binary category, meaning that all times were coded as one of the two categories.
 A period was considered missing if the infant went out of view; missing periods were excluded from further analysis.
@@
 We applied the random forest algorithm (Breiman, 2001) in modeling using the *DecisionTree* package (Sadeghi et al., 2022) in Julia.
-We used a leave-one-out cross-validation approach, where we held out one session at each iteration as testing set and trained the model using the other sessions, to evaluate how well the model generalized to unseen data.
+We used a leave-one-out cross-validation approach, where we held out one session at each iteration as a testing set and trained the model using the other sessions, to evaluate how well the model generalized to unseen data.
 After validation, an optimal window length was picked[,]{.underline} and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.
@@
 A negative prediction error indicates the model estimated less amount of time infants were restrained compared to human annotation.
-The horizontal line in the box denotes the median; the boarder lines of the box denotes the first (Q1) quartile and the third (Q3) quartile; and the lines extended denotes the min and max data points within the range of 1.5 IQR (i.e., Q3-Q1).
+The horizontal line in the box denotes the median; the borderlines of the box denotes the first (Q1) quartile and the third (Q3) quartile; and the lines extended denotes the min and max data points within the range of 1.5 IQR (i.e., Q3-Q1).
 
@@
 
-@fig-prevalencegap A shows that, at the session level, all window lengths provided reliable estimate of restrained time.
+@fig-prevalencegap A shows that, at the session level, all window lengths provided reliable estimates of restrained time.
 We defined prevalence as the proportion of windows classified as restrained among all windows for each session.
@@
 
-### Models were biased towards predicted unrestrained
+### Models Were Biased Toward Predicted Unrestrained
 
 Despite the overall strong accuracy, models using all three window lengths were biased towards predicting that infants were unrestrained rather than restrained.
-@fig-cibtwctg shows the difference of a sensitivity and PPV between restrained class and unrestrained class within a single window length.
+@fig-cibtwctg shows the difference of sensitivity and PPV between restrained class and unrestrained class within a single window length.
 Larger difference scores (away from 0) indicate larger asymmetry in predicting one category rather than the other.
@@
 
-We applied the best model based on the analyses in the previous (30-s window length) to create full-day predictions for all sessions that included full-day recording.
-We excluded eleven sessions whose recording time was less than 3 hours per session, which were useful for training and validating the models but insufficient for measuring day-long experience.
+We applied the best model based on the analyses in the previous section (30-s window length) to create full-day predictions for all sessions that included full-day recording.
+We excluded eleven sessions whose recording time was less than 3 hours, which were useful for training and validating the models but insufficient for measuring day-long experience.
 We also excluded another two sessions where the family took the garment off for a substantial portion of time but did not log the time stamp of removal.
@@
 Consider the contextual nature of restraint: infants may exhibit similar body movement whether restrained or not for most of the time, but one subtle occurrence of distinctive movement can signal restraint.
-For example, an infant can be sitting on the floor unrestrained, on caregiver’s lap restrained, or strapped in a seat restrained.
+For example, an infant can be sitting on the floor unrestrained, on a caregiver’s lap restrained, or strapped in a seat restrained.
 It is hard to distinguish if the infant is restrained while sitting still, but when the infant moves their limbs, restraint may present differences, such as a stronger damping in the motion decay because the support (i.e., caregiver’s laps, or seats) absorbs the motion.
@@
 Another explanation is that a longer window length can improve generalization across a variety of restraint forms.
-Infants can be restrained in various ways—such as in stationary seats, in jumpers, in swings, or in caregivers’ arms, either still or while being carried around.
+Infants can be restrained in various ways—such as in stationary seats, jumpers, swings, or caregivers’ arms, either still or while being carried around.
 Different restraint forms are presented by different motion features.
@@
 Infants’ outdoor experiences remain largely underexplored, whereas outdoor environments provide infants with rich and unique opportunities for learning and exploration [@Schneider2025d; @Herrington1998o; @Fjortoft2000o].
-How much time do infants spend restrained in strollers or car seats and while worn by parents in wearable carriers?
+How much time do infants spend restrained in strollers, car seats, or in carriers worn by caregivers?
 Which positions are they frequently in when restrained out of the house?
@@
 
-## Advances in Understanding Full-day Real-time Restraint
+## Advances in Understanding Full-Day Real-Time Restraint
 
@@
 Having a full-day continuous profile of infant restraint is also essential in understanding infant real-time sensorimotor experiences.
-An overall estimate of restraint time is not enough to examine the mechanisms between restraint practices and development outcomes.
+An overall estimate of restraint time is not enough to examine the mechanisms between restraint practices and developmental outcomes.
 Take motor development as an example.
@@
 Questions can be answered such as: Does adult word count vary based on the duration of each restraint bout?
-How do specific language inputs (e.g., referential words) distribute throughout a day based on restraint states and restraint types?
+How are specific language inputs (e.g., referential words) distributed throughout a day based on restraint states and restraint types?
 Additionally, real-time restraint can also be linked to infants’ self-initiated explorations.

### 2025-11-21 manuscript_restrained_classification validation.qmd
fixed typo

@@
 However, previous survey-based and video-based methods cannot simultaneously provide continuous, full-day accounts of infant restraint experiences.
-This study developed and validated a machine learning model to quantify infants’ restraint time moment-to-moment across the full day in the home environment using wearable sensor movement data derived from an existing data set.
-The data set includes 142 home-visit sessions from 65 infants, with 29 younger infants aged 4-7 months, and 36 older infants aged 11-14 months.
+This study developed and validated a machine learning model to quantify infants’ restraint time moment-to-moment across the full day in the home environment using wearable sensor data.
+We used a data set that includes 142 home-visit sessions from 65 infants, with 29 younger infants aged 4-7 months, and 36 older infants aged 11-14 months.
 We annotated infants’ restrained states in the first 1.5-hour video recordings of each session as ground truth labels.
 The supervised machine-learning model achieved high accuracy (89%) and substantial kappa agreement (κ = .73) compared with human-coded ground truth.
-The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when using a longer data-aggregating window length.
+The model presented a slight bias towards overestimating unrestrained periods compared to restrained periods, but the bias was mitigated when we used a longer data-aggregating window length.
 The model also showed convergent validity by corroborating prior studies that showed an age-related decrease in infants' overall restraint time throughout the day.
@@
 keywords: [infant, restraint, everyday experiences, wearable sensors, machine learning]
-bibliography: [references.bib]
+bibliography: [references.bib, packages.bib]
 format:
@@
 Restraint also alters experiences that are relevant for perceptual and cognitive development by determining what is accessible in the physical world.
-For example, restraint is associated with fewer object holding experiences [@Franchak2024l], reduced amount of language input [@Malachowski2023a], and increased opportunities for joint attention [@Karasik2025h].
+For example, restraint is associated with fewer object holding experiences [@Franchak2024l], reduced amounts of language input [@Malachowski2023a], and increased opportunities for joint attention [@Karasik2025h].
 Therefore, characterizing infants’ restrained experiences helps profile infants’ daily learning contexts and situate their motor, perceptual, and cognitive development within everyday experiences.
@@
 
-Despite different approaches, prior works converge to show that restraint is common in infant daily life, and that the time infants spend restrained changes over development.
+Despite different approaches, prior work converges to show that restraint is common in infant daily life, and that the time infants spend restrained changes over development.
 In terms of caregiver restraint, 2- to 6-month-old infants spent around 40% of their awake time being held or carried [@Malachowski2023a; @Siddicky2020f].
 Older 10- to 13-month-olds spent around 19% of their awake time being held or carried [@Franchak2024l].
-In terms of device restraint, 4- to 6-month-olds spent around 20% of awake time in seating devices, and 35% in placement with physical support [@Callahan1997s; @Malachowski2023a].
+In terms of device restraint, 4- to 6-month-olds spent around 20% of their awake time in seating devices, and 35% in placement with physical support [@Callahan1997s; @Malachowski2023a].
 Older 10- to 13-month-olds spent around 27% of awake time in restraining furniture or devices [@Franchak2024l].
@@
 However, this method might still be sensitive to memory loss and thus not as accurate in providing data.
-EMA survey techniques reduce caregivers’ memory burden by making momentary phone-survey prompts to caregivers multiple times (e.g., 5, 10, or 12 times) randomly throughout a day [@Franchak2019o; @Franchak2024l; @Malachowski2023a].
+EMA survey techniques reduce caregivers’ memory burden by sending momentary phone-survey prompts to caregivers multiple times (e.g., 5, 10, or 12 times) randomly throughout a day [@Franchak2019o; @Franchak2024l; @Malachowski2023a].
 Despite the benefits of gathering immediate observations rather than retrospective reports, EMA cannot provide continuous time-series data.
@@
 The presence of a researcher in the room may affect the caregiver’s and infant’s behaviors.
-Another limitation is that video recording and video annotation are both labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies.
+Another limitation is that video recording and annotation are both labor-intensive and time-consuming, which limits its use in large-scale or long-duration studies.
 Indeed, prior work that videotaped and manually annotated infants’ restraint had short durations of recording (e.g., half an hour in @Springfield2025p, and one hour in @Wang2025v).
@@
 
-However, one of the challenges researchers continue to face is the reduced model performance in naturalistic home settings [@Airaksinen2024z; @Franchak2024e; @Yao2019q], compared to more controlled laboratory environments [@Airaksinen2020j; @Franchak2021k].
+However, one of the challenges researchers continue to face is the reduced model performance in naturalistic home settings [@Airaksinen2024z; @Franchak2024e; @Yao2019q], compared with more controlled laboratory environments [@Airaksinen2020j; @Franchak2021k].
 Indeed, infants may exhibit more complex and unexpected movements at home versus in lab, which requires the models to be able to generalize to scenarios that go beyond the models’ training data set.
@@
 The third step was to train and validate supervised machine learning models that classify the motion features into restrained status.
-We applied the random forest algorithm (Breiman, 2001) in modeling using the *DecisionTree* package (Sadeghi et al., 2022) in Julia.
+We applied the random forest algorithm (Breiman, 2001) in modeling using the DecisionTree package (Sadeghi et al., 2022) in Julia.
 We used a leave-one-out cross-validation approach, where we held out one session at each iteration as a testing set and trained the model using the other sessions, to evaluate how well the model generalized to unseen data.
-After validation, an optimal window length was picked[,]{.underline} and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.
+After validation, an optimal window length was picked, and a final model was trained using all the human annotations to process whole-day sensor data of all sessions.
 
@@
 Machine-learning process was conducted in Julia version 1.11.1 using the DecisionTree package (Sadeghi et al., 2022).
-This reproducible manuscript, including the model evaluation, figures, and statistical analyses, was made with Quarto version 1.5.57 (Allaire et al., 2024) using R version 4.5.0 (RStudioTeam, 2024) with the R packages tidyverse [@tidyverse], flextable [@flextable], and papaja [@papaja], as well as the apa quarto template (Schneider, n.d.).
+This reproducible manuscript, including the model evaluation, figures, and statistical analyses, was made with Quarto version 1.5.57 (Posit, PBC, 2024) using R version 4.5.0 [@R-base] with the R packages tidyverse [@R-tidyverse], flextable [@R-flextable], and papaja (Aust & Barth, 2024), as well as the apa quarto template (Schneider, n.d.).
 Human annotation results, which were used for model training and validation, together with processing scripts are publicly shared on OSF (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
@@
 #| fig-height: 5
-#| apa-note: Each vertical line indicates the 95%CI of the metric's difference between restrained prediction and unrestrained prediction.
+#| apa-note: Each vertical line indicates the 95% CI of the metric's difference between restrained prediction and unrestrained prediction.
 The dot on each line denotes the mean of the difference (restrained - unrestrained).
@@
 All window lengths also had high F1 scores—a balanced measure that combines sensitivity and PPV—ranging from `r apa_num(min(table_btw_mdl[7,3:5]))` to `r apa_num(max(table_btw_mdl[7,3:5]))` for restrained prediction, and `r apa_num(min(table_btw_mdl[8,3:5]))` to `r apa_num(max(table_btw_mdl[8,3:5]))` for unrestrained prediction.
-Moreover, the three window lengths did not significantly differ for any of the metrics (*p*s \>= `r apa_num(min(table_btw_mdl[1:8,8]))`).
+Moreover, the three window lengths did not significantly differ for any of the metrics (*p*s $\ge$ `r apa_num(min(table_btw_mdl[1:8,8]))`).
 
@@
 We explored the window length in the data aggregating procedure of modeling and proved that a longer window enhanced the model’s balanced performance.
-Together, the current study advanced the method for measuring infant full-day, real-time movement, and opened avenues for research on how restraint imposed by caregivers shape early development.
+Together, the current study advanced the method for measuring infant full-day, real-time movement, and opened avenues for research on how restraint imposed by caregivers can shape early development.
 
 # Declarations
+
 ## Funding
+
 This work was supported by National Science Foundation Grant BCS #1941449.
 
-## Conflicts of interest
+## Conflicts of Interest
+
 The authors report no conflicts of interest related to this article.
 
-## Ethics approval
+## Ethics Approval
+
 All study procedures were reviewed and approved by the Institutional Review Board at the University of California, Riverside (Protocol HS-15-050).
 
-## Consent to participate
+## Consent to Participate
+
 Written informed consent was obtained from all caregivers before the study began.
 
-## Consent for publication
+## Consent for Publication
+
 Additional written consent was obtained from caregivers to allow the sharing of audio and video data.
 
-## Availability of data and materials
+## Availability of Data and Materials
+
 Human annotation results, which were used for model training and validation are publicly shared on OSF (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
 
-## Code availability
+## Code Availability
+
 All the processing scripts are also publicly shared on OSF.

### 2025-11-21 manuscript_restrained_classification validation.qmd
edited declarations

@@
 
-Human annotation results, which were used for model training and validation are publicly shared on OSF (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
+Sensor data and video recordings are publicly shared on Databrary (Franchak, 2023).
+Compiled human annotation results are shared on OSF (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
 
@@
 
-All the processing scripts are also publicly shared on OSF.
+All processing scripts, including model training and validation analyses, are publicly shared on OSF as well (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
 

### 2025-11-21 manuscript_restrained_classification validation.qmd
edited declarations
fixed typo

@@
 For infants, there have been established models that classify infants’ body postures in general—including holding and carrying [@Airaksinen2020j; @Franchak2021k; @Franchak2024e], and models that specifically detect holding and carrying periods [@Airaksinen2024z; @Yao2019q].
-Insights from prior studies suggest effective practices, such as using multiple sensors and using sensors that record both gyroscope and accelerator data [@Airaksinen2025l; @Franchak2024e].
+Insights from prior studies suggest effective practices, such as using multiple sensors and using sensors that record both gyroscope and accelerometer data [@Airaksinen2025l; @Franchak2024e].
 
@@
 Each sensor collects raw data as accelerometer signals (in x, y, and z orientations) and gyroscope signals (in roll, pitch, and yaw orientations).
-The sampling rate of 62.5 Hz allows for high resolution motion recording, and the sufficient battery life and built-in storage allow for daylong recording.
+The sampling rate of 62.5 Hz allows for high-resolution motion recording, and the sufficient battery life and built-in storage allow for daylong recording.
 An action camera fixed on a mini tripod (GoPro HERO9 Black) was used to record synchronized third-person video of the first 1.5 hours into the session as ground truth.
@@
 The caregiver was instructed to set up the camera to capture the main room and to move it as needed to ensure the infant remained in view.
-The caregiver put the garment on the infant and positioned the infant in various postures (including holding and sitting in a restraint device) for at least one minute each for model training purpose.
+The caregiver put the garment on the infant and positioned the infant in various postures (including holding and sitting in a restraint device) for at least one minute each for model training purposes.
 Afterwards, the infant continued wearing the garment during their daily routines.
@@
 
-The third step was to train and validate supervised machine learning models that classify the motion features into restrained status.
+The third step was to train and validate supervised machine learning models that classify the motion features into restrained statuses.
 We applied the random forest algorithm (Breiman, 2001) in modeling using the DecisionTree package (Sadeghi et al., 2022) in Julia.
@@
 Machine-learning process was conducted in Julia version 1.11.1 using the DecisionTree package (Sadeghi et al., 2022).
-This reproducible manuscript, including the model evaluation, figures, and statistical analyses, was made with Quarto version 1.5.57 (Posit, PBC, 2024) using R version 4.5.0 [@R-base] with the R packages tidyverse [@R-tidyverse], flextable [@R-flextable], and papaja (Aust & Barth, 2024), as well as the apa quarto template (Schneider, n.d.).
-Human annotation results, which were used for model training and validation, together with processing scripts are publicly shared on OSF (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
+This reproducible manuscript, including the model evaluation, figures, and statistical analyses, was made with Quarto version 1.5.57 (Posit, PBC, 2024) using R version 4.5.0 [@R-base] with the R packages tidyverse [@R-tidyverse], flextable [@R-flextable], and papaja (Aust & Barth, 2024), as well as the APA Quarto template (Schneider, n.d.).Human annotation results, processing scripts, and analyses codes are publicly shared on OSF (https://osf.io/a698g/overview?view_only=221d1f2bc8d34b5d9838354224d6da8f).
 
@@
 
-We first compared model performance to the human annotations of the 1.5-hour video recorded period across all 142 sessions using leave-one-out cross-validation approach.
-We calculated performance metrics including accuracy, kappa, sensitivity, positive predictive value, F1, and the correlation of overall estimate time of restraint between the model and human annotation.
-We also examined prediction biases model and how they vary across window lengths.
+We first compared model performance to the human annotations of the 1.5-hour video recorded period across all 142 sessions using the leave-one-out cross-validation approach.
+We calculated performance metrics, including accuracy, kappa, sensitivity, positive predictive value, F1, and the correlation of the overall estimate time of restraint between the model and human annotation.
+We also examined the prediction biases of the model and how they vary across window lengths.
 Second, we applied the model with the optimal window length to process infants’ full-day sensor data and examined the model’s convergent validity by replicating a previously-reported effect: that restraint time decreases with age [@Carson2022n].
@@
 
-This work was supported by National Science Foundation Grant BCS #1941449.
+This work was supported by National Science Foundation Grants BCS #1941449 and BCS #2521429.
 
